
%% \chapter{Overview}
\section{Overview}

Graphs, used in ML and other \textbf{scientific} domains, as well as \textbf{industrial} and \textbf{general} programming, 
do \textbf{not} presently exist in the C++ standard. In ML, a graph forms the underlying structure of an \textbf{artificial neural network} (ANN). 
In a \textbf{game}, a graph can be used to represent the \textbf{map} of a game world. In \textbf{business} environments, graphs arise as 
\textbf{entity relationship diagrams} (ERD) or \textbf{data flow diagrams} (DFD). In the realm of \textbf{social media}, a graph represents a 
\textbf{social network}.

All documents, taken as a whole for a Graph Library, proposes the addition of \textbf{graph algorithms, operators, views, adaptors}, the 
\textbf{graph container interface} and a \textbf{graph container implementation} to the C++ library to support \textbf{machine learning} (ML), 
as well as other applications. ML is a large and growing field, both in the \textbf{research community} and \textbf{industry}, that has 
received a great deal of attention in recent years. This documents presents an \textbf{interface} of the proposed algorithms, operators, adaptors,
views, graph functions and containers.

\section{Goals and Priorities}

Because graphs and their algorithms cover a broad range of capabilities and implementations, we have defined a focused set of 
goals and priorities that will provide an initial set of useful functionality, as well as a sound foundation for future work.

\begin{itemize}
      \item Provide a firm theoretical foundation for the library.
      \item Follow the separation of algorithms, ranges, views and containers established by the standard library.
      \item Include a rich enough set of algorithms for the library to be useful.
            \begin{itemize}
                  \item The syntax for an algorithm's implementation should be simple, expressive and easy to understand. 
                  \item The ability to write high-performance algorithms should not be compromised.
                  \item In the initial proposal, algorithms expect vertices to be in a random access range with an integral vertex id. 
            \end{itemize}
      \item Include views for common traversals of a graph's vertices and edges that is concise and consistant without having to use a lower level interface.
            \begin{itemize}
                  \item Simple views include vertexlist, incidence edges on a vertes, neighbors of a vertex, and edges of a graph.
                  \item Complex views include depth-first search, breath-first search, and topological sort.
            \end{itemize}
      \item A Graph Container Interface, used by Views and Algorithms, that provides a consistent interface for different graph data structures. The interface
            includes concepts, types, traits and functions and provides a similar role to the Ranges library for standard containers.
            \begin{itemize}
                  \item Descriptors that are used to project a consistent data model for vertex, edge and neighbor by views and edge lists.

                  \item Adjacency list, an outer range of vertices with an inner range of outgoing edges on each vertex.
                        \begin{itemize}
                              \item A set of customization point functions to access a graph data structure. 
                              \item Integration of existing graph data structures by overriding a small number of functions. Reasonable default implementations 
                                    are provided to reduce the work required. When using a standard container to define a graph, no additional work is required.
                              \item Support for optional user-defined value types on an edge, vertex and/or the graph itself.
                              \item Support bipartite and multipartite graphs.
                        \end{itemize}

                  \item Edge list, which is a range of edge descriptors.
                        \begin{itemize}
                              \item From an \tcode{edgelist} view.
                              \item From a user-defined range that uses \tcode{std::ranges::transform_view}.
                        \end{itemize}
            \end{itemize}
      \item A graph container that can be used with the algorithms.
            \begin{itemize}
                  \item A high-performance \tcode{compressed_graph} container, based on the Compressed Sparse Row matrix.
                  \item The ability to create simple graph container from standard containers, e.g. \tcode{vector<vector<int>>}.
            \end{itemize}
\end{itemize}

The design should not hinder the ability to extend the functionality to support expanded functionality identified in the future roadmap
that follows.

\subsection{Future Roadmap}

The following are areas we'd like to see in future proposals, after the initial proposal is accepted. We endeavor to investigate
these to assure the existing design will support them.
\begin{itemize}
      \item Additional graph algorithms. The Graph Algorithms paper identifies algorithms additional algorithms we'd like to see added.
      \item Parallel algorithms.
      \item Support for sparse vertex ids, implying the use of bi-directional containers such as \tcode{map} and \tcode{unordered_map} for vertices.
      \item Bi-directional graphs, where vertices have incoming and outgoing edges.
      \item Non-integral vertex ids.
      \item Constexpr graphs, where vertices and edges are stored in \tcode{std::array} or other const-friendly container.
\end{itemize}


\section{Examples}

%\andrew{Where do examples really belong?  In P2300 they are up front here, but I think there is too much forward referencing for that.}

The following code demonstrates how a simple graph can be created as a range of ranges, using the standard containers. 

\phil{Duplicated in Introduction. OK?}

{\small
  \lstinputlisting[firstline=26,lastline=48]{D9901/src/bacon.cpp}
}

\tcode{target_id(g,uv)} defines the required function to get a target\_id for an edge in the graph \tcode{G}. Other functions can also
be overridden to allow a developer to adapt their own graph data structures to the library.

\section{What this proposal is  \textbf{not}}

The Graph Library proposal limits itself to adjacency graphs and edgelists only. An adjacency graph is an outer range of vertices with an inner range of outgoing
edges on each vertex. An edgelist is a view of edges, which is either all the edges in the adjacency graph or a projection of a user-defined range.

Parallel versions of the algorithms are not included for several reasons. The executors proposal in P2300r5 \cite{REF_P2300r5} is expected to introduce new 
and better ways to do parallel algorithms beyond that used in the parallel STL algorithms and we would like to wait for finalization of that proposal before 
committing to parallel implementations. Secondly, many graph algorithms don't benefit from parallel implementations so there is less need to offer an implementation. 
Lastly, it will help limit the size of this proposal which is already looking to be large without it. It is expected that future proposals will be submitted for parallel graph algorithms. 

Incoming edges on a vertex are not included, though it is hoped that a future proposal will be made for them.

The algorithms and views in this proposal expect that vertex\_ids are densly assigned in a random access range, but it does not exclude the possibility of 
sparsely-defined vertex\_ids stored in containers like \lstinline{std::map} or \tcode{std::unordered_map} in future proposals. 

The algorithms and views in this proposal expect that vertex\_ids are integal, but it does not exclude non-integral or user-defined types in future proposals.

Hypergraphs are not supported.

\section{Impact on the Standard}
This proposal is a pure \textbf{library} extension.

\section{Interaction wtih Other Papers}
Other than the papers identified as part of the Graph Libary, there is no interaction with other proposals to the standard.

\section{Implementation Experience}
The github \href{https://github.com/stdgraph}{github.com/stdgraph} repository contains an implementation for this proposal.

\section{Usage Experience}
There is no current use of the library. There are plans to begin using it in 2024 in a commercial setting.

\section{Deployment Experience}
There is no current deployment experience of the library. There are plans for this to follow the usage experience.

\section{Performance Considerations}
The algorithms are being ported from NWGraph to the \href{https://github.com/stdgraph}{github.com/stdgraph} implementation used for this proposal. 
Performance analysis from those algorithms can be found in the peer-reviewed papers for NWGraph~\cite{REF_nwgraph_paper,gapbs_2023}.

\section{Prior Art}
\textbf{boost::graph} has been an important C++ graph implementation since 2001. It was developed with the goal of providing
a modern (at the time) generic library that addressed all the needs someone would want of a graph library. It is still a viable library used today, attesting to the value it brings.

However, boost::graph was written using C++98 in an ``expert-friendly'' style, adding many abstractions and using sophisticated tempate metaprogramming, making it difficult to use by a casual developer.

(Andrew is a co-author of boost::graph.)

\medskip

\textbf{NWGraph} (\cite{REF_nwgraph_library} and \cite{REF_nwgraph_paper}) was published in 2022
by Lumsdaine et al, bringing additional experience gained since creating boost::graph, to create a modern graph library using C++20 for its implementation 
that was more accessible to the average developer. % with the latest algorithms.

While NWGraph made important strides to introduce the idea of the graph as a range-of-ranges and implemented many important algorithms,
there are some areas it didn't address that come a practical use in the field. For instance, it didn't have a well-defined API for graph
data structures that could be applied to existing graphs, and there wasn't a uniform approach to properties.

This proposal takes the best of NWGraph, with previous work done for P1709 to define a Graph Container Interface, to provide a library that
embraces performance, ease-of-use and the ability to use the algorithms and views on externally defined graph containers.

\section{Alternatives}
We're unaware of any other library that meets the same requirements and uses concepts and ranges from C++20.

%Later revisions should include this:
%\section{Changes Library Evolution previously requested}

\section{Feature Test Macro}
The \tcode{__cpp_lib_graph} feature test macro is recommended to represent all features in this proposal including algorithms, views, concepts, traits, types, functions and graph container(s).

\section{Freestanding}
We believe this library can be used in a freestanding C++ implementation.

\section{Namespaces}
Graph containers and their views and algorithms are not interchangeable with existing containers and algorithms.
Additionally, there are some domain-specific terms that may clash with existing or future names, such as 
\tcode{degree} and \tcode{partition_id}.
For these reasons, we recommend their own namespaces. The following assumption is used in this proposal.
\begin{itemize}
\item[]\tcode{std::graph} and \tcode{std::graph::views}
\end{itemize}

\noindent
Alternative locations include the following:
\begin{itemize}
\item[]\tcode{std::ranges} and \tcode{std::ranges::views}
\item[]\tcode{std::ranges/graph} and \tcode{std::ranges::graph::views}
\end{itemize}
The advantage of these two options are that there would be no requirement to use the ranges:: prefix for things
in the std::ranges namespace, a common occurance.

\section{Notes and Considerations}
There are some interesting observations that can be made about graphs and how they compare and contrast to the 
standard library that may not be obvious.
\begin{itemize}
      \item The adjacency list, the primary data structure for this proposal, is a compound data structure of a
          range of ranges. This introduces a new form of container beyond a simple range.
      \item There is more than one possible value type, one each for edge, vertex and graph. Each is optional.
            This is in contrast to existing practice where the value type is the distinguishing difference between
            different containers, such as for \tcode{set} and \tcode{map}.
      \item Algorithms will often use views, though they can use the GCI functions when needed.
      \item Algorithms and Views often need to allocate memory internally to achieve their purpose. This is a departure from
            common practice in the standard.
      \item Storing vertices in a \tcode{map} (bi-directional range) requires a different style of programming 
            algorithms, compared to being kept in a \tcode{vector} (random access range). When using a \tcode{vector},
            \tcode{edges(g,uid)} would normally be used without much thought. Using that with a \tcode{map} would
            incur a $\mathcal{O}(\log(V))$ cost. Instead, it will use vertex id once to get the vertex reference 
            and then use \tcode{edges(g,uv)}. This is expected to result in overloading of existing algorithms based on the
            range type of container, distinguished with concepts.
\end{itemize}


\andrew{Compare code with boost::graph}
