%% \chapter{Algorithms}

\clearpage

\section{Algorithm Selection} \label{other_algo}

% \andrew{I've tagged the algorithms below as Tier 2 or Tier 3 -- denoting whether they should be done right now or done later or done much later.}

% \andrew{I've used NetworkX as inspiration for organization.  Oddly, NetworkX only has DFS as an adaptor (view).}
% \phil{If the use of Yield is any indicator, then NetworkX implements topological sort as adaptor also.}

% Algorithms require \tcode{adjacency_list<G>} unless specifically constrained to
% \tcode{mapped_adjacency_list<G>}, \tcode{index_adjacency_list<G>}, or a bidirectional variant.

When determining the algorithms to propose we split them into different tiers. Tier 1 algorithms are included
in this proposal. The algorithms selected are a result of balancing a few things:
\begin{itemdescr}
\begin{itemize}
      \item Include a rich enough set of algorithms for the library to be useful.
      \item Include algorithms with well-defined functionality and agreed-upon algorithmic description.
      \item Don't include so many that the proposal will get bogged down for years and years.
\end{itemize}
\end{itemdescr}

\subsection{Tier 1 Algorithms}
\begin{multicols}{3}
      \emph{Traversal}
      \begin{itemize}
            \item Breadth-First search
            \item Depth-First search
            \item Topological sort
      \end{itemize}
      \emph{Shortest Paths}
      \begin{itemize}
            %\item Shortest paths (driver interface)
            \item Dijkstra's algorithm
            \item Bellman-Ford algorithm
      \end{itemize}
      \emph{Clustering}
      \begin{itemize}
            \item Triangle counting
      \end{itemize}
\columnbreak
      \emph{Communities}
      \begin{itemize}
            \item Label propagation
      \end{itemize}
      \emph{Components}
      \begin{itemize}
            \item Articulation points
            \item Connected components
            \item Biconnected components
            \item Kosaraju's Strongly CC
            \item Tarjan's Strongly CC
      \end{itemize}
\columnbreak
      \emph{Maximal Independent Set}
      \begin{itemize}
            \item Maximal independent set
      \end{itemize}
      \emph{Link Analysis}
      \begin{itemize}
            \item Jaccard coefficient
      \end{itemize}
      \emph{Minimal Spanning Tree}
      \begin{itemize}
            \item Kruskal's MST
            \item Prim's MST
      \end{itemize}

\vfill\null
\end{multicols}

Traversal and Shortest Paths algorithms include single-source and multi-source versions with multiple targets.

\subsection{Other Algorithms}
Additional algorithms that were considered but not included in this proposal are shown in Table \ref{tab:other_algorithms}. 
Tier X algorithms are variations of shortest paths algorithms that complement the Single Source, Multiple Target algorithms 
in this proposal.
It is assumed that future proposals will include them, though the exact mix for each proposal will depend on feedback received
and our experience with the current proposal.

\phil{We may want to revisit the Driver idea in the future after we have more algorithms.}
%The Shortest Paths Driver is an idea of having a unified interface that chooses the best Shortest Path algorithm
%based on characteristics like non-negative edge weight, multi-threading, etc.

\begin{table}[ht]
\begin{center}
%\resizebox{\textwidth}{!}
{\begin{tabular}{l|l|l}
\hline
    \textbf{Tier 2} & \textbf{Tier 3} & \textbf{Tier X} \\
\hline
    All Pairs Shortest Paths & Jones Plassman & Single Source, Single Target: Shortest Paths Driver\\
    Floyd-Warshall & Cores: k-cores & Single Source, Single Target: BFS \\
    Johnson & Cores: k-truss & Single Source, Single Target: Dijkstra \\
    Centrality: Betweenness Centrality & Subgraph Isomorphism & Single Source, Single Target: Bellman-Ford \\
    Coloring: Greedy & & Single Source, Single Target: Delta Stepping \\
    Communities: Louvain & &  \\
    Connectivity: Minimum Cuts & & Multiple Source: Shortest Paths Driver \\
    Transitive Closure & & Multiple Source: BFS \\
    Flows: Edmunds Karp & & Multiple Source: Dijkstra \\
    Flows: Push Relabel & & Multiple Source: Bellman-Ford \\
    Flows: Boykov Kolmogorov & & Multiple Source: Delta Stepping \\
    Link Analysis: Adamic-Adar Index & &  \\
    Pathfinding: A* & & Multiple Source, Single Target: Shortest Paths Driver\\
    Best-first search & & Multiple Source, Single Target: BFS \\
    & & Multiple Source, Single Target: Dijkstra \\
    & & Multiple Source, Single Target: Bellman-Ford \\
    & & Multiple Source, Single Target: Delta Stepping \\
\hline
\end{tabular}}
\caption{Other Algorithms}
\label{tab:other_algorithms}
\end{center}
\end{table}

\andrew{All Pairs: Tier 2?  People bring this up a lot -- but it is very expensive in terms of computation and memory.}
\phil{If it's useful to enough people it should be included. Users can make their own determination of whether they want to use it, based on the cost.}

\phil{The same variations for Shortest Paths algorithms can also be useful for topological sort.}

% Floyd-Warshall $\mathcal{O}(N^2)$}
% Johnson $\mathcal{O}(N^2)$}

\clearpage
\section{Common Algorithm Definitions}

%\andrew{Need to develop this ala CppCon 2021 talk.}

%The abstraction that is used for describing and analyzing almost all graph algorithms is the adjacency list.  Naturally then implementations of graph algorithms in C++ will operate on a data structure representing an adjacency list.  And generic algorithms will be written in terms of concepts that capture the essential operations that a concrete data structure must provide in order to be used as an abstraction of an adjacency list.

%Most fundamentally (as illustrated above), an adjacency list is a collection of vertices, each of which has a collection of outgoing edges.  In terms of existing C++ concepts, we can consider an adjacency list to be a range of ranges (or, more specifically, a random access range of forward ranges).  The outer range is the collection of vertices, and the inner ranges are the collections of outgoing edges.

%\andrew{Is it better to list the concepts here or forward reference them?  Kind of a circularity.  But maybe the sequence of concepts - algorithms - concrete data types is the right one.  OTOH, std::ranges use ordering overview library-concepts-containers-algorithms.  Since a graph is a range of ranges, maybe we should follow that.}

%\phil{I think this is a good place, just before they're used in the text for this section.}
% Additional concepts used by algorithms.

Common concepts used by algorithms are in this section, extending those in the Graph Container Interface.

\subsection{Value Function and Edge Weight Concepts}

Value function concepts constrain callables that extract per-vertex or per-edge values.
Edge weights are intrinsic numeric type (\lstinline{edge_weight_function}), but could be any type 
for more general use (\lstinline{basic_edge_weight_function}).

\begin{lstlisting}
// For exposition only: base concept for any 2-arg vertex value function.
template <class VVF, class Graph, class VertexDescriptor>
concept vertex_value_function =
    invocable<VVF, const Graph&, VertexDescriptor> &&
    (!is_void_v<invoke_result_t<VVF, const Graph&, VertexDescriptor>>);

// For exposition only: base concept for any 2-arg edge value function.
template <class WF, class G, class E>
concept edge_value_function = // For exposition only
      std::invocable<WF, const G&, E> && !std::is_void_v<std::invoke_result_t<WF, const G&, E>>;

// For exposition only
template <class G, class WF, class DistanceValue, class Compare, class Combine>
concept basic_edge_weight_function = // e.g. weight(g, uv)
      edge_value_function<WF, std::remove_reference_t<G>, edge_t<G>> &&
      strict_weak_order<Compare, DistanceValue, DistanceValue> &&
      assignable_from<add_lvalue_reference_t<DistanceValue>,
            invoke_result_t<Combine, DistanceValue,
                  invoke_result_t<WF, const std::remove_reference_t<G>&, edge_t<G>>>>;

// For exposition only
template <class G, class WF, class DistanceValue>
concept edge_weight_function = // e.g. weight(g, uv)
      is_arithmetic_v<DistanceValue> &&
      is_arithmetic_v<invoke_result_t<WF, const std::remove_reference_t<G>&, edge_t<G>>> &&
      basic_edge_weight_function<G, WF, DistanceValue, less<DistanceValue>, plus<DistanceValue>>;
\end{lstlisting}

\begin{comment}
      \phil{Queueable isn't being used.}
      \begin{lstlisting}
      // queueableQ can represent std::queue and std::priority\_queue
      template <class Q>
      concept queueable = requires(Q&& q, Q::value_type value) {
      Q::value_type;
      Q::size_type;
      Q::reference;

      {q.top()};
      {q.push(value)};
      {q.pop()};
      {q.empty()};
      {q.size()};
      };
      \end{lstlisting}
\end{comment}

\subsection{Ordered Vertex Edges Concept}
The \lstinline{ordered_vertex_edges} concept requires that the outgoing edges of each vertex
are sorted in ascending order of \lstinline{target_id}. This is needed by algorithms such as
\lstinline{triangle_count} that rely on sorted adjacency lists for efficient intersection.
This is a semantic requirement that is not enforced by the graph container interface, but it 
is documented as a requirement for algorithms that need it.

\begin{lstlisting}
// For exposition only
template <class G>
concept ordered_vertex_edges = adjacency_list<G> &&
  requires(G& g, vertex_t<G> u) {
    requires forward_iterator<decltype(ranges::begin(out_edges(g, u)))>;
  };
\end{lstlisting}

\subsection{Vertex Property Function Concepts}
Most algorithms maintain per-vertex state (distances, predecessors, component labels, etc.)
through callable \emph{function objects} of the form
\tcode{(const G\&, vertex\_id\_t<G>) -> value\&}.
This function-first design is more flexible than direct container access: property values may
reside in vertex properties, in external containers, or in any custom storage.

\subsubsection{\tcode{vertex_property_fn_for} and \tcode{vertex_fn_value_t}}
\tcode{vertex_property_fn_for<VF,G>} is the general concept for per-vertex property functions,
requiring a callable that returns a mutable lvalue reference to the stored value.
It is used directly by \tcode{connected\_components}, \tcode{kosaraju}, \tcode{tarjan\_scc},
and \tcode{label\_propagation}.
\tcode{vertex_fn_value_t<VF,G>} extracts the corresponding value type.

\begin{lstlisting}
// For exposition only
template <class VF, class G>
concept vertex_property_fn_for =
      invocable<VF&, const remove_reference_t<G>&, const vertex_id_t<G>&> &&
      is_lvalue_reference_v<invoke_result_t<VF&, const remove_reference_t<G>&, const vertex_id_t<G>&>>;

// For exposition only
template <class VF, class G>
using vertex_fn_value_t = remove_cvref_t<
    invoke_result_t<VF&, const remove_reference_t<G>&, const vertex_id_t<G>&>>;
\end{lstlisting}

\subsubsection{\tcode{distance_fn_for} and \tcode{predecessor_fn_for}}
\tcode{distance_fn_for} and \tcode{predecessor_fn_for} are concept aliases for
\tcode{vertex_property_fn_for}, giving semantic names to the distance and predecessor
function parameters of shortest-path and MST algorithms.
\tcode{distance_fn_value_t} and \tcode{predecessor_fn_value_t} are corresponding
type aliases for \tcode{vertex_fn_value_t}.

\begin{lstlisting}
// For exposition only
template <class DF, class G>
concept distance_fn_for = vertex_property_fn_for<DF, G>;

// For exposition only
template <class DF, class G>
using distance_fn_value_t = vertex_fn_value_t<DF, G>;

// For exposition only
template <class PF, class G>
concept predecessor_fn_for = vertex_property_fn_for<PF, G>;

// For exposition only
template <class PF, class G>
using predecessor_fn_value_t = vertex_fn_value_t<PF, G>;
\end{lstlisting}

\subsubsection{\tcode{container_value_fn} Adaptor}
\tcode{container_value_fn<Container>} adapts any subscriptable container into a property
function, making it straightforward to reuse a \tcode{std::vector} or
\tcode{std::unordered\_map}:
\begin{lstlisting}[texcl=false]
std::vector<double> dist_vec(num_vertices(g), infinite_distance<double>());
auto dist_fn = container_value_fn{dist_vec}; // dist_fn(g, uid) -> dist_vec[uid] (by ref)
dijkstra_shortest_paths(g, source, dist_fn, _null_predecessor, weight_fn);
\end{lstlisting}

\subsubsection{Legacy: \tcode{vertex_property_map_for} Concept}
The container-subscript concept \tcode{vertex_property_map_for<M,G>} is retained for
the \tcode{init\_shortest\_paths} helper functions, which take
subscriptable containers directly.  It is satisfied by \tcode{std::vector<T>} for index graphs
and \tcode{std::unordered\_map<vertex\_id\_t<G>,T>} for mapped graphs.

\subsection{Vertex Property Map}
Algorithms frequently need per-vertex storage for quantities such as distances, predecessor
IDs, component labels, and visited flags.  Because the graph library supports both
\emph{index-based} vertex containers (\tcode{std::vector}, \tcode{std::deque}) and
\emph{key-based} vertex containers (\tcode{std::map}, \tcode{std::unordered\_map}), the
appropriate external container differs:
\begin{itemize}
      \item For index graphs, a \tcode{std::vector<T>} addressed by \tcode{vertex\_id\_t<G>} provides $\mathcal{O}(1)$ access.
      \item For mapped graphs, a \tcode{std::unordered\_map<vertex\_id\_t<G>, T>} is required
            because vertex IDs are non-contiguous keys.
\end{itemize}

The \tcode{vertex\_property\_map} facility provides a uniform abstraction that selects the
correct container automatically and offers factory functions and access helpers so that
algorithms and users need not distinguish between the two cases.

\subsubsection{Synopsis}
{\small
      \lstinputlisting{D3128_Algorithms/src/vertex_property_map.hpp}
}

\subsubsection{\tcode{vertex\_property\_map<G, T>} Type Alias}
A conditional type alias that resolves to \tcode{std::vector<T>} when the graph \tcode{G}
satisfies \tcode{index\_vertex\_range<G>}, and to
\tcode{std::unordered\_map<vertex\_id\_t<G>, T>} otherwise.

\subsubsection{\tcode{make\_vertex\_property\_map} Factory Functions}
Two overloads create a \tcode{vertex\_property\_map}:
\begin{itemize}
      \item \textbf{Eager initialization} ---
            \tcode{make\_vertex\_property\_map(g, init\_value)} pre-populates every vertex with
            \tcode{init\_value}.  Useful when an algorithm reads all entries before writing
            (e.g.\ component labels).  $\mathcal{O}(|V|)$.
      \item \textbf{Lazy initialization} ---
            \tcode{make\_vertex\_property\_map<G, T>(g)} creates a default-constructed
            (index graphs) or empty-but-reserved (mapped graphs) container.
            Useful with \tcode{vertex\_property\_map\_get} when absence has a semantic meaning
            (e.g.\ infinity for distances, \tcode{false} for visited flags).
\end{itemize}

\subsubsection{Access Helpers}
\begin{itemize}
      \item \tcode{vertex\_property\_map\_contains(m, uid)} ---
            Tests whether vertex~\tcode{uid} has an entry.
            For \tcode{vector}: \tcode{uid < size(m)}.
            For \tcode{unordered\_map}: \tcode{m.contains(uid)}.
      \item \tcode{vertex\_property\_map\_get(m, uid, default\_val)} ---
            Returns the stored value if present, otherwise returns \tcode{default\_val}
            \emph{without insertion}.  This avoids accidentally growing an
            \tcode{unordered\_map} through \tcode{operator[]}.
\end{itemize}

\subsubsection{\tcode{vertex\_property\_map\_value\_t<Container>}}
A trait that extracts the per-vertex value type from a \tcode{vertex\_property\_map} container:
\tcode{T} for \tcode{vector<T>}, \tcode{V} for \tcode{unordered\_map<K,V>}.

\subsubsection{\tcode{vertex\_property\_map\_for<M, G>} Concept}
Requires that \tcode{m[uid]} is valid for \tcode{uid} of type \tcode{vertex\_id\_t<G>} and
the result is convertible to \tcode{vertex\_property\_map\_value\_t<M>}.
This concept constrains algorithm parameters that accept subscriptable containers directly,
such as the \tcode{init\_shortest\_paths} helpers.

\subsection{Visitor Concepts and Classes}
Visitors are optional member functions on a user-defined class that are called during the execution of an algorithm. 
Each algorithm has its own set of visitor events that it supports, and each event function must match the visitor concepts 
shown in this section.

The visitor events mimic those used in the Boost Graph Library.

\subsubsection{Vertex Visitor Concepts}
{\small
      \lstinputlisting{D3128_Algorithms/src/visitor_vertex.hpp}
}
Each vertex event concept exists in two variants.
The \textit{descriptor} variant (e.g.\ \lstinline{has_on_initialize_vertex}) requires an event
function whose second argument is \lstinline{const vertex_t<G>&}.
The \textit{ID} variant (e.g.\ \lstinline{has_on_initialize_vertex_id}) requires an event
function whose second argument is \lstinline{const vertex_id_t<G>&}.
The ID variant is useful when the visitor only needs the vertex identifier rather than the full
vertex descriptor.  In both variants the first argument is \lstinline{const G&}.

The vertex events are called under the following conditions.
\begin{itemize}
      \item \lstinline{on_initialize_vertex(g, vdesc)} (or \lstinline{on_initialize_vertex(g, uid)})
            is called once for each vertex before the algorithm is run.
      \item \lstinline{on_discover_vertex(g, vdesc)} (or \lstinline{on_discover_vertex(g, uid)})
            is called once for each source vertex passed to the algorithm.
      \item \lstinline{on_examine_vertex(g, vdesc)} (or \lstinline{on_examine_vertex(g, uid)})
            is called for a vertex before any of its outgoing edges are examined.  It is possible
            that it will be called multiple times for the same vertex if paths are found to it
            from other vertices with a shorter distance.
      \item \lstinline{on_start_vertex(g, vdesc)} (or \lstinline{on_start_vertex(g, uid)})
            is called for a vertex that is being examined.
      \item \lstinline{on_finish_vertex(g, vdesc)} (or \lstinline{on_finish_vertex(g, uid)})
            is called for a vertex that is being examined, after all its outgoing edges have been
            examined.
\end{itemize}

\subsubsection{Edge Visitor Concepts}
{\small
      \lstinputlisting{D3128_Algorithms/src/visitor_edge.hpp}
}
The edge events are called under the following conditions.
\begin{itemize}
      \item \lstinline{on_examine_edge(edesc)} is called for edge of the source vertex that is
            being examined.
      \item \lstinline{on_edge_relaxed(edesc)} is called when the distance to the target vertex of the edge
            is relaxed, or decreased.
      \item \lstinline{on_edge_not_relaxed(edesc)} is called when the distance to the target vertex of the edge
            is not relaxed, or not decreased.
      \item \lstinline{on_edge_minimized(edesc)} is called when the distance to the target vertex of the edge
            is minimized, or decreased to the minimum value.
      \item \lstinline{on_edge_not_minimized(edesc)} is called when the distance to the target vertex of the edge
            is not minimized, or not decreased to the minimum value.
      \item \lstinline{on_tree_edge(edesc)} is called when the edge is added to the tree.
      \item \lstinline{on_back_edge(edesc)} is called when the edge is a back edge.
      \item \lstinline{on_forward_or_cross_edge(edesc)} is called when the edge is a forward or cross edge.
      \item \lstinline{on_finish_edge(edesc)} is called when the edge is finished being examined.
\end{itemize}


\subsubsection{Visitor Classes}
\tcode{empty_visitor} is used when no visitor is needed. It is a no-op struct that does nothing.
{\small
      \lstinputlisting{D3128_Algorithms/src/visitor_other.hpp}
}



\section{Traversal}

\subsection{Breadth-First Search}
Compute the breadth-first path from one or more \tcode{source} vertices to all reachable vertices in \tcode{graph}.

\begin{table}[ht]
\setcellgapes{3pt}
\makegapedcells
\centering
\begin{tabular}{|P{0.30\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|}
\hline
      \multirowcell{2}{
            \textbf{Complexity} \\
            $\mathcal{O}(|V| + |E|)$
            }
      & \textbf{Directed?} Yes & \textbf{Cycles?} No & \textbf{Throws?} Yes \\
      & \textbf{Multi-edge?} No & \textbf{Self-loops} Yes & \\
\hline
\end{tabular}
%\caption{Algorithm Example}
\end{table}

%\phil{BGL doesn't include distances or predecessors.}

\subsubsection{Single Source Breadth-First Search}

{\small
      \lstinputlisting{D3128_Algorithms/src/breadth_first_search.hpp}
}

\subsubsection{Multi-Source Breadth-First Search}

{\small
      \lstinputlisting{D3128_Algorithms/src/breadth_first_search_multi.hpp}
}

\begin{itemdescr}
      \pnum\mandates
            \begin{itemize}
                  \item \tcode{G} satisfies \tcode{adjacency_list<G>}.
                  \item For the multi-source version, \tcode{Sources} satisfies \tcode{input_range}
                        with \tcode{range_value_t<Sources>} convertible to \tcode{vertex_id_t<G>}.
            \end{itemize}
      \pnum\preconditions
            \begin{itemize}
                  \item
                        \lstinline{0 <= source < num_vertices(graph)}, for each \lstinline{source} in \lstinline{sources},
                                   for the multi-source version.
            \end{itemize}
      \pnum\hardprecond
            \begin{itemize}
                  \item
                        \lstinline{0 <= source < num_vertices(graph)} for the single-source version.
            \end{itemize}
      \pnum\effects
            \begin{itemize}
                  \item Does not modify the graph \lstinline{g}.
                  \item Member functions on the \lstinline{visitor} parameter are called during the algorithm's execution.
                        The functions are optional and, when included, must follow the visitor concepts for the events.
                        No overhead is incurred if the functions are not included.
                        The events supported are \lstinline{on_initialize_vertex}, \lstinline{on_discover_vertex},
                        \lstinline{on_examine_vertex}, \lstinline{on_finish_vertex}, and \lstinline{on_examine_edge}.
            \end{itemize}
      \pnum\postconditions
            \begin{itemize}
                  \item All vertices reachable from any source are visited exactly once.
                  \item Visitor callbacks are invoked in BFS (level-order) traversal order.
                  \item The graph \lstinline{g} is unchanged.
            \end{itemize}
      %\pnum\returns \lstinline{void} \\
      \pnum\throws
            \begin{itemize}
                  \item \lstinline{std::bad_alloc} if the visited array or queue cannot be allocated.
                  \item Any exception thrown by \lstinline{visitor} callbacks is propagated.
                  \item \textbf{Exception guarantee:} Basic --- \lstinline{g} remains unchanged;
                        partial traversal may have occurred.
            \end{itemize}
      \pnum\complexity
            \begin{itemize}
                  \item \textbf{Time:} $\mathcal{O}(|V| + |E|)$ --- BFS uses a FIFO queue; each vertex and edge is visited at most once.
                  \item \textbf{Space:} $\mathcal{O}(|V|)$ auxiliary --- visited array and queue.
            \end{itemize}
      \pnum\remarks
            \begin{itemize}
                  \item \tcode{dijkstra_shortest_paths} provides extended functionality if \tcode{breadth_first_search} doesn't have
                         enough capability.
                  \item The optional \lstinline{alloc} parameter supplies an allocator used for
                        the internal FIFO queue.  Defaults to \tcode{std::allocator<std::byte>}.
                        Provide a custom allocator (e.g.~a pool allocator) to control the memory
                        used by the traversal queue.
            \end{itemize}
      %\pnum\errors
\end{itemdescr}





\subsection{Depth-First Search}
Traverse the depth-first path from one or more \tcode{source} vertices to all reachable vertices in \tcode{graph}.

\begin{table}[ht]
\setcellgapes{3pt}
\makegapedcells
\centering
\begin{tabular}{|P{0.30\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|}
\hline
      \multirowcell{2}{
            \textbf{Complexity} \\
            $\mathcal{O}(|V| + |E|)$
            }
      & \textbf{Directed?} Yes & \textbf{Cycles?} No & \textbf{Throws?} No \\
      & \textbf{Multi-edge?} Yes & \textbf{Self-loops} Yes & \\
\hline
\end{tabular}
%\caption{Algorithm Example}
\end{table}

\subsubsection{Single Source Depth-First Search}

{\small
      \lstinputlisting{D3128_Algorithms/src/depth_first_search.hpp}
}

% \subsubsection{Multi-Source Depth-First Search}

% {\small
%       \lstinputlisting{D3128_Algorithms/src/depth_first_search_multi.hpp}
% }

\begin{itemdescr}
      \pnum\mandates
            \begin{itemize}
                  \item \tcode{G} satisfies \tcode{adjacency_list<G>}.
            \end{itemize}
      \pnum\preconditions
            \begin{itemize}
                  \item \lstinline{source} is a valid vertex ID in \lstinline{g}.
                  \item The graph \lstinline{g} is not modified during traversal.
            \end{itemize}
      \pnum\hardprecond
            \begin{itemize}
                  \item
                        \lstinline{0 <= source < num_vertices(graph)}.
            \end{itemize}
      \pnum\effects
            \begin{itemize}
                  \item Does not modify the graph \lstinline{g}.
                  \item Classifies every edge reachable from \lstinline{source} as tree, back, or forward/cross.
                  \item Member functions on the \lstinline{visitor} parameter are called during the algorithm's execution.
                        The functions are optional and, when included, must follow the visitor concepts for the events.
                        No overhead is incurred if the functions are not included.
                        The events supported are \lstinline{on_initialize_vertex}, \lstinline{on_start_vertex},
                        \lstinline{on_discover_vertex}, \lstinline{on_finish_vertex},
                        \lstinline{on_examine_edge}, \lstinline{on_tree_edge}, \lstinline{on_back_edge},
                        \lstinline{on_forward_or_cross_edge}, and \lstinline{on_finish_edge}.
            \end{itemize}
      \pnum\postconditions
            \begin{itemize}
                  \item All vertices reachable from \lstinline{source} are visited exactly once.
                  \item \lstinline{on_finish_vertex} is called in reverse topological order for DAGs.
                  \item The graph \lstinline{g} is unchanged.
            \end{itemize}
      %\pnum\returns \lstinline{void} \\
      \pnum\throws
            \begin{itemize}
                  \item \lstinline{std::bad_alloc} if the color array or stack cannot be allocated.
                  \item Any exception thrown by \lstinline{visitor} callbacks is propagated.
                  \item \textbf{Exception guarantee:} Basic --- \lstinline{g} remains unchanged;
                        partial traversal may have occurred.
            \end{itemize}
      \pnum\complexity
            \begin{itemize}
                  \item \textbf{Time:} $\mathcal{O}(|V| + |E|)$ --- each vertex and edge is visited at most once.
                  \item \textbf{Space:} $\mathcal{O}(|V|)$ auxiliary --- color array and DFS stack.
            \end{itemize}
      \pnum\remarks
            \begin{itemize}
                  \item Uses iterative DFS with explicit stack to avoid recursion-depth limits.
                  \item Edge iterators are stored on the stack for $\mathcal{O}(1)$ resume after backtracking.
                  \item The optional \lstinline{alloc} parameter supplies an allocator used for
                        the internal DFS stack.  Defaults to \tcode{std::allocator<std::byte>}.
                        Provide a custom allocator (e.g.~a pool allocator) to control the memory
                        used by the traversal stack.
            \end{itemize}
      %\pnum\errors
\end{itemdescr}

\subsection{Topological Sort}
A linear ordering of vertices such that for every directed edge (u,v) from vertex u to vertex v, u comes before v in the ordering.

\begin{table}[ht]
\setcellgapes{3pt}
\makegapedcells
\centering
\begin{tabular}{|P{0.30\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|}
\hline
      \multirowcell{2}{
            \textbf{Complexity} \\
                  $\mathcal{O}(|E|+|V|)$
            }
      & \textbf{Directed?} Yes & \textbf{Cycles?} No & \textbf{Throws?} No \\
      & \textbf{Multi-edge?} No & \textbf{Self-loops} Yes & \\
\hline
\end{tabular}
%\caption{Algorithm Example}
\end{table}

\subsubsection{Full-Graph Topological Sort}

{\small
      \lstinputlisting{D3128_Algorithms/src/topological_sort_helpers.hpp}
}

\subsubsection{Single Source Topological Sort}

{\small
      \lstinputlisting{D3128_Algorithms/src/topological_sort.hpp}
}

\subsubsection{Multi-Source Topological Sort}

{\small
      \lstinputlisting{D3128_Algorithms/src/topological_sort_multi.hpp}
}

\begin{itemdescr}
      \pnum\mandates
            \begin{itemize}
                  \item \tcode{G} satisfies \tcode{adjacency_list<G>}.
                  \item \tcode{OutputIterator} satisfies \tcode{output_iterator<OutputIterator, vertex_id_t<G>>}.
            \end{itemize}
      \pnum\preconditions
            \begin{itemize}
                  \item
                        \lstinline{0 <= source < num_vertices(graph)} for the single-source version.
                  \item
                        \lstinline{0 <= source < num_vertices(graph)}, for each \lstinline{source} in \lstinline{sources},
                        for the multi-source version.
            \end{itemize}
      \pnum\effects
            \begin{itemize}
                  \item
                        Vertices reachable from the source(s) are written to the output iterator in reverse
                        topological (finish-time) order.
                  \item
                        The full-graph overload visits all vertices regardless of reachability from a single source.
            \end{itemize}
      \pnum\returns
            \begin{itemize}
                  \item \lstinline{true} if the topological ordering succeeded.
                  \item \lstinline{false} if a cycle was detected in the (sub)graph reachable from the source(s).
                  \item Attribute: \lstinline{[[nodiscard]]}.
            \end{itemize}
      \pnum\complexity
            \begin{itemize}
                  \item \textbf{Time:} $\mathcal{O}(|V| + |E|)$ --- DFS-based; each vertex and edge is visited at most once.
                  \item \textbf{Space:} $\mathcal{O}(|V|)$ auxiliary --- color array, finish-order vector, and DFS stack.
            \end{itemize}
      \pnum\remarks
            \begin{itemize}
                  \item Duplicate sources do not affect the algorithm's complexity or correctness.
                  \item The full-graph overload visits all vertices regardless of reachability from a single source.
                  \item The optional \lstinline{alloc} parameter supplies an allocator used for
                        the internal finish-order vector and DFS stack.  Defaults to
                        \tcode{std::allocator<std::byte>}.  All three overloads (single-source,
                        multi-source, full-graph) accept this parameter.
            \end{itemize}
      %\pnum\errors
\end{itemdescr}



\section{Shortest Paths}

\begin{comment}
      \subsection{Driver Interface}

      \andrew{I am not sure we should have the unified interface.  We need to be more parsimonious in our interfaces.  Users can read the documentation for which algorithms to use.  And, if they are using graph algorithms, we should assume a certain level of knowledge about graph algorithms.  OTOH, it is only a handful of algorithms.}

      \andrew{I am also not sure we should have ``shortest distance'' variants.  That doubles the number of functions in the interface.
            For each function we have shortest paths, s-t paths, multi-source paths, parallel = 6X variants for each base function.  If we add shortest distances, that will make 12X.  OTOH, we could consider not having s-t paths or not having multi-source paths -- which would leave 4X for each base function.  However, I think people will want s-t and multi-source.
      }
      \phil{\tcode{dijkstra_shortest_distances} includes predecessor and distances, so excluding \tcode{dijkstra_shortest_distances} won't impact 
      the user much.}


      {\small
            \lstinputlisting{D3128_Algorithms/src/shortest_paths.hpp}
      }

      \andrew{The variety of algorithms was inspired by networkx....  Which also had ``distance'' variants.}

      \phil{I assume \tcode{adjacency_list_graph} is the same as our \tcode{adjacency_list}. \tcode{bidirectional_adjacency_list_graph} is new; what to do with it?}
\end{comment}

\subsection{Initialization}

{\small
      \lstinputlisting{D3128_Algorithms/src/shortest_paths_helpers.hpp}
}
\begin{itemdescr}
      \pnum\effects
            \begin{itemize}
                  \item
                        \lstinline{init_shortest_paths(g, distances)} sets \lstinline{distances[v] = infinite_distance()}
                        for each vertex \lstinline{v} in \lstinline{g}.
                  \item
                        \lstinline{init_shortest_paths(g, distances, predecessors)} does the same and additionally
                        sets \lstinline{predecessors[v] = v} for each vertex \lstinline{v} in \lstinline{g}.
            \end{itemize}
      \pnum\returns 
            \begin{itemize}
                  \item \lstinline{infinite_distance<T>()} returns the largest value of type \lstinline{T},
                        typically \lstinline{numeric_limits<T>::max()} for numeric types.
                  \item \lstinline{zero_distance<T>()} returns a zero-valued instance of type \lstinline{T},
                        typically \lstinline{T()} for numeric types.
            \end{itemize}
\end{itemdescr}

\subsection{Dijkstra Shortest Paths and Shortest Distances}

Compute the shortest path and associated distance from vertex \tcode{source} to all reachable vertices in \tcode{graph}
using non-negative weights.

\begin{table}[ht]
\setcellgapes{3pt}
\makegapedcells
\centering
\begin{tabular}{|P{0.30\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|}
\hline
      \multirowcell{2}{
            \textbf{Complexity} \\
            $\mathcal{O}((|E| + |V|)\log{|V|})$
            }
      & \textbf{Directed?} Yes & \textbf{Cycles?} No & \textbf{Throws?} Yes \\
      & \textbf{Multi-edge?} No & \textbf{Self-loops} Yes & \\
\hline
\end{tabular}
%\caption{Algorithm Example}
\end{table}

Note that complexity may be $\mathcal{O}(|E| + |V|\log{|V|)}$ for certain implementations that use a Fibonacci heap
instead of a binary heap implemented with \tcode{std::priority_queue}.

\subsubsection{Dijkstra Shortest Paths}

\phil{Feedback: use std::invocable instead of std::function to avoid performance penalty. 
      (Is this really an issue, when the function parameter has a default lambda?)}

\paragraph{Single-Source Shortest Paths}
{\small
      \lstinputlisting{D3128_Algorithms/src/dijkstra_shortest_paths.hpp}
}

\paragraph{Multi-Source Shortest Paths}
{\small
      \lstinputlisting{D3128_Algorithms/src/dijkstra_shortest_paths_multi.hpp}
}

%\andrew{Should we include the check for negative weights (performance)?}
%\phil{It's the cost of safety. BGL did it, so there is precedent. The check isn't done if weights are unsigned 
%integral values.}
%\phil{For the standard, no, because it is an O(n) check.  But we can add it as an option in the future for the
%      non-standard version.}

\begin{itemdescr}
      \pnum\mandates
            \begin{itemize}
                  \item \tcode{G} satisfies \tcode{adjacency_list<G>}.
                  \item \tcode{DistanceFn} satisfies \tcode{distance_fn_for<DistanceFn, G>}
                        with \tcode{is_arithmetic_v<distance_fn_value_t<DistanceFn, G>>}.
                  \item \tcode{PredecessorFn} satisfies \tcode{predecessor_fn_for<PredecessorFn, G>}.
                  \item \tcode{WF} satisfies \tcode{basic_edge_weight_function<G, WF, distance_fn_value_t<DistanceFn, G>, Compare, Combine>}.
            \end{itemize}
      \pnum\constantwhen
            \begin{itemize}
                  \item All operations on the distance, predecessor, and weight types
                        are constant subexpressions, and no visitor callback throws.
            \end{itemize}
      \pnum\preconditions
            \begin{itemize}
                  \item
                        Each \lstinline{source} in \lstinline{sources} is a valid vertex ID in \lstinline{g},
                        for the multi-source version.
                  \item
                        \lstinline{distance(g, vid) = infinite_distance()} for every vertex \lstinline{vid} in \lstinline{g}.
                  \item
                        \lstinline{predecessor(g, vid) = vid} for every vertex \lstinline{vid} in \lstinline{g}.
                  \item
                        The weight function \lstinline{weight} must return a non-negative value.
            \end{itemize}
      \pnum\hardprecond
            \begin{itemize}
                  \item \lstinline{source} is a valid vertex ID in \lstinline{g} for the single-source version.
            \end{itemize}
      \pnum\effects
            \begin{itemize}
                  \item
                        If vertex with index \lstinline{i} is reachable from vertex \lstinline{source}, then
                        \lstinline{distance(g, i)} will contain the distance from \lstinline{source} to vertex
                        \lstinline{i}.  Otherwise \lstinline{distance(g, i)} will contain
                        \lstinline{infinite_distance()}.
                  \item
                        If vertex with index \lstinline{i} is reachable
                        from vertex \lstinline{source}, then \lstinline{predecessor(g, i)} will contain the
                        predecessor vertex of vertex \lstinline{i}. Otherwise \lstinline{predecessor(g, i)} will contain
                        \lstinline{i}.
                  \item Does not modify the graph \lstinline{g}.
                  \item Member functions on the \lstinline{visitor} parameter are called during the algorithm's execution.
                        The functions are optional and, when included, must follow the visitor concepts for the events.
                        No overhead is incurred if the functions are not included.
                        The events supported are \lstinline{on_initialize_vertex}, \lstinline{on_discover_vertex},
                        \lstinline{on_examine_vertex}, \lstinline{on_finish_vertex}, \lstinline{on_examine_edge},
                        \lstinline{on_edge_relaxed}, and \lstinline{on_edge_not_relaxed}.
            \end{itemize}
      \pnum\postconditions
            \begin{itemize}
                  \item \lstinline{distance(g, s) == 0} for all sources \lstinline{s}.
                  \item For every reachable vertex \lstinline{v}, \lstinline{distance(g, v)} contains the
                        shortest distance from the nearest source.
                  \item For every reachable vertex \lstinline{v}, \lstinline{predecessor(g, v)} contains the
                        predecessor on a shortest path tree.
                  \item For every unreachable vertex \lstinline{v},
                        \lstinline{distance(g, v) == infinite_distance<distance_fn_value_t<DistanceFn,G>>()}.
            \end{itemize}
      %\pnum\returns \lstinline{void} \\
      \pnum\throws
            \begin{itemize}
                  \item \lstinline{std::out_of_range} if any source vertex ID is not in \lstinline{vertices(g)}.
                  \item \lstinline{std::out_of_range} if a negative edge weight is encountered
                        (for signed weight types).
                  \item Any exception thrown by \lstinline{visitor} callbacks is propagated.
                  \item \textbf{Exception guarantee:} Basic --- \lstinline{g} remains unchanged;
                        \lstinline{distances} and \lstinline{predecessor} may be partially modified.
            \end{itemize}
      \pnum\complexity
            \begin{itemize}
                  \item \textbf{Time:} $\mathcal{O}((|E| + |V|)\log{|V|})$ based on using the binary heap in \tcode{std::priority_queue}.
                  \item \textbf{Space:} $\mathcal{O}(|V|)$ auxiliary --- priority queue and bookkeeping arrays.
                  \item An implementation may choose to use a Fibonacci heap for a time complexity of $\mathcal{O}(|E| + |V|\log{|V|})$.
            \end{itemize}
      \pnum\remarks 
            \begin{itemize}
                  \item Duplicate sources do not affect the algorithm's complexity or correctness.
                  \item Bellman-Ford Shortest Paths allows negative weights with the consequence of greater complexity.
                  \item \tcode{DistanceFn} and \tcode{PredecessorFn} must satisfy \tcode{distance_fn_for}
                        and \tcode{predecessor_fn_for} respectively.  For index graphs, wrap a \tcode{std::vector}
                        using \tcode{container_value_fn}; for mapped graphs, wrap a \tcode{std::unordered_map}
                        the same way (see §\textit{Vertex Property Function Concepts}).
                  \item Pass \lstinline{_null_predecessor} as \lstinline{predecessor} when predecessor tracking
                        is not needed (avoids storing results).
                  \item The optional \lstinline{alloc} parameter supplies an allocator used for
                        the internal priority queue.  Defaults to \tcode{std::allocator<std::byte>}.
                        Provide a custom allocator (e.g.~a pool allocator) to control the memory
                        used by the priority queue during traversal.
            \end{itemize}
      %\pnum\errors
\end{itemdescr}

\paragraph{Example: Dijkstra on a Mapped Graph}
The following example demonstrates \tcode{dijkstra_shortest_paths} on a graph whose
vertex IDs are \tcode{std::string} values.  Lambda closures over \tcode{std::unordered_map}
instances serve as the \tcode{DistanceFn} and \tcode{PredecessorFn} arguments.  All other
algorithm code is identical to the index-graph case.
{\small
      \lstinputlisting{D3128_Algorithms/src/dijkstra_example_mapped.hpp}
}

\subsubsection{Dijkstra Shortest Distances}
This is the same as \textit{Shortest Paths} except that it excludes the predecessors,
giving a small performance improvement with lower memory overhead.

\paragraph{Single-Source Shortest Distances}
{\small
      \lstinputlisting{D3128_Algorithms/src/dijkstra_shortest_dists.hpp}
}

\paragraph{Multi-Source Shortest Distances}
{\small
      \lstinputlisting{D3128_Algorithms/src/dijkstra_shortest_dists_multi.hpp}
}

\begin{itemdescr}
      \pnum\mandates
            \begin{itemize}
                  \item \tcode{G} satisfies \tcode{adjacency_list<G>}.
                  \item \tcode{DistanceFn} satisfies \tcode{distance_fn_for<DistanceFn, G>}
                        with \tcode{is_arithmetic_v<distance_fn_value_t<DistanceFn, G>>}.
                  \item \tcode{WF} satisfies \tcode{basic_edge_weight_function<G, WF, distance_fn_value_t<DistanceFn, G>, Compare, Combine>}.
            \end{itemize}
      \pnum\constantwhen
            \begin{itemize}
                  \item All operations on the distance and weight types
                        are constant subexpressions, and no visitor callback throws.
            \end{itemize}
      \pnum\preconditions
            \begin{itemize}
                  \item
                        Each \lstinline{source} in \lstinline{sources} is a valid vertex ID in \lstinline{g},
                        for the multi-source version.
                  \item
                        \lstinline{distance(g, vid) = infinite_distance()} for every vertex \lstinline{vid} in \lstinline{g}.
                  \item
                        The weight function \lstinline{weight} must return a non-negative value.
            \end{itemize}
      \pnum\hardprecond
            \begin{itemize}
                  \item \lstinline{source} is a valid vertex ID in \lstinline{g} for the single-source version.
            \end{itemize}
      \pnum\effects
            \begin{itemize}
                  \item
                        If vertex with index \lstinline{i} is reachable from vertex \lstinline{source}, then
                        \lstinline{distance(g, i)} will contain the distance from \lstinline{source} to vertex
                        \lstinline{i}.  Otherwise \lstinline{distance(g, i)} will contain
                        \lstinline{infinite_distance()}.
                  \item Does not modify the graph \lstinline{g}.
                  \item Member functions on the \lstinline{visitor} parameter are called during the algorithm's execution.
                        The functions are optional and, when included, must follow the visitor concepts for the events.
                        No overhead is incurred if the functions are not included.
                        The events supported are \lstinline{on_initialize_vertex}, \lstinline{on_discover_vertex},
                        \lstinline{on_examine_vertex}, \lstinline{on_finish_vertex}, \lstinline{on_examine_edge},
                        \lstinline{on_edge_relaxed}, and \lstinline{on_edge_not_relaxed}.
            \end{itemize}
      \pnum\postconditions
            \begin{itemize}
                  \item \lstinline{distance(g, s) == 0} for all sources \lstinline{s}.
                  \item For every reachable vertex \lstinline{v}, \lstinline{distance(g, v)} contains the
                        shortest distance from the nearest source.
                  \item For every unreachable vertex \lstinline{v},
                        \lstinline{distance(g, v) == infinite_distance<distance_fn_value_t<DistanceFn,G>>()}.
            \end{itemize}
      %\pnum\returns \lstinline{void} \\
      \pnum\throws
            \begin{itemize}
                  \item \lstinline{std::out_of_range} if any source vertex ID is not in \lstinline{vertices(g)}.
                  \item \lstinline{std::out_of_range} if a negative edge weight is encountered
                        (for signed weight types).
                  \item Any exception thrown by \lstinline{visitor} callbacks is propagated.
                  \item \textbf{Exception guarantee:} Basic --- \lstinline{g} remains unchanged;
                        \lstinline{distances} may be partially modified.
            \end{itemize}
      %\pnum\result
      %\pnum\returns \lstinline{void} \\
      \pnum\complexity
            \begin{itemize}
                  \item \textbf{Time:} $\mathcal{O}((|E| + |V|)\log{|V|})$ based on using the binary heap in \tcode{std::priority_queue}.
                  \item \textbf{Space:} $\mathcal{O}(|V|)$ auxiliary --- priority queue and bookkeeping arrays.
                  \item An implementation may choose to use a Fibonacci heap for a time complexity of $\mathcal{O}(|E| + |V|\log{|V|})$.
            \end{itemize}
      \pnum\remarks 
            \begin{itemize}
                  \item Duplicate sources do not affect the algorithm's complexity or correctness.
                  \item Bellman-Ford Shortest Distances allows negative weights with the consequence of greater complexity.
                  \item \tcode{DistanceFn} must satisfy \tcode{distance_fn_for<DistanceFn,G>}.  Wrap a
                        \tcode{std::vector} or \tcode{std::unordered_map} with \tcode{container_value_fn}
                        (see §\textit{Vertex Property Function Concepts}).
                  \item The optional \lstinline{alloc} parameter supplies an allocator used for
                        the internal priority queue.  Defaults to \tcode{std::allocator<std::byte>}.
                        Provide a custom allocator (e.g.~a pool allocator) to control the memory
                        used by the priority queue during traversal.
            \end{itemize}
      %\pnum\errors
\end{itemdescr}

\subsection{Bellman-Ford Shortest Paths and Shortest Distances}
Compute the shortest path and associated distance from vertex \tcode{source} to all reachable vertices in \tcode{graph}.

\begin{table}[ht]
\setcellgapes{3pt}
\makegapedcells
\centering
\begin{tabular}{|P{0.30\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|}
\hline
      \multirowcell{2}{
            \textbf{Complexity} \\
                  $\mathcal{O}(|E| \cdot |V|)$ \\
            }
      & \textbf{Directed?} Yes & \textbf{Cycles?} No & \textbf{Throws?} Yes \\
      & \textbf{Multi-edge?} No & \textbf{Self-loops} Yes & \\
\hline
\end{tabular}
%\caption{Algorithm Example}
\end{table}

%\andrew{Complexity is really is atrocious.  Suitable only for really small graphs.}

The Bellman-Ford algorithm supports the use of negative edge weights, at cost in performance. 
Because of its complexity, it can only be used for small graphs.
If a user can guarantee that a graph has positive edge weights then Dijkstra's algorithm 
provides far better performance.

There is a special case where edges form a negative cycle. 
"If a graph contains a 'negative cycle' (i.e. a cycle whose edges sum to a negative value) 
that is reachable from the source, then there is no cheapest path: any path that has a point 
on the negative cycle can be made cheaper by one more walk around the negative cycle. In such 
a case, the Bellman–Ford algorithm can detect and report the negative cycle." 
\textbf{[Wikipedia, Bellman–Ford algorithm]}

\tcode{find_negative_cycle} can be called after calling \tcode{bellman_ford_shortest_paths} to
get the vertex ids of the negative weight cycle.

\subsubsection{Bellman-Ford Shortest Paths}

\paragraph{Single-Source Shortest Paths}
{\small
      \lstinputlisting{D3128_Algorithms/src/bellman_shortest_paths.hpp}
}

\paragraph{Multi-Source Shortest Paths}
{\small
      \lstinputlisting{D3128_Algorithms/src/bellman_shortest_paths_multi.hpp}
}

\begin{itemdescr}
      \pnum\mandates
            \begin{itemize}
                  \item \tcode{G} satisfies \tcode{adjacency_list<G>}.
                  \item \tcode{DistanceFn} satisfies \tcode{distance_fn_for<DistanceFn, G>}
                        with \tcode{is_arithmetic_v<distance_fn_value_t<DistanceFn, G>>}.
                  \item \tcode{PredecessorFn} satisfies \tcode{predecessor_fn_for<PredecessorFn, G>}.
                  \item \tcode{WF} satisfies \tcode{basic_edge_weight_function<G, WF, distance_fn_value_t<DistanceFn, G>, Compare, Combine>}.
            \end{itemize}
      \pnum\constantwhen
            \begin{itemize}
                  \item All operations on the distance, predecessor, and weight types
                        are constant subexpressions, and no visitor callback throws.
            \end{itemize}
      \pnum\preconditions
            \begin{itemize}
                  \item
                        Each \lstinline{source} in \lstinline{sources} is a valid vertex ID in \lstinline{g},
                        for the multi-source version.
                  \item
                        \lstinline{distance(g, vid) = infinite_distance()} for every vertex \lstinline{vid} in \lstinline{g}.
                  \item
                        \lstinline{predecessor(g, vid) = vid} for every vertex \lstinline{vid} in \lstinline{g}.
            \end{itemize}
      \pnum\hardprecond
            \begin{itemize}
                  \item \lstinline{source} is a valid vertex ID in \lstinline{g} for the single-source version.
            \end{itemize}
      \pnum\effects
            \begin{itemize}
                  \item
                        If vertex with index \lstinline{i} is reachable from vertex \lstinline{source}, then
                        \lstinline{distance(g, i)} will contain the distance from \lstinline{source} to vertex
                        \lstinline{i}.  Otherwise \lstinline{distance(g, i)} will contain
                        \lstinline{infinite_distance()}.
                  \item
                        If vertex with index \lstinline{i} is reachable
                        from vertex \lstinline{source}, then \lstinline{predecessor(g, i)} will contain the
                        predecessor vertex of vertex \lstinline{i}. Otherwise \lstinline{predecessor(g, i)} will contain
                        \lstinline{i}.
                  \item Does not modify the graph \lstinline{g}.
                  \item Member functions on the \lstinline{visitor} parameter are called during the algorithm's execution.
                        The functions are optional and, when included, must follow the visitor concepts for the events.
                        No overhead is incurred if the functions are not included.
                        The events supported are \lstinline{on_examine_edge}, \lstinline{on_edge_relaxed}, \lstinline{on_edge_not_relaxed},
                        \lstinline{on_edge_minimized}, and \lstinline{on_edge_not_minimized}.
            \end{itemize}
      \pnum\postconditions
            \begin{itemize}
                  \item \lstinline{distance(g, s) == 0} for all sources \lstinline{s}.
                  \item If no negative cycle is detected: for every reachable vertex \lstinline{v},
                        \lstinline{distance(g, v)} contains the shortest distance from the nearest source.
                  \item If a negative cycle is detected, \lstinline{distance} and \lstinline{predecessor}
                        may contain intermediate values.
                  \item For every unreachable vertex \lstinline{v},
                        \lstinline{distance(g, v) == infinite_distance<distance_fn_value_t<DistanceFn,G>>()}.
            \end{itemize}
      \pnum\returns 
            \begin{itemize}
                  \item \lstinline{optional<vertex_id_t<G>>} If no negative weight cycle is found, 
                        there is no associated vertex id. If a negative weight cycle is found, a
                        vertex id in the cycle is returned. \lstinline{find_negative_cycle}
                        can be called to get the vertex ids of the cycle.
                  \item Attribute: \lstinline{[[nodiscard]]}.
            \end{itemize}
      \pnum\throws
            \begin{itemize}
                  \item \lstinline{std::out_of_range} if any source vertex ID is not in \lstinline{vertices(g)}.
                  \item \textbf{Exception guarantee:} Basic --- \lstinline{g} remains unchanged;
                        \lstinline{distances} and \lstinline{predecessors} may be partially modified.
            \end{itemize}
      \pnum\complexity
            \begin{itemize}
                  \item \textbf{Time:} $\mathcal{O}(|E| \cdot |V|)$. Complexity may also be affected when visitor events are called.
                  \item \textbf{Space:} $\mathcal{O}(1)$ auxiliary beyond caller-provided arrays.
            \end{itemize}
      \pnum\remarks 
            \begin{itemize}
                  \item Duplicate sources do not affect the algorithm's complexity or correctness.
                  \item Unlike Dijkstra's algorithm, Bellman-Ford allows negative edge weights.
                        Performance constraints limit this to smaller graphs.
                  \item \tcode{DistanceFn} and \tcode{PredecessorFn} must satisfy \tcode{distance_fn_for}
                        and \tcode{predecessor_fn_for} respectively.  Wrap containers with
                        \tcode{container_value_fn} (see §\textit{Vertex Property Function Concepts}).
            \end{itemize}
      %\pnum\errors
\end{itemdescr}

\subsubsection{Bellman-Ford Shortest Distances}

This is the same as \textit{Shortest Paths} except that it excludes the predecessors,
giving a small performance improvement with lower memory overhead.

\paragraph{Single-Source Shortest Distances}
{\small
      \lstinputlisting{D3128_Algorithms/src/bellman_shortest_dists.hpp}
}

\paragraph{Multi-Source Shortest Distances}
{\small
      \lstinputlisting{D3128_Algorithms/src/bellman_shortest_dists_multi.hpp}
}

\begin{itemdescr}
      \pnum\mandates
            \begin{itemize}
                  \item \tcode{G} satisfies \tcode{adjacency_list<G>}.
                  \item \tcode{DistanceFn} satisfies \tcode{distance_fn_for<DistanceFn, G>}
                        with \tcode{is_arithmetic_v<distance_fn_value_t<DistanceFn, G>>}.
                  \item \tcode{WF} satisfies \tcode{basic_edge_weight_function<G, WF, distance_fn_value_t<DistanceFn, G>, Compare, Combine>}.
            \end{itemize}
      \pnum\constantwhen
            \begin{itemize}
                  \item All operations on the distance and weight types
                        are constant subexpressions, and no visitor callback throws.
            \end{itemize}
      \pnum\preconditions
            \begin{itemize}
                  \item
                        Each \lstinline{source} in \lstinline{sources} is a valid vertex ID in \lstinline{g},
                        for the multi-source version.
                  \item
                        \lstinline{distance(g, vid) = infinite_distance()} for every vertex \lstinline{vid} in \lstinline{g}.
            \end{itemize}
      \pnum\hardprecond
            \begin{itemize}
                  \item
                        \lstinline{source} is a valid vertex ID in \lstinline{g} for the single-source version.
            \end{itemize}
      \pnum\effects
            \begin{itemize}
                  \item
                        If vertex with index \lstinline{i} is reachable from vertex \lstinline{source}, then
                        \lstinline{distance(g, i)} will contain the distance from \lstinline{source} to vertex
                        \lstinline{i}.  Otherwise \lstinline{distance(g, i)} will contain
                        \lstinline{infinite_distance()}.
                  \item Does not modify the graph \lstinline{g}.
                  \item Member functions on the \lstinline{visitor} parameter are called during the algorithm's execution.
                        The functions are optional and, when included, must follow the visitor concepts for the events.
                        No overhead is incurred if the functions are not included.
                        The events supported are \lstinline{on_examine_edge}, \lstinline{on_edge_relaxed}, \lstinline{on_edge_not_relaxed},
                        \lstinline{on_edge_minimized}, and \lstinline{on_edge_not_minimized}.
            \end{itemize}
      \pnum\postconditions
            \begin{itemize}
                  \item \lstinline{distance(g, s) == 0} for all sources \lstinline{s}.
                  \item If no negative cycle is detected: for every reachable vertex \lstinline{v},
                        \lstinline{distance(g, v)} contains the shortest distance from the nearest source.
                  \item For every unreachable vertex \lstinline{v},
                        \lstinline{distance(g, v) == infinite_distance<distance_fn_value_t<DistanceFn,G>>()}.
            \end{itemize}
      \pnum\returns 
            \begin{itemize}
                  \item \lstinline{optional<vertex_id_t<G>>} If no negative weight cycle is found, 
                        there is no associated vertex id. If a negative weight cycle is found, a
                        vertex id in the cycle is returned. \lstinline{bellman_ford_shortest_paths} must be used
                        to get the predecessors if it is important to get the vertex ids of the cycle
                        using \lstinline{find_negative_cycle}.
                  \item Attribute: \lstinline{[[nodiscard]]}.
            \end{itemize}
      \pnum\throws
            \begin{itemize}
                  \item \lstinline{std::out_of_range} if any source vertex ID is not in \lstinline{vertices(g)}.
                  \item \textbf{Exception guarantee:} Basic --- \lstinline{g} remains unchanged;
                        \lstinline{distances} may be partially modified.
            \end{itemize}
      \pnum\complexity
            \begin{itemize}
                  \item \textbf{Time:} $\mathcal{O}(|E| \cdot |V|)$. Complexity may also be affected when visitor events are called.
                  \item \textbf{Space:} $\mathcal{O}(1)$ auxiliary beyond caller-provided arrays.
            \end{itemize}
      \pnum\remarks 
            \begin{itemize}
                  \item Duplicate sources do not affect the algorithm's complexity or correctness.
                  \item Unlike Dijkstra's algorithm, Bellman-Ford allows negative edge weights.
                        Performance constraints limit this to smaller graphs.
                  \item \tcode{DistanceFn} must satisfy \tcode{distance_fn_for<DistanceFn,G>}.  Wrap a
                        container with \tcode{container_value_fn}
                        (see §\textit{Vertex Property Function Concepts}).
            \end{itemize}
      %\pnum\errors
\end{itemdescr}

\subsubsection{Finding the Negative Cycle}
If a cycle with negative weights is found, it's possible to get the vertex ids of the cycle
using \tcode{find_negative_cycle} after calling \tcode{bellman_ford_shortest_paths}. It is
not possible to get the cycle from \tcode{bellman_ford_shortest_distances} because it does
not evaluate predecessors.

{\small
      \lstinputlisting{D3128_Algorithms/src/find_negative_cycle.hpp}
}
\begin{itemdescr}
      \pnum\mandates
            \begin{itemize}
                  \item \tcode{G} satisfies \tcode{adjacency_list<G>}.
                  \item \tcode{Predecessors} satisfies \tcode{forward_range}.
                  \item \tcode{OutputIterator} satisfies \tcode{output_iterator<OutputIterator, vertex_id_t<G>>}.
            \end{itemize}
      \pnum\preconditions
            \begin{itemize}
                  \item \lstinline{predecessors} must be evaluated by \lstinline{bellman_ford_shortest_paths}.
                  \item \lstinline{cycle_vertex_id} is the return value of \lstinline{bellman_ford_shortest_paths}.
            \end{itemize}
      \pnum\effects
            \begin{itemize}
                  \item
                        All vertex ids in the negative weight cycle are written to the \lstinline{out_cycle} output
                        iterator.
            \end{itemize}
      %\pnum\result
      %\pnum\returns 
      %\pnum\throws 
      \pnum\complexity
            \begin{itemize}
                  \item \textbf{Time:} $\mathcal{O}(|E|+|V|)$.
                  \item \textbf{Space:} $\mathcal{O}(|V|)$ auxiliary --- cycle-tracking array.
            \end{itemize}
      %\pnum\remarks 
      %\pnum\errors
\end{itemdescr}

\section{Clustering}
\subsection{Triangle Counting}
Compute the number of triangles in a graph. A triangle is a set of three vertices $\{u, v, w\}$
where each pair is connected by an edge. Two functions are provided:
\tcode{triangle_count} counts triangles (3-cliques) in undirected graphs, and
\tcode{directed_triangle_count} counts directed 3-cycles in directed graphs.

Both algorithms use a merge-based set intersection approach on sorted adjacency lists,
which is more efficient than nested-loop or hash-based methods for sparse graphs.
The algorithms require the graph to satisfy both \tcode{adjacency_list<G>} and
\tcode{ordered_vertex_edges<G>}.

\subsubsection{Triangle Count (Undirected)}

Count the number of triangles (3-cliques) in an undirected graph with sorted adjacency lists.

\begin{table}[ht]
\setcellgapes{3pt}
\makegapedcells
\centering
\begin{tabular}{|P{0.30\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|}
\hline
      \multirowcell{2}{
            \textbf{Complexity} \\
                  $\mathcal{O}(m^{3/2})$ \\
            }
      & \textbf{Directed?} No & \textbf{Cycles?} Yes & \textbf{Throws?} No \\
      & \textbf{Multi-edge?} No & \textbf{Self-loops} No & \\
\hline
\end{tabular}
%\caption{Algorithm Example}
\end{table}

{\small
      \lstinputlisting[firstline=4,lastline=6]{D3128_Algorithms/src/tc.hpp}
}
\begin{itemdescr}
      \pnum\mandates
            \begin{itemize}
                  \item \tcode{G} satisfies \tcode{adjacency_list<G>}.
                  \item \tcode{G} satisfies \tcode{ordered_vertex_edges<G>}: the outgoing edges of each vertex
                        must be sorted in ascending order of \tcode{target_id}.
            \end{itemize}
      \pnum\preconditions
            \begin{itemize}
                  \item The graph represents an undirected graph, with edges stored bidirectionally
                        (both $(u,v)$ and $(v,u)$).
                  \item Adjacency lists are sorted by \tcode{target_id} in ascending order.
            \end{itemize}
      \pnum\effects
            \begin{itemize}
                  \item For each edge $(u,v)$ where \tcode{vertex_id(g, u)} $<$ \tcode{vertex_id(g, v)}:
                  \begin{enumerate}
                        \item Get the sorted adjacency lists for $u$ and $v$.
                        \item Perform a merge-based intersection of the two lists
                              (analogous to \tcode{std::set_intersection}).
                        \item For each common neighbor $w$ where \tcode{vertex_id(g, w)} $>$ \tcode{vertex_id(g, v)},
                              increment the triangle count.
                  \end{enumerate}
                  \item The ordering constraint
                        \tcode{vertex_id(g, u)} $<$ \tcode{vertex_id(g, v)} $<$ \tcode{vertex_id(g, w)}
                        guarantees each triangle is counted exactly once.
            \end{itemize}
      %\pnum\result
      \pnum\returns 
            \begin{itemize}
                  \item \lstinline{[[nodiscard]]}. The total number of triangles in the graph.
                  \item Returns 0 for empty graphs or graphs with fewer than 3 vertices.
            \end{itemize}
      \pnum\throws This function is \tcode{noexcept}.
      \pnum\complexity
            \begin{itemize}
                  \item \textbf{Time:} $\mathcal{O}(m^{3/2})$ average, where $m = |E|$. Best case
                        $\mathcal{O}(|V|+|E|)$ for triangle-free graphs. Worst case
                        $\mathcal{O}(|V| \cdot d_{\max}^2)$ where $d_{\max}$ is the maximum vertex degree.
                  \item \textbf{Space:} $\mathcal{O}(1)$ auxiliary (excluding graph storage).
            \end{itemize}
      \pnum\remarks 
            \begin{itemize}
                  \item Edge weights, if present, are ignored.
                  \item Self-loops are ignored (a vertex cannot form a triangle with itself).
                  \item Multi-edges: each parallel edge contributes to the triangle count separately.
                  \item Works correctly on both connected and disconnected graphs.
                  \item The \tcode{ordered_vertex_edges} requirement enables $\mathcal{O}(d)$ set intersection
                        per edge rather than $\mathcal{O}(d^2)$ nested loops, which is critical for
                        performance on high-degree vertices. Graph types using \tcode{std::set} or
                        \tcode{std::map} edge containers (e.g.\ \tcode{vos}, \tcode{uos}, \tcode{dos},
                        \tcode{mos}) naturally satisfy this requirement.
            \end{itemize}
      %\pnum\errors
\end{itemdescr}

\subsubsection{Directed Triangle Count}

Count the number of directed 3-cycles in a directed graph with sorted adjacency lists.
A directed 3-cycle is a set of three vertices $\{u, v, w\}$ where directed edges
$u \rightarrow v$, $v \rightarrow w$, and $u \rightarrow w$ all exist.

\begin{table}[ht]
\setcellgapes{3pt}
\makegapedcells
\centering
\begin{tabular}{|P{0.30\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|}
\hline
      \multirowcell{2}{
            \textbf{Complexity} \\
                  $\mathcal{O}(m^{3/2})$ \\
            }
      & \textbf{Directed?} Yes & \textbf{Cycles?} Yes & \textbf{Throws?} No \\
      & \textbf{Multi-edge?} No & \textbf{Self-loops} No & \\
\hline
\end{tabular}
%\caption{Algorithm Example}
\end{table}

{\small
      \lstinputlisting[firstline=4,lastline=6]{D3128_Algorithms/src/directed_triangle_count.hpp}
}
\begin{itemdescr}
      \pnum\mandates
            \begin{itemize}
                  \item \tcode{G} satisfies \tcode{adjacency_list<G>}.
                  \item \tcode{G} satisfies \tcode{ordered_vertex_edges<G>}: the outgoing edges of each vertex
                        must be sorted in ascending order of \tcode{target_id}.
            \end{itemize}
      \pnum\preconditions
            \begin{itemize}
                  \item Adjacency lists (out-neighbor lists) are sorted by \tcode{target_id} in ascending order.
            \end{itemize}
      \pnum\effects
            \begin{itemize}
                  \item For each vertex $u$, for each out-neighbor $v$ ($v \neq u$):
                  \begin{enumerate}
                        \item Get the sorted out-neighbor lists for $u$ and $v$.
                        \item Perform a merge-based intersection of the two lists.
                        \item For each common out-neighbor $w$ ($w \neq u$ and $w \neq v$),
                              increment the count.
                  \end{enumerate}
                  \item Each directed 3-cycle $(u \rightarrow v, v \rightarrow w, u \rightarrow w)$
                        is counted exactly once.
                  \item Self-loops are skipped.
            \end{itemize}
      %\pnum\result
      \pnum\returns 
            \begin{itemize}
                  \item \lstinline{[[nodiscard]]}. The total number of directed 3-cycles in the graph.
            \end{itemize}
      \pnum\throws This function is \tcode{noexcept}.
      \pnum\complexity
            \begin{itemize}
                  \item \textbf{Time:} $\mathcal{O}(m^{3/2})$ average, where $m = |E|$. Worst case
                        $\mathcal{O}(|V| \cdot d_{\max}^2)$ where $d_{\max}$ is the maximum out-degree.
                  \item \textbf{Space:} $\mathcal{O}(1)$ auxiliary (excluding graph storage).
            \end{itemize}
      \pnum\remarks 
            \begin{itemize}
                  \item For an undirected graph stored with bidirectional edges, this function
                        counts each undirected triangle 6 times (once per permutation of the 3
                        vertices). Use \tcode{triangle_count} instead for undirected graphs.
                  \item Unlike \tcode{triangle_count}, this function does not impose vertex ID
                        ordering constraints on the counted cycles.
                  \item Edge weights, if present, are ignored.
            \end{itemize}
      %\pnum\errors
\end{itemdescr}


\section{Communities}
\subsection{Label Propagation}
Propagate vertex labels for community detection by majority voting among neighbours.
Each iteration shuffles the vertex processing order, then sets every vertex's label to the most
popular label among its neighbours. Ties are broken randomly using the supplied random-number
generator. The algorithm iterates until no label changes (convergence) or until \tcode{max_iters}
iterations have been performed. $\mathcal{O}(M)$ complexity, where $M = |E|$, is based on one
iteration (each edge is examined once); the number of iterations required for convergence is
typically small relative to graph size.

A second overload accepts an \tcode{empty_label} sentinel for semi-supervised propagation:
vertices whose label equals \tcode{empty_label} are treated as unlabelled --- they do not vote
and are not counted in neighbour tallies, but can acquire a label from a labelled neighbour.
Vertices in components with no labelled vertex retain \tcode{empty_label} after convergence.

\begin{table}[ht]
\setcellgapes{3pt}
\makegapedcells
\centering
\begin{tabular}{|P{0.30\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|}
\hline
      \multirowcell{2}{
            \textbf{Complexity} \\
                  $\mathcal{O}(M)$ \\
            }
      & \textbf{Directed?} Yes & \textbf{Cycles?} Yes & \textbf{Throws?} Yes \\
      & \textbf{Multi-edge?} Yes & \textbf{Self-loops} Yes & \\
\hline
\end{tabular}
%\caption{Algorithm Example}
\end{table}

\paragraph{Basic Label Propagation}
{\small
      \lstinputlisting[firstline=4,lastline=14]{D3128_Algorithms/src/lp.hpp}
}

\paragraph{Label Propagation with Empty-Label Sentinel}
{\small
      \lstinputlisting[firstline=16,lastline=27]{D3128_Algorithms/src/lp.hpp}
}

\begin{itemdescr}
      \pnum\mandates
            \begin{itemize}
                  \item \tcode{G} satisfies \tcode{adjacency_list<G>}.
                  \item \tcode{LabelFn} satisfies \tcode{vertex_property_fn_for<LabelFn, G>}.
                  \item \tcode{vertex_fn_value_t<LabelFn, G>} satisfies
                        \tcode{std::equality_comparable}.
                  \item \tcode{Gen} satisfies \tcode{std::uniform_random_bit_generator}.
            \end{itemize}
      \pnum\preconditions
            \begin{itemize}
                  \item \lstinline{label(g, v)} returns a meaningful initial label for every vertex \lstinline{v} in \lstinline{g}.
                  \item For the sentinel overload, vertices to be treated as unlabelled must have
                        their label initialised to \lstinline{empty_label}.
            \end{itemize}
      \pnum\effects
            \begin{itemize}
                  \item Each iteration: shuffles the vertex processing order using \lstinline{rng},
                        then for each vertex $u$ in that order:
                  \begin{enumerate}
                        \item Tallies the labels of all neighbours of $u$ (skipping neighbours
                              with \lstinline{empty_label} in the sentinel overload).
                        \item Assigns \lstinline{label(g, u)} to the most frequent neighbour label,
                              breaking ties randomly via \lstinline{rng}.
                        \item Isolated vertices (no neighbours, or all neighbours unlabelled)
                              retain their current label.
                  \end{enumerate}
                  \item Iteration stops when no label changes in a full pass (convergence) or
                        \lstinline{max_iters} is reached.
                  \item The graph \lstinline{g} is not modified.
            \end{itemize}
      %\pnum\result
      %\pnum\returns
      \pnum\throws
            \begin{itemize}
                  \item \lstinline{std::bad_alloc} from internal frequency map and candidate vector allocations.
                  \item \textbf{Exception guarantee:} Basic --- \lstinline{g} remains unchanged;
                        \lstinline{label} may be partially modified.
            \end{itemize}
      \pnum\complexity
            \begin{itemize}
                  \item \textbf{Time:} $\mathcal{O}(|E|)$ per iteration; convergence typically
                        requires few iterations relative to graph size.
                  \item \textbf{Space:} $\mathcal{O}(|V|)$ auxiliary --- shuffled vertex-ID vector and
                        per-vertex frequency map.
            \end{itemize}
      \pnum\remarks
            \begin{itemize}
                  \item User is responsible for providing meaningful initial vertex labels.
                  \item Some label propagation implementations use vertex IDs as initial labels;
                        this is not done here because the label type can be more general than the 
                        vertex ID type.
                  \item Directed edges are followed as-is; for undirected semantics the graph
                        should store both $(u,v)$ and $(v,u)$.
                  \item Multi-edges and self-loops are counted in the neighbour tally and may
                        bias the label assignment.
                  \item When \lstinline{empty_label} is not present in \lstinline{label}, the
                        sentinel overload behaves identically to the basic overload.
            \end{itemize}
      %\pnum\errors
\end{itemdescr}

\section{Components}
\subsection{Articulation Points}
Find articulation points (cut vertices) of a graph. An articulation point is a vertex whose removal,
along with its incident edges, disconnects the graph into two or more connected components.
The algorithm uses an iterative Hopcroft-Tarjan DFS, tracking discovery times (\tcode{disc[v]}) and
low-link values (\tcode{low[v]}), defined as the minimum discovery time reachable from the subtree
rooted at $v$ via back-edges. A vertex $u$ is an articulation point under one of two rules:
\begin{itemize}
      \item \textbf{Root rule:} $u$ is the root of a DFS tree and has two or more DFS children.
      \item \textbf{Non-root rule:} $u$ is not a root and has a child $v$ with \tcode{low[v]} $\geq$ \tcode{disc[u]}.
\end{itemize}

\begin{table}[ht]
\setcellgapes{3pt}
\makegapedcells
\centering
\begin{tabular}{|P{0.30\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|}
\hline
      \multirowcell{2}{
            \textbf{Complexity} \\
                  $\mathcal{O}(|E|+|V|)$
            }
      & \textbf{Directed?} Yes & \textbf{Cycles?} Yes & \textbf{Throws?} Yes \\
      & \textbf{Multi-edge?} Yes & \textbf{Self-loops} Yes & \\
\hline
\end{tabular}
%\caption{Algorithm Example}
\end{table}

{\small
     \lstinputlisting[firstline=4,lastline=6]{D3128_Algorithms/src/connected_components.hpp}
}

\begin{itemdescr}
      \pnum\mandates
            \begin{itemize}
                  \item \tcode{G} satisfies \tcode{adjacency_list<G>}.
                  \item \tcode{Iter} satisfies \tcode{std::output\_iterator<Iter, vertex\_id\_t<G>>}.
            \end{itemize}
      \pnum\preconditions
            \begin{itemize}
                  \item For undirected graph semantics, each edge $\{u,v\}$ must be stored as
                        both $(u,v)$ and $(v,u)$ in \lstinline{g}.
            \end{itemize}
      \pnum\effects
            \begin{itemize}
                  \item Performs an iterative DFS over all vertices (including disconnected components).
                  \item Writes each articulation point vertex ID to \lstinline{cut_vertices} exactly once.
                  \item \lstinline{g} is not modified.
                  \item Self-loops are ignored and do not affect articulation point detection.
                  \item Multi-edges: only the first reverse edge to the DFS parent is treated as the
                        tree edge; additional parallel edges are treated as back-edges that update
                        low-link values.
            \end{itemize}
      %\pnum\result
      %\pnum\returns
      \pnum\throws
            \begin{itemize}
                  \item \lstinline{std::bad_alloc} if internal vector allocation fails.
                  \item \textbf{Exception guarantee:} Basic --- \lstinline{g} remains unchanged;
                        the output iterator may be partially written.
            \end{itemize}
      \pnum\complexity
            \begin{itemize}
                  \item \textbf{Time:} $\mathcal{O}(|V|+|E|)$ --- each vertex and edge is visited exactly once.
                  \item \textbf{Space:} $\mathcal{O}(|V|)$ auxiliary --- discovery time, low-link, parent,
                        child-count, and emitted arrays, plus DFS stack.
            \end{itemize}
      \pnum\remarks
            \begin{itemize}
                  \item There is no ordering guarantee on the articulation points written to
                        \lstinline{cut_vertices}.
                  \item Disconnected graphs are handled correctly; the outer loop restarts DFS
                        from each unvisited vertex.
                  \item The optional \lstinline{alloc} parameter supplies an allocator used for
                        the internal DFS stack and discovery/low-link arrays.  Defaults to
                        \tcode{std::allocator<std::byte>}.
            \end{itemize}
      %\pnum\errors
\end{itemdescr}

\subsection{BiConnected Components}
Find the biconnected components of a graph. A biconnected component (also called a
2-connected component) is a maximal biconnected subgraph --- one that is connected and has
no articulation points. Equivalently, any two vertices in a biconnected component lie on a
common simple cycle.

The algorithm extends the Hopcroft-Tarjan DFS with an explicit edge stack. During the DFS,
each tree edge and back edge is pushed onto the edge stack. When a biconnected-component
boundary is detected on backtrack (i.e., \tcode{low[v]} $\geq$ \tcode{disc[u]} for child $v$
and parent $u$), the edge stack is popped down to and including edge $(u,v)$ and the unique
vertex IDs from those edges form one biconnected component. Isolated vertices are emitted as
trivial single-vertex components. Articulation-point vertices appear in more than one component.

\begin{table}[ht]
\setcellgapes{3pt}
\makegapedcells
\centering
\begin{tabular}{|P{0.30\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|}
\hline
      \multirowcell{2}{
            \textbf{Complexity} \\
                  $\mathcal{O}(|E|+|V|)$
            }
      & \textbf{Directed?} Yes & \textbf{Cycles?} Yes & \textbf{Throws?} Yes \\
      & \textbf{Multi-edge?} Yes & \textbf{Self-loops} Yes & \\
\hline
\end{tabular}
%\caption{Algorithm Example}
\end{table}

{\small
     \lstinputlisting[firstline=11,lastline=12]{D3128_Algorithms/src/connected_components.hpp}
}

\begin{itemdescr}
      \pnum\mandates
            \begin{itemize}
                  \item \tcode{G} satisfies \tcode{adjacency_list<G>}.
                  \item \tcode{OuterContainer} supports \tcode{push_back} with an inner container
                        type constructible from a pair of iterators over \tcode{vertex_id_t<G>}.
            \end{itemize}
      \pnum\preconditions
            \begin{itemize}
                  \item For undirected semantics, each edge $\{u,v\}$ must be stored as both $(u,v)$
                        and $(v,u)$ in \lstinline{g}.
            \end{itemize}
      \pnum\effects
            \begin{itemize}
                  \item Performs an iterative Hopcroft-Tarjan DFS over all vertices.
                  \item For each biconnected component detected, one inner container of vertex IDs
                        is \tcode{push_back}'d into \tcode{components}.
                  \item Isolated vertices (degree~0) are emitted as trivial single-vertex components.
                  \item Every vertex appears in at least one component; articulation-point vertices
                        appear in more than one component.
                  \item Self-loops are ignored and do not affect component detection.
                  \item Multi-edges: only the first reverse edge to the DFS parent is treated as
                        the tree edge; additional parallel edges are treated as back-edges.
                  \item \lstinline{g} is not modified.
            \end{itemize}
      %\pnum\result
      %\pnum\returns
      \pnum\throws
            \begin{itemize}
                  \item \lstinline{std::bad_alloc} from internal vector, set, or stack allocations.
                  \item \textbf{Exception guarantee:} Basic --- \lstinline{g} remains unchanged;
                        \lstinline{components} may be partially written.
            \end{itemize}
      \pnum\complexity
            \begin{itemize}
                  \item \textbf{Time:} $\mathcal{O}(|V|+|E|)$ --- each vertex and edge is visited exactly once.
                  \item \textbf{Space:} $\mathcal{O}(|V|+|E|)$ auxiliary --- discovery/low-link arrays
                        and parent map ($\mathcal{O}(|V|)$), DFS frame stack ($\mathcal{O}(|V|)$),
                        and edge stack ($\mathcal{O}(|E|)$).
            \end{itemize}
      \pnum\remarks
            \begin{itemize}
                  \item There is no ordering guarantee on the components or on vertex IDs within
                        a component.
                  \item Disconnected graphs are handled correctly; the outer loop restarts DFS
                        from each unvisited vertex.
                  \item The implementation uses iterative DFS with stored edge iterators to avoid
                        recursion-depth limits and $\mathcal{O}(\text{degree})$ re-scans on resume.
                  \item The optional \lstinline{alloc} parameter supplies an allocator used for
                        the internal DFS stack, edge stack, and discovery/low-link arrays.  Defaults
                        to \tcode{std::allocator<std::byte>}.
            \end{itemize}
      %\pnum\errors
\end{itemdescr}

\subsection{Connected Components}
Find weakly connected components of a graph. Weakly connected components are subgraphs where a path exists between all pairs of vertices when ignoring edge direction.

\begin{table}[ht]
\setcellgapes{3pt}
\makegapedcells
\centering
\begin{tabular}{|P{0.30\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|}
\hline
      \multirowcell{2}{
            \textbf{Complexity} \\
                  $\mathcal{O}(|E|+|V|)$
            }
      & \textbf{Directed?} No & \textbf{Cycles?} Yes & \textbf{Throws?} No \\
      & \textbf{Multi-edge?} No & \textbf{Self-loops} Yes & \\
\hline
\end{tabular}
%\caption{Algorithm Example}
\end{table}

{
\small
     \lstinputlisting[firstline=17,lastline=19]{D3128_Algorithms/src/connected_components.hpp}
}

\begin{itemdescr}
      \pnum\mandates
            \begin{itemize}
                  \item \tcode{G} satisfies \tcode{adjacency_list<G>}.
                  \item \tcode{ComponentFn} satisfies \tcode{vertex_property_fn_for<ComponentFn, G>}.
            \end{itemize}
      \pnum\preconditions
            \begin{itemize}
                  \item
                        \lstinline{component(g, v)} is callable for each vertex \lstinline{v} in \lstinline{g}.
            \end{itemize}
      \pnum\effects
            \begin{itemize}
                  \item
                        \lstinline{component(g, v)} is set to the connected component id of vertex \lstinline{v}.
                  \item
                        There is at least one Connected Component, with component id of \lstinline{0}, for \lstinline{num_vertices(g) > 0}.
            \end{itemize}
      %\pnum\result
      \pnum\returns The number of connected components found.
      %\pnum\throws
      \pnum\complexity
            \begin{itemize}
                  \item \textbf{Time:} $\mathcal{O}(|E|+|V|)$.
                  \item \textbf{Space:} $\mathcal{O}(|V|)$ auxiliary --- component array and DFS stack.
            \end{itemize}
      \pnum\remarks 
            \begin{itemize}
                  \item
                        If the maximum component id is 0, then all vertices are in a single connected component.
                  \item
                        If the maximum component id is \lstinline{num_vertices(g)-1}, each vertex is the sole vertex in its 
                        own connected component.
                  \item The optional \lstinline{alloc} parameter supplies an allocator used for
                        the internal DFS stack.  Defaults to \tcode{std::allocator<std::byte>}.
            \end{itemize}
      %\pnum\errors
\end{itemdescr}

% Afforest Connected Components is intentionally excluded from this proposal.
% It is a parallel sampling-based algorithm that belongs in a future parallel-algorithms paper.
% Do not re-add it here.

\subsection{Strongly Connected Components}
\subsubsection{Kosaraju}
Find strongly connected components of a graph using Kosaraju's algorithm. Strongly connected components are subgraphs where a path exists between all pairs of vertices.

\begin{table}[ht]
\setcellgapes{3pt}
\makegapedcells
\centering
\begin{tabular}{|P{0.30\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|}
\hline
      \multirowcell{2}{
            \textbf{Complexity} \\
                  $\mathcal{O}(|E|+|V|)$
            }
      & \textbf{Directed?} Yes & \textbf{Cycles?} Yes & \textbf{Throws?} No \\
      & \textbf{Multi-edge?} No & \textbf{Self-loops} Yes & \\
\hline
\end{tabular}
%\caption{Algorithm Example}
\end{table}

{\small
      \lstinputlisting[firstline=35, lastline=43]{D3128_Algorithms/src/connected_components.hpp}
}

\begin{itemdescr}
      \pnum\mandates
            \begin{itemize}
                  \item \tcode{G} satisfies \tcode{adjacency_list<G>}.
                  \item \tcode{ComponentFn} satisfies \tcode{vertex_property_fn_for<ComponentFn, G>}.
            \end{itemize}
      \pnum\preconditions
            \begin{itemize}
                  \item
                        For the two-graph overload, \lstinline{g_t} is the transpose of \lstinline{g}, where edge \lstinline{uv} in \lstinline{g} implies edge \lstinline{vu} in \lstinline{g_t}.
            \end{itemize}
      \pnum\hardprecond
            \begin{itemize}
                  \item
                        \lstinline{num_vertices(g) == num_vertices(g_t)} for the two-graph overload.
            \end{itemize}
      \pnum\effects
            \begin{itemize}
                  \item
                        \lstinline{component(g, v)} is set to the strongly connected component id of vertex \lstinline{v}.
                  \item
                        The component id is in the range \lstinline{0 <= component(g, v) < num_vertices(g)}.
            \end{itemize}
      %\pnum\result
      %\pnum\returns
      %\pnum\throws
      \pnum\complexity
            \begin{itemize}
                  \item \textbf{Time:} $\mathcal{O}(|E|+|V|)$.
                  \item \textbf{Space:} $\mathcal{O}(|V|)$ auxiliary --- visited array and finish order.
            \end{itemize}
      \pnum\remarks 
            \begin{itemize}
                  \item
                        If the maximum component id is 0, then all vertices are in a single strongly connected component.
                  \item
                        If the maximum component id is \lstinline{num_vertices(g)-1}, each vertex is the sole vertex in its 
                        own strongly connected component.
                  \item The optional \lstinline{alloc} parameter supplies an allocator used for
                        the internal finish-order vector and DFS stacks.  Defaults to
                        \tcode{std::allocator<std::byte>}.
            \end{itemize}
      %\pnum\errors
\end{itemdescr}

\subsubsection{Tarjan's SCC}
Find strongly connected components of a directed graph using Tarjan's algorithm.
A strongly connected component (SCC) is a maximal set of vertices such that there is a
directed path from every vertex in the set to every other vertex in the set.

\begin{table}[ht]
\setcellgapes{3pt}
\makegapedcells
\centering
\begin{tabular}{|P{0.30\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|}
\hline
      \multirowcell{2}{
            \textbf{Complexity} \\
                  $\mathcal{O}(|E|+|V|)$
            }
      & \textbf{Directed?} Yes & \textbf{Cycles?} Yes & \textbf{Throws?} Yes \\
      & \textbf{Multi-edge?} Yes & \textbf{Self-loops} Yes & \\
\hline
\end{tabular}
\end{table}

{\small
      \lstinputlisting{D3128_Algorithms/src/tarjan_scc.hpp}
}

\begin{itemdescr}
      \pnum\mandates
            \begin{itemize}
                  \item \tcode{G} satisfies \tcode{adjacency_list<G>}.
                  \item \tcode{ComponentFn} satisfies \tcode{vertex_property_fn_for<ComponentFn, G>}.
            \end{itemize}
      \pnum\preconditions
            \begin{itemize}
                  \item \tcode{component} must be callable for each vertex of \tcode{g}.
            \end{itemize}
      \pnum\hardprecond
            \begin{itemize}
                  \item \tcode{g} must be a directed graph.
            \end{itemize}
      \pnum\effects
            \begin{itemize}
                  \item \lstinline{component(g, uid)} is set to the SCC id of vertex \lstinline{uid}
                        for every vertex in \lstinline{g}.
                  \item Component ids are assigned in the range
                        \lstinline{0 <= component(g, uid) < num_vertices(g)}, numbered in
                        reverse topological order of the condensed SCC DAG.
                  \item Does not modify the graph \lstinline{g}.
            \end{itemize}
      \pnum\returns The number of strongly connected components found.
      \pnum\throws
            \begin{itemize}
                  \item \lstinline{std::bad_alloc} if internal allocations (discovery time, low-link,
                        on-stack flag, DFS stack, or SCC stack) fail.
                  \item \textbf{Exception guarantee:} Basic --- \lstinline{g} is unchanged;
                        partial component assignments may have occurred.
            \end{itemize}
      \pnum\complexity
            \begin{itemize}
                  \item \textbf{Time:} $\mathcal{O}(|V|+|E|)$ --- single DFS pass, each vertex
                        and edge visited at most once.
                  \item \textbf{Space:} $\mathcal{O}(|V|)$ auxiliary --- discovery time, low-link,
                        on-stack flag arrays, DFS stack, and SCC stack.
            \end{itemize}
      \pnum\remarks
            \begin{itemize}
                  \item Compared to \tcode{kosaraju}, Tarjan's algorithm requires only a single
                        DFS pass and does not require (or accept) a transpose graph. It therefore
                        works on any \tcode{adjacency_list} graph, whereas the single-graph
                        overload of \tcode{kosaraju} requires \tcode{bidirectional_adjacency_list}.
                  \item Uses an iterative DFS with an explicit stack to avoid recursion-depth limits.
                  \item A vertex \lstinline{u} is the root of an SCC when
                        \lstinline{disc[u] == low[u]} after all of its outgoing edges have been
                        processed.
                  \item Low-link values track the earliest reachable discovery time in the
                        current DFS subtree.  Back and cross edges to vertices already in a
                        \emph{completed} SCC do not update low-link values.
                  \item If the return value is \lstinline{1}, all vertices belong to a single
                        strongly connected component (the graph is strongly connected).
                  \item If the return value equals \lstinline{num_vertices(g)}, every vertex
                        is its own SCC (a DAG).
                  \item The optional \lstinline{alloc} parameter supplies an allocator used for
                        the internal DFS stack, SCC stack, and discovery/low-link arrays.  Defaults
                        to \tcode{std::allocator<std::byte>}.
            \end{itemize}
\end{itemdescr}

\section{Maximal Independent Set}
\subsection{Maximal Independent Set}
Find a maximally independent set of vertices in a graph starting from a source vertex. An independent vertex set indicates no pair of vertices in the set are adjacent.

\begin{table}[ht]
\setcellgapes{3pt}
\makegapedcells
\centering
\begin{tabular}{|P{0.30\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|}
\hline
      \multirowcell{2}{
            \textbf{Complexity} \\
                  $\mathcal{O}(|E|)$
            }
      & \textbf{Directed?} Yes & \textbf{Cycles?} No & \textbf{Throws?} No \\
      & \textbf{Multi-edge?} No & \textbf{Self-loops} No & \\
\hline
\end{tabular}
%\caption{Algorithm Example}
\end{table}

{\small
      \lstinputlisting{D3128_Algorithms/src/mis.hpp}
}
\begin{itemdescr}
      \pnum\mandates
            \begin{itemize}
                  \item \tcode{G} satisfies \tcode{adjacency_list<G>}.
                  \item \tcode{Iter} satisfies \tcode{output_iterator<Iter, vertex_id_t<G>>}.
            \end{itemize}
      \pnum\hardprecond
            \begin{itemize}
                  \item
                        \lstinline{0 <= seed < num_vertices(graph)}.
            \end{itemize}
      \pnum\effects
            \begin{itemize}
                  \item
                        Output iterator \lstinline{mis} contains the maximal independent set of vertices that
                        includes \lstinline{seed}, which is a subset of \lstinline{vertices(graph)}.
            \end{itemize}
      %\pnum\result
      \pnum\returns The number of vertices in the maximal independent set.
      %\pnum\throws
      \pnum\complexity $\mathcal{O}(|E|)$ \\
      %\pnum\remarks 
      %\pnum\errors
\end{itemdescr}

\section{Link Analysis}

% pagerank removed because there are many variations and corner cases to cover
% It fits better in an example. 08-Nov-2023

\subsection{Jaccard Coefficient}
Calculate the Jaccard coefficient of a graph

\begin{table}[ht]
\setcellgapes{3pt}
\makegapedcells
\centering
\begin{tabular}{|P{0.30\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|}
\hline
      \multirowcell{2}{
            \textbf{Complexity} \\
                  $\mathcal{O}(|N|^3)$
            }
      & \textbf{Directed?} Yes & \textbf{Cycles?} No & \textbf{Throws?} No \\
      & \textbf{Multi-edge?} No & \textbf{Self-loops} No & \\
\hline
\end{tabular}
%\caption{Algorithm Example}
\end{table}

{\small
      \lstinputlisting[firstline=4,lastline=6]{D3128_Algorithms/src/jaccard.hpp}
}

\phil{Consider using \tcode{out(uid,vid,uv,val)} as the function descriptor in Preconditions to make it more readable.}

\phil{Would an output iterator be appropriate? \tcode{pair<pair<vertex_id_t<G>,vertex_id_t<G>>,T>} might work for the output type. This starts to get into the same realm of the views
as to whether the edge reference is useful by the consumer or not (e.g. basic\_ vs. regular versions).}

\begin{itemdescr}
      \pnum\mandates
            \begin{itemize}
                  \item \tcode{G} satisfies \tcode{adjacency_list<G>}.
                  \item \tcode{OutOp} satisfies \tcode{invocable<OutOp, vertex_id_t<G>, vertex_id_t<G>, edge_t<G>\&, T>}.
            \end{itemize}
      \pnum\preconditions
            \begin{itemize}
                  \item 
                        \lstinline{out} is an operator for setting the resulting Jaccard coefficient. This function 
                        is expected to be of the form \lstinline{out(vertex_id_t<G> uid, vertex_id_t<G> vid, edge_t<G> uv, T val)}.
            \end{itemize}
      \pnum\effects 
            \begin{itemize}
                  \item
                        For every pair of neighboring vertices \lstinline{(uid, vid)}, the function \lstinline{out} is called, passing the vertex ids, 
                        the edge \lstinline{uv} between them, and the calculated Jaccard coefficient.
            \end{itemize}
      %\pnum\result
      %\pnum\returns
      %\pnum\throws
      \pnum\complexity
            \begin{itemize}
                  \item \textbf{Time:} $\mathcal{O}(|V| + |E| \cdot d_{\min})$, where $d_{\min}$ is the minimum degree of either endpoint.
                  \item \textbf{Space:} $\mathcal{O}(|V|+|E|)$ auxiliary --- precomputed neighbor sets.
            \end{itemize}
      %\pnum\remarks 
      %\pnum\errors
\end{itemdescr}

\section{Minimum Spanning Tree}
%\andrew{Simplified ``driver'' for these also?}

\subsection{Kruskal Minimum Spanning Tree}
Find the minimum weight spanning tree of a graph using Kruskal's algorithm.

\begin{table}[ht]
\setcellgapes{3pt}
\makegapedcells
\centering
\begin{tabular}{|P{0.30\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|}
\hline
      \multirowcell{2}{
            \textbf{Complexity} \\
            $\mathcal{O}(|E|)$
            }
      & \textbf{Directed?} Yes & \textbf{Cycles?} No & \textbf{Throws?} No \\
      & \textbf{Multi-edge?} No & \textbf{Self-loops} No & \\
\hline
\end{tabular}
%\caption{Algorithm Example}
\end{table}

{\small
      \lstinputlisting[firstline=4,lastline=10]{D3128_Algorithms/src/mst.hpp}
}
\begin{itemdescr}
      \pnum\mandates
            \begin{itemize}
                  \item \tcode{IELR} satisfies \tcode{x_index_edgelist_range}.
                  \item \tcode{OELR} satisfies \tcode{x_index_edgelist_range}.
            \end{itemize}
      \pnum\preconditions
            \begin{itemize}
                  %\item
                  %      \lstinline{e} is an \lstinline{edgelist}.
                  \item
                        \lstinline{compare} operator is a valid comparison operation on two edge values of type \lstinline{range_value_t<EL>::value_type} 
                        which returns a bool.
            \end{itemize}
      \pnum\effects 
            \begin{itemize}
                  \item
                        Edgelist \lstinline{t} contains edges representing a spanning tree or forest, which minimize the comparison operator. 
                        When \lstinline{compare} is \lstinline{<}, \lstinline{t} represents a minimum weight spanning tree.
            \end{itemize}
      %\pnum\result
      \pnum\returns A \lstinline{pair<EV, size_t>} containing the total weight of the spanning tree or
            forest and the number of connected components.
      %\pnum\throws
      \pnum\complexity
            \begin{itemize}
                  \item \textbf{Time:} $\mathcal{O}(|E|\log{|E|})$.
                  \item \textbf{Space:} $\mathcal{O}(|E|+|V|)$ auxiliary --- edge copy and union-find.
            \end{itemize}
      \pnum\remarks The optional \lstinline{alloc} parameter supplies an allocator used for
            the internal disjoint-set (union-find) storage.  Defaults to
            \tcode{std::allocator<std::byte>}.
      %\pnum\errors
\end{itemdescr}

\subsection{Inplace Kruskal Minimum Spanning Tree}
Memory-efficient variant of Kruskal's algorithm that sorts the input edge list in place
rather than making a copy.  The input range must satisfy \lstinline{permutable<iterator_t<IELR>>}.

\begin{table}[ht]
\setcellgapes{3pt}
\makegapedcells
\centering
\begin{tabular}{|P{0.30\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|}
\hline
      \multirowcell{2}{
            \textbf{Complexity} \\
            $\mathcal{O}(|E|\log{|E|})$
            }
      & \textbf{Directed?} Yes & \textbf{Cycles?} No & \textbf{Throws?} No \\
      & \textbf{Multi-edge?} No & \textbf{Self-loops} No & \\
\hline
\end{tabular}
\label{tab:algo_inplace_kruskal}
\end{table}

{\small
      \lstinputlisting[firstline=15,lastline=23]{D3128_Algorithms/src/mst.hpp}
}
\begin{itemdescr}
      \pnum\mandates
            \begin{itemize}
                  \item \tcode{IELR} satisfies \tcode{x_index_edgelist_range}.
                  \item \tcode{OELR} satisfies \tcode{x_index_edgelist_range}.
                  \item \tcode{iterator_t<IELR>} satisfies \tcode{permutable}.
            \end{itemize}
      \pnum\preconditions
            \begin{itemize}
                  \item
                        \lstinline{compare} operator, when provided, is a valid comparison operation on two edge values
                        of type \lstinline{range_value_t<EL>::value_type} which returns a bool.
            \end{itemize}
      \pnum\effects 
            \begin{itemize}
                  \item
                        The input edge list \lstinline{e} is sorted by edge weight (or by \lstinline{compare}
                        when provided).
                  \item
                        Edgelist \lstinline{t} contains edges representing a spanning tree or forest, which
                        minimize the comparison operator.  When \lstinline{compare} is \lstinline{<},
                        \lstinline{t} represents a minimum weight spanning tree.
            \end{itemize}
      %\pnum\result
      \pnum\returns A \lstinline{pair<EV, size_t>} containing the total weight of the spanning tree or
            forest and the number of connected components.
      %\pnum\throws
      \pnum\complexity
            \begin{itemize}
                  \item \textbf{Time:} $\mathcal{O}(|E|\log{|E|})$.
                  \item \textbf{Space:} $\mathcal{O}(|V|)$ auxiliary --- union-find only; no edge copy.
            \end{itemize}
      \pnum\remarks The input edge list is modified (sorted).  If the caller needs to
            preserve the original order, use \lstinline{kruskal} instead.
            The optional \lstinline{alloc} parameter supplies an allocator used for
            the internal disjoint-set (union-find) storage.  Defaults to
            \tcode{std::allocator<std::byte>}.
      %\pnum\errors
\end{itemdescr}

\subsection{Prim Minimum Spanning Tree}
Find the minimum weight spanning tree of a graph using Prim's algorithm.
Delegates to \lstinline{dijkstra_shortest_paths} with a projecting combine function that uses the edge weight directly rather than accumulating path distance.

\begin{table}[ht]
\setcellgapes{3pt}
\makegapedcells
\centering
\begin{tabular}{|P{0.30\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|P{0.20\textwidth}|}
\hline
      \multirowcell{2}{
            \textbf{Complexity} \\
            $\mathcal{O}(|E|log|V|)$
            }
      & \textbf{Directed?} Yes & \textbf{Cycles?} No & \textbf{Throws?} Yes \\
      & \textbf{Multi-edge?} No & \textbf{Self-loops} Yes & \\
\hline
\end{tabular}
%\caption{Algorithm Example}
\end{table}

{\small
      \lstinputlisting[firstline=28,lastline=46]{D3128_Algorithms/src/mst.hpp}
}

\begin{itemdescr}
      \pnum\mandates
            \begin{itemize}
                  \item \tcode{G} satisfies \tcode{adjacency_list<G>}.
                  \item \tcode{WeightFn} satisfies \tcode{distance_fn_for<WeightFn, G>}
                        with \tcode{is_arithmetic_v<distance_fn_value_t<WeightFn, G>>}.
                  \item \tcode{PredecessorFn} satisfies \tcode{predecessor_fn_for<PredecessorFn, G>}.
                  \item \tcode{WF} satisfies
                        \tcode{basic_edge_weight_function<G, WF, distance_fn_value_t<WeightFn, G>, CompareOp,}
                        \tcode{plus<distance_fn_value_t<WeightFn, G>>>}.
                  \item \lstinline{compare} is a valid strict weak ordering on values of type \lstinline{distance_fn_value_t<WeightFn, G>}.
            \end{itemize}
      \pnum\preconditions
            \begin{itemize}
                  \item
                        When \lstinline{weight_fn} is omitted, the default is \lstinline{edge_value(g, uv)}.
            \end{itemize}
      \pnum\hardprecond
            \begin{itemize}
                  \item
                        \lstinline{seed} is a valid vertex ID in \lstinline{g}.
                  \item
                        \lstinline{weight(g, v)} and \lstinline{predecessor(g, v)} have been initialized for
                        each vertex \lstinline{v} in \lstinline{g} before calling
                        (e.g.~via \lstinline{init_shortest_paths(g, weight_cont, pred_cont)} on the
                        underlying containers, then wrapping with \lstinline{container_value_fn}).
            \end{itemize}
      \pnum\effects
            \begin{itemize}
                  \item
                        \lstinline{predecessor(g, v)} is the parent vertex of \lstinline{v} in a tree rooted at \lstinline{seed}
                        and \lstinline{weight(g, v)} is the value of the edge between \lstinline{v} and \lstinline{predecessor(g, v)}
                        in the tree.  \lstinline{predecessor(g, seed) == seed}.
                        When \lstinline{compare} is \lstinline{<}, \lstinline{predecessor} represents a minimum weight spanning tree.
                  \item
                        For vertices not reachable from \lstinline{seed}, \lstinline{predecessor(g, v)} and \lstinline{weight(g, v)}
                        retain their initialized values.
            \end{itemize}
      %\pnum\result
      \pnum\returns The total weight of the minimum spanning tree rooted at \lstinline{seed}.
      \pnum\throws
            \begin{itemize}
                  \item \lstinline{std::out_of_range} if \lstinline{seed} is not a valid vertex ID in \lstinline{g}.
                  \item \lstinline{std::out_of_range} if \lstinline{predecessor} or \lstinline{weight} are undersized for the graph.
                  \item \lstinline{std::out_of_range} if a negative edge weight is encountered (for signed weight types).
                  \item \lstinline{std::logic_error} if an internal invariant violation is detected.
            \end{itemize}
      \pnum\complexity
            \begin{itemize}
                  \item \textbf{Time:} $\mathcal{O}(|E|\log{|V|})$.
                  \item \textbf{Space:} $\mathcal{O}(|V|)$ auxiliary --- priority queue and bookkeeping arrays.
            \end{itemize}
      \pnum\remarks Only produces a spanning tree for the connected component containing \lstinline{seed}.
            For disconnected graphs, call \lstinline{prim} once per component with a seed vertex from each component.
            The optional \lstinline{alloc} parameter supplies an allocator used for
            the internal priority queue (forwarded to \lstinline{dijkstra_shortest_paths}).
            Defaults to \tcode{std::allocator<std::byte>}.
      %\pnum\errors
\end{itemdescr}

