\chapter*{Revision History}
\section*{P1709R5}
Extensions and refinements to r4
\begin{itemize}
\item Added basic\_* versions for depth first search, breadth first search and topoloical sort views. Also renamed the views to use bfs and dfs to avoid long names.
\item Replace \tcode{adjacency_list} with \tcode{index_adjacency_list} in appropriate algorithms.
\item Refined algorithm defintions.
\end{itemize}

\section*{P1709R4}
This was a major redesign that incorporated all the experience and input from the past four years.
\begin{itemize}
\item Revisit the algorithms to be considered.
\item Reduce the scope to focus on an adjacency list with outgoing edges, edge list, and remove mutable interface functions. 
\item Replace directed and undirected concepts with overridable types of unordered\_edge for a graph type.
\item Simplify the Graph Container types and functions. In particular, const and non-const variations were consolidated to a single definition to handle both cases 
when appropriate.
\item All Graph Container Interface functions are customization points.
\item Introduce Views, inspired by NWGraph design, resulting in simpler and cleaner interfaces to traverse a graph, and simplifying the container interface design.
\item Add support for bipartite and multipartite graphs.
\item Replace the two container implementations with compressed\_graph, based on the Compressed Sparse Row matrix, a commonly used 
      data structure for high-performance graphs.
\end{itemize}

\section*{P1709R3}
A simple status revision to say a major change is coming soon.

\section*{P1709R2}
Define the \textbf{uniform API} for undirected and directed algorithms (an extended API also exists for directed graphs). Added \textbf{concepts} for undirected, directed and bidirected graphs. Refined \textbf{DFS} and \textbf{BFS} range definitions from prototype experience. Refined \textbf{shortest paths} and \textbf{transitive closure} algorithms from input and prototype experience.

\section*{P1709R1}
Rewrite with a focus on a \textbf{purely functional design}, emphasizing the algorithms and graph API. Also added \textbf{concepts} and \textbf{ranges} into the design. Addressed concerns from Cologne review to change to functional design.

\section*{P1709R0}
Focus on \textbf{object-oriented API} for data structures and example code for a few algorithms.
