
\clearpage
%% \chapter{Graph Container Implementations}

\section{compressed\_graph Graph Container}
\tcode{compressed_graph} is a graph container being proposed for the standard library. It is a high-performance data structure that 
uses \href{https://en.wikipedia.org/wiki/Sparse_matrix#Compressed_sparse_row_\%28CSR\%2C_CRS_or_Yale_format\%29}{Compressed Sparse Row} 
(CSR) format to store its vertices, edges and associated values. Once constructed, vertices and edges cannot be added or deleted 
but values on vertices and edges can be modified.

There are a number of features added beyond the typical CSR implementation:
\begin{itemize}
    \item   \textbf{User-defined values} The typical CSR implementation stores values on edges (columns) by defining the \tcode{EV}
            template parameter. \tcode{compressed_graph} extends that to also allow values on vertices (rows) and the graph itself 
            by defining the \tcode{VV} and \tcode{GV} template arguments respectively. If a type is void, no memory overhead is incurred.
    \item   \textbf{Index type sizes} The size of the integral indexes into the internal vertex (row) and edge (column) structures can be 
            controlled by the \tcode{VId} and \tcode{EIndex} template arguments respectively to give a balance between capacity, memory 
            usage and performance.
    \item   \textbf{Multi-partite graphs} The vertices can optionally be partitioned into multiple partitions by passing the 
            starting vertex id of each partition in the \tcode{partition_start_ids} argument in the constructors. If no partitions are
            specified, the graph is single-partite.
\end{itemize}

The listings in the following sections show the prototypes for the \tcode{compressed_graph} when the graph value type \tcode{GV} is 
non-\tcode{void} (section \ref{compressed_full}) and a class template specialization when it is \tcode{void} (section \ref{compressed_specialization}).

Only the constructors and destructor shown for \tcode{compressed_graph} are public. All other types and functions related to the graph
are only accessible through the types and functions in the Graph Container Interface.

\begin{table}[h]
    \setcellgapes{3pt}
    \makegapedcells
    \centering
    \begin{tabular}{|P{0.38\textwidth}|P{0.35\textwidth}|P{0.25\textwidth}|}
    \hline
    \textbf{vertex\_id assignment:} Contiguous & \textbf{\tcode{has_edges(g)}} $O(1)$ & \textbf{Append vertices?} No \\
    \textbf{Vertices range:} Contiguous & \textbf{\tcode{num_edges(g)}} $O(1)$ & \textbf{Append edges?} No \\
    \textbf{Edge range:} Contiguous & \textbf{\tcode{partition_id(g,u)}} $O(log(P+1))$ & \textbf{Partitions?} Yes\\
    & & \textbf{is\_directed?} No \\
    %O(log(P+1)) is the complexity of std::upper_bound; +1 is for the number of vertices added at the end of partition_start_ids
    \hline
    \end{tabular}
    %\caption{Jaccard Coefficient Summary}
    \label{tab:compressed_graph_summary}
\end{table}
$P$ is the number of partitions and is expected to be small, e.g. $P = 2$ for bipartite and $P \leq 10$ for typical 
multi-partite graphs.

The \tcode{is_directed} trait is not supported. If \tcode{compressed_graph} is intended to be used for an undirected
graph, then the edge pairs must be included for both directions, (uid,vid) and (vid,uid), when constructing the graph.

A public \tcode{operator[](vertex_id_t<G>)} is \emph{not} provided. The only value it could return is an
internal CSR element that is not useful on its own, and exposing it would breach the rule that
\tcode{compressed_graph} is accessed solely through the Graph Container Interface. The equivalent functionality
is already available generically via \tcode{vertex_value(g,uid)} for the vertex value and
\tcode{out_edges(g,*find_vertex(g,uid))} for the incident edges.

\subsection{compressed\_graph when \tcode{GV} is not \tcode{void}} \label{compressed_full}
{\small
      \lstinputlisting{D3131_Containers/src/compressed_graph.hpp}
}

\subsection{compressed\_graph specialization when \tcode{GV} is \tcode{void}} \label{compressed_specialization}
When \tcode{GV} is void the number of constructors decreases significantly as shown in the following listing.

{\small
      \lstinputlisting{D3131_Containers/src/compressed_graph_gvoid.hpp}
}

\subsection{compressed\_graph description}

\tcode{EV} and \tcode{VV} are currently required to be copyable. Supporting move-only \tcode{EV}/\tcode{VV}
is possible future work but is not proposed here, because it would: (1) disable the
\tcode{initializer_list} constructor (an \tcode{initializer_list} exposes only \tcode{const} elements, so its
values can only be copied); (2) restrict construction to the rvalue-range overloads that consume an owning
input range (the \tcode{const}-lvalue-range overloads copy each value); and (3) make the whole
\tcode{compressed_graph} move-only for those instantiations, since its defaulted copy constructor would be
deleted once the internal value vectors are non-copyable. It would also require a corresponding relaxation of
the \tcode{copyable_edge}/\tcode{copyable_vertex} requirements in the Graph Container Interface. A non-\tcode{void}
\tcode{GV} is already permitted to be movable (which likewise makes the graph move-only).

\begin{itemdescr}
    \pnum\mandates
        \begin{itemize}
            \item The \tcode{EV} template argument for an edge value must be a copyable type or \tcode{void}.
            \item The \tcode{VV} template argument for a vertex value must be a copyable type or \tcode{void}.
            \item When the \tcode{GV} template argument for a graph value is not \tcode{void} it can be movable or copyable. It must have a 
                  default constructor if it is not passed in a \tcode{compressed_graph} constructor.
            \item The \tcode{EProj} template argument must be a projection that returns a value of \tcode{copyable_edge_t<VId, EV>} type
                  given a value of \tcode{erng}. If the value type of \tcode{ERng} is already a \tcode{copyable_edge_t<VId, EV>} type, 
                  then \tcode{EProj} can be \tcode{identity}.
            \item The \tcode{VProj} template argument must be a projection that returns a value of \tcode{copyable_vertex_t<VId, VV>} type,
                  given a value of \tcode{vrng}. If the value type of \tcode{Vrng} is already a \tcode{copyable_vertex_t<VId, VV>} type, 
                  then \tcode{VProj} can be \tcode{identity}.
        \end{itemize}
    \pnum\preconditions
        \begin{itemize}
            \item The \tcode{VId} template argument must be able to store a value of |V|+1, where |V| is the
                  number of vertices in the graph. The size of this type impacts the size of the \textit{edges}.
            \item The \tcode{EIndex} template argument must be able to store a value of |E|+1, where |E| is the
                  number of edges in the graph. The size of this type impact the size of the \textit{vertices}.
            \item The \tcode{EProj} and \tcode{VProj} template arguments must be valid projections.
            \item The \tcode{partition_start_ids} range includes the starting vertex id for each partition. If it is empty, then the graph is single-partite
                    and the number of partitions is 1. If it is not empty, then the number of partitions is the size of the range, where the first element
                    must be 0 and all elements are in strictly increasing order. A vertex id in the range must not exceed the number of 
                    vertices in the graph. Any violation of these conditions throws \tcode{graph_error}.
        \end{itemize}

    \pnum\effects
        \begin{itemize}
            \item When \tcode{EV}, \tcode{VV}, or \tcode{GV} are \tcode{void}, no extra memory overhead is incurred for that type.
         \end{itemize}
    %\pnum\result
    %\pnum\returns \lstinline{void} \\
    %\pnum\throws \tcode{out_of_range} is thrown when \tcode{source} is not in the range \tcode{0 <= source < num_vertices(graph)}.  \\
    %\pnum\complexity \\
    \pnum\remarks 
        \begin{itemize}
            \item  The \tcode{VId} and \tcode{EIndex} template arguments impact the capacity, internal storage requirements and performance.
                   The default of \tcode{uint32_t} is sufficient for most graphs and provides a good balance between storage and performance.

                   The memory requirements are roughly, 
                        \[|V| \times (sizeof(EIndex)+sizeof(VV)) + |E| \times (sizeof(VId) + sizeof(EV)) + sizeof(GV)\]
                   where $|V|$ is the number of vertices and $|E|$ is the number of edges in the graph. $sizeof$ \tcode{void} is 0
                   when considering $sizeof$ for \tcode{VV}, \tcode{EV}, and \tcode{GV}. Alignment and overhead for internal vectors are 
                   not included in this calculation.
            \item The allocator passed to constructors is rebound for different types used by different internal containers.
        \end{itemize}
%\pnum\errors
\end{itemdescr}

\section{Using Existing Data Structures}
Reasonable defaults have been defined for the adjacency list and edgelist using data structures and types
in the standard library, with some adaptation for externally defined containers, out of the box that require
no function overrides.

\subsection{Adjacency List Data Structures}

\subsubsection{Using Standard Containers for an Adjacency List}
When the graph is defined using standard containers, the GCI functions can be used without any function overrides.

In this example, we'll use \tcode{G = vector<forward_list<tuple<int,double>>>} to define the graph, where \tcode{g} 
is an instance of \tcode{G}. \tcode{tuple<int,double>} defines the target\_id and weight property respectively. We
can write loops to go through the vertices, and edges within each vertex, as follows.

\begin{lstlisting}
using G = vector<forward_list<tuple<int,double>>>;

G g;
// ... populate g with vertices and weighted edges ...

// A weight accessor following the f(const graph, descriptor) convention.
// edge value is the second tuple element (the double).
auto weight = [](const G& g, edge_t<G> uv) { return edge_value(g, uv); };

// Using GCI functions
for (auto u : vertices(g)) {        // u is a vertex descriptor
  for (auto uv : edges(g, u)) {      // uv is an edge descriptor
    auto vid = target_id(g, uv);     // the target id (an int)
    auto w   = weight(g, uv);        // the weight (a double)
    // do something...
  }
}
\end{lstlisting}

\tcode{edge_t<G>} is defined as \tcode{tuple<int,double>} in the example; the only requirement is that the first element
in the \tcode{tuple} is integral and is used as the target\_id. The \tcode{edge_value(g,uv)} function is automatically
available for a recognized \tcode{tuple} edge and returns the remaining value (the \tcode{double} here); defining it
explicitly is therefore optional, since algorithms typically accept a value lambda to extract the value from the edge.

If all you need is the target\_id without any values, you can use \tcode{G = vector<forward_list<int>>}. The \tcode{int}
is used as the target\_id. Again, the only requirement is that it be \tcode{integral}.

Note that no function override was required. More formally, the two patterns recognized are \tcode{random_access_range<forward_range<integral>>}
and \tcode{random_access_range<forward_range<tuple<integral,...>>>}. This extends to any range type. For instance, boost::containers can be used
just as easily as std containers.

Table \ref{tab:simple_graph} shows how the types are defined for the example above. Because this proposal uses
boost-like descriptors, \tcode{vertices(g)} and \tcode{edges(g,u)} yield \emph{descriptors}; the standard-container
elements (\tcode{forward_list<...>}, \tcode{tuple<int,double>}) are the \emph{underlying} values reached through those
descriptors.
\begin{table}[h!]
    \begin{center}
    \resizebox{\textwidth}{!}
    {\begin{tabular}{l L{9.0cm}}
    \hline
        \textbf{Function or Value} & \textbf{Concrete Type} \\
    \hline
        \tcode{vertices(g)} & vertex range of descriptors over \tcode{vector<forward_list<tuple<int,double>>>} \\
        \tcode{u} & a \tcode{vertex_t<G>} descriptor; underlying value \tcode{forward_list<tuple<int,double>>} \\
        \tcode{edges(g,u)} & edge range of descriptors over the vertex's \tcode{forward_list<tuple<int,double>>} \\
        \tcode{uv} & an \tcode{edge_t<G>} descriptor; underlying value \tcode{tuple<int,double>} \\
        \tcode{target_id(g,uv)} & \tcode{int} --- \tcode{get<0>(uv)} (the integral target\_id) \\
        \tcode{edge_value(g,uv)} & \tcode{double} --- \tcode{get<1>(uv)} (optional; the remaining tuple value) \\
    \hline
    \end{tabular}}
    \caption{Types When Using Standard Containers}
    \label{tab:simple_graph}
    \end{center}
\end{table}

\paragraph{Catalog of recognized standard containers}
The Graph Container Interface proposal specifies the normative patterns that trigger automatic
recognition (the random-access and associative vertex patterns, and the integral/\tcode{pair}/\tcode{tuple}
edge element patterns). This section catalogs the concrete standard containers that match those
patterns and their trade-offs. Detection is concept-based, so non-standard containers (e.g.\
\tcode{boost} containers) that satisfy the same patterns work identically.

Vertex containers in the \textbf{random-access} pattern (\tcode{vertex_id_t<G>} is the integral index,
$\mathcal{O}(1)$ \tcode{find_vertex}):
\begin{itemize}
    \item \tcode{vector<vector<int>>} --- vertices in a \tcode{vector}, edges as \tcode{int} target ids.
    \item \tcode{vector<list<pair<int,double>>>} --- vertices in a \tcode{vector}, edges as weighted pairs in a \tcode{list}.
    \item \tcode{deque<forward_list<int>>} --- vertices in a \tcode{deque}, edges in a \tcode{forward_list}.
\end{itemize}

Vertex containers in the \textbf{associative} pattern (\tcode{vertex_id_t<G>} is the \tcode{key_type},
\tcode{find_vertex} via \tcode{g.find(uid)}):
\begin{itemize}
    \item \tcode{map<int, vector<int>>} --- sparse vertex ids in a sorted \tcode{map}, edges as target ids.
    \item \tcode{unordered_map<int, vector<pair<int,double>>>} --- hash-based vertex lookup, weighted edges.
\end{itemize}

The associative pattern uses the same GCI functions as the random-access pattern, again with no overrides.
The difference is that the vertex id is the map \emph{key} rather than a dense index, so it may be sparse (or a
non-integral type such as \tcode{string}). For \tcode{G = map<int, vector<pair<int,double>>>} the inner
\tcode{vector<pair<int,double>>} is the edge range reached through \tcode{edges(g,u)}, the \tcode{pair.first} is
the target\_id and the \tcode{pair.second} is the edge value.

\begin{lstlisting}
using G = map<int, vector<pair<int,double>>>;  // sparse vertex ids, weighted edges

G g;
// ... populate g, keyed by vertex id ...

// Same accessor convention as the random-access example.
auto weight = [](const G& g, edge_t<G> uv) { return edge_value(g, uv); };

// Using GCI functions -- identical loop shape to the vector-based example.
for (auto u : vertices(g)) {          // u is a vertex descriptor
  auto uid = vertex_id(g, u);          // the map key (a sparse int)
  for (auto uv : edges(g, u)) {        // uv is an edge descriptor over the mapped value
    auto vid = target_id(g, uv);       // the target id (pair.first)
    auto w   = weight(g, uv);          // the weight (pair.second)
    // do something...
  }
}
\end{lstlisting}

Such a graph models \tcode{adjacency_list} and (because the vertex id is hashable) \tcode{mapped_adjacency_list},
but \emph{not} \tcode{index_adjacency_list}: \tcode{find_vertex(g,uid)} is $\mathcal{O}(\log |V|)$ for \tcode{map}
or amortized $\mathcal{O}(1)$ for \tcode{unordered_map} rather than $\mathcal{O}(1)$ indexed access. Algorithms in
this proposal that require \tcode{index_adjacency_list} therefore need an index-based container; the associative
form is intended for adapting existing keyed data in place, as described in the Graph Container Interface proposal.

The inner edge range returned by \tcode{edges(g,u)} may be any \tcode{forward_range} whose element matches a
recognized edge element pattern. Table~\ref{tab:edge_container_patterns} groups the standard containers by
category and summarises their trade-offs as edge containers.

\begin{table}[h!]
\begin{center}
\resizebox{\textwidth}{!}
{\begin{tabular}{l l L{7.5cm}}
\hline
    \textbf{Category} & \textbf{Container} & \textbf{Characteristics} \\
\hline
    Sequence &
    \tcode{vector<E>} &
    Random-access, cache-friendly, amortized $\mathcal{O}(1)$ push-back. \\
    &
    \tcode{deque<E>} &
    Random-access, stable references on push-back. \\
    &
    \tcode{list<E>} &
    Bidirectional, $\mathcal{O}(1)$ splice, stable iterators. \\
    &
    \tcode{forward_list<E>} &
    Forward-only, minimal overhead, $\mathcal{O}(1)$ push-front. \\
\hdashline
    Ordered associative &
    \tcode{set<E>} &
    Sorted, automatic deduplication, $\mathcal{O}(\log n)$ lookup; element is the edge itself. \\
    &
    \tcode{map<VId, EV>} &
    Sorted by target id, $\mathcal{O}(\log n)$ lookup, no parallel edges; \tcode{.first} is
    target id, \tcode{.second} is edge value. \\
\hdashline
    Unordered associative &
    \tcode{unordered_set<E>} &
    Hash-based deduplication, amortized $\mathcal{O}(1)$ insert/lookup. \\
    &
    \tcode{unordered_map<VId, EV>} &
    Hash-based by target id, amortized $\mathcal{O}(1)$ lookup, no parallel edges;
    \tcode{.first} is target id, \tcode{.second} is edge value. \\
\hline
\end{tabular}}
\caption{Recognized Edge Container Patterns}
\label{tab:edge_container_patterns}
\end{center}
\end{table}

These vertex and edge containers combine freely: any recognized vertex container can hold any recognized
edge container, and any recognized edge container can hold any recognized edge element type. For example,
\tcode{vector<vector<pair<int,double>>>} is a random-access graph with weighted edges,
\tcode{vector<set<int>>} is a random-access graph with sorted deduplicated unweighted edges,
\tcode{vector<map<int,double>>} is a random-access graph with $\mathcal{O}(\log n)$ edge lookup by target,
and \tcode{map<string, list<int>>} is an associative graph with unweighted edges.
When the element type does not match a recognized pattern, the user must override the appropriate CPOs for
the graph type, as described next.

\subsubsection{Using Other Graph Data Structures}
For other graph data structures function overrides are required. Table \ref{tab:cmn_cpo_overrides} shows the 
common function overrides anticipated for most cases, keeping in mind that all functions can be overridden
if the default implementation is not suitable. 
When they are defined they must be in the same namespace as the data structures.

\begin{table}[h!]
    \begin{center}
    %\resizebox{\textwidth}{!}
    {\begin{tabular}{l l p{1.5cm} L{7.0cm}}
    \hline
        \textbf{Function} & \textbf{Comment} \\
    \hline
        \tcode{vertices(g)} & \\
        \tcode{edges(g,u)} & \\
        \tcode{target_id(g,uv)} & \\
        \tcode{edge_value(g,uv)} & If edges have value(s) in the graph \\
        \tcode{vertex_value(g,u)} & If vertices have value(s) in the graph \\
        \tcode{graph_value(g)} & If the graph has value(s) \\
    \hline
        \multicolumn{2}{c}{When edges have the optional source\_id on an edge} \\
    \hdashline
        \tcode{source_id(g,uv)} & \\
    \hline
        \multicolumn{2}{c}{When the graph supports multiple partitions} \\
    \hdashline
        \tcode{num_partitions(g)} & \\
        \tcode{partition_id(g,u)} & \\
        \tcode{vertices(g,pid)} & \\
    \hline
    \end{tabular}}
    \caption{Common CPO Function Overrides}
    \label{tab:cmn_cpo_overrides}
    \end{center}
\end{table}

\subsection{Edgelist Data Structures}
\subsubsection{Using Standard Containers for an Edgelist}
Like the adjacency list, an edgelist can be defined using standard containers and types without 
requiring any function overrides.

\begin{lstlisting}
    using EL = vector<tuple<int, int, double>>;
    using E  = std::range_value_t<EL>;
    EL el{{1, 2, 11.1}, {1, 4, 22.2}, {2, 3, 3.33}, {2, 4, 4.44}};
    for (auto&& e : el) {
      int    sid = source_id(el, e);
      int    tid = target_id(el, e);
      double val = edge_value(el, e);
    }
\end{lstlisting}

An alternative is to use the \tcode{edge_data} used in this proposal. Notice that the only difference
is the definition of the edgelist type. All other code is identical to the previous example.
\begin{lstlisting}
    using EL = vector<edge_data<int, true, void, double>>;
    using E  = std::range_value_t<EL>;
    EL el{{1, 2, 11.1}, {1, 4, 22.2}, {2, 3, 3.33}, {2, 4, 4.44}};
    for (auto&& e : el) {
      int    sid = source_id(el, e);
      int    tid = target_id(el, e);
      double val = edge_value(el, e);
    }
\end{lstlisting}

While these examples show the optional edge\_value that is a \tcode{double}, it can be omitted if the edges 
do not have values.

Type alias are in the namespace \tcode{std::graph::edge_list} to avoid conflicts with adjacency\_list
types.

\subsubsection{Using Other Graph Data Structures}
If you have different edge type not covered by the standard types, you can override the
\tcode{source_id(e)}, \tcode{target_id(e)} and \tcode{edge_value(E)} functions for that type.
The functions must be in the same namespace as the edge data structure you want to use.
