\section{Revision History}

\subsection*{\paperno r0}

\begin{itemize}
      \item Split from P1709r5. Added \textit{Getting Started} section.
      \item Move text for graph data structures created from std containers from Graph Container Interface to 
            Container Implementation paper.
      \item GCI overloads are no longer required for adjacency lists constructed with standard containers. Data structures that follow
            the pattern \tcode{random_access_range<forward_range<integral>>} and \tcode{random_access_range<forward_range<tuple<integral,...>>>}
            are automatically recognized as an adjacency list, including containers from non-standard libraries. The \tcode{integral}
            value is used as the target\_id.
\end{itemize}

\subsection*{\paperno r1}
\begin{itemize}
      \item Added feature summary of \tcode{compressed_graph} beyond the typical CSR implementation.
      \item Added complexity for \tcode{num_edges(g)} and \tcode{has_edge(g)} functions in \tcode{compressed_graph}.
      \item Add constructors to \tcode{compressed_graph} to complement the removal of the load functions from 
            \href{https://www.wg21.link/P3130r1}{P3130r1 Graph Container Interface}. An optional \tcode{partition_start_ids} 
            parameter is also included.
\end{itemize}

\subsection*{\paperno r2}

\begin{itemize}
      \item Add the edgelist as an abstract data structure as a peer to the adjacency list. 
            A section on edgelists has been added to Using Existing Data Structures.
      \item Add \tcode{is_directed} item in the feature summary box to \tcode{compressed_graph}.
\end{itemize}

\subsection*{\paperno r3}

\begin{itemize}
      \item Changes related to boost::graph-like descriptors as described in \href{https://www.wg21.link/P3130r3}{P3130r3 Graph Container Interface}.
            Related changes that affect this paper include:
            \begin{itemize}
                  \item Rename \tcode{descriptor} structs to \tcode{info} structs for new boost::graph-like descriptors.
                  \item Retain \tcode{copyable_edge} and \tcode{copyable_vertex} in the \tcode{compressed_graph}
                        \tcode{requires} clauses. These constrain the \emph{result} of the \tcode{EProj}/\tcode{VProj}
                        projection to be convertible to \tcode{copyable_edge_t<VId,EV>} / \tcode{copyable_vertex_t<VId,VV>};
                        they validate the projected value's type and are unrelated to descriptor copyability.
            \end{itemize}
      
\end{itemize}

\subsection*{\paperno r4}
\begin{itemize}
      \item Received the std-container catalog from the Graph Container Interface proposal: the concrete
            container example lists, the edge-container trade-off table, and the combined examples now live
            here; the normative recognition patterns remain in the GCI paper, which this paper references.
      \item Consistency fixes against the reference implementation: \tcode{edge_info} renamed to
            \tcode{edge_data}; edgelist CPOs shown in 2-argument form \tcode{source_id(el,e)} /
            \tcode{target_id(el,e)} / \tcode{edge_value(el,e)}; edgelist namespace corrected to
            \tcode{std::graph::edge_list}; \tcode{has_edge(g)} corrected to \tcode{has_edges(g)};
            \tcode{partition_id(g,u)} (descriptor form); \tcode{vertices(g,pid)} override; and the
            \tcode{copyable_edge_t<VId,EV>} / \tcode{copyable_vertex_t<VId,VV>} projection result types.
      \item Corrected the \tcode{compressed_graph} partition preconditions: invalid
            \tcode{partition_start_ids} (first element not 0, not strictly increasing, or exceeding the
            vertex count) throw \tcode{graph_error} rather than being undefined behavior.
      \item Updated the standard-container adjacency example and its type table to the descriptor model.
\end{itemize}
