
\clearpage
%% \chapter{Graph Container Interface}
\section{Graph Container Interface}

The Graph Container Interface (GCI) defines the primitive concepts, traits, types and functions used to define and access adjacency lists (aka graphs)
and edgelists, no matter their internal design and organization. For instance, an adjacency list can be a vector of lists from standard containers, 
CSR-based graph and adjacency matrix. Likewise, an edgelist can be a range of edges from a standard container or externally defined edge types, 
provided they have a source\_id, target\_id and optional edge\_value.

The GCI covers two peer abstract data types, each with its own namespace:
\begin{itemize}
    \item \tcode{std::graph::adj\_list} --- concepts, type aliases, CPOs and descriptors for adjacency lists.
    \item \tcode{std::graph::edge\_list} --- concepts, type aliases and CPOs for edge lists.
\end{itemize}
All adjacency list symbols are additionally re-exported into \tcode{std::graph} for convenience. Edge list
symbols are \emph{not} re-exported because several names (notably \tcode{vertex\_id\_t}) have different
definitions in each namespace and cannot coexist at the root level.

If there is a desire to use the algorithms against externally defined data structures, the GCI exposes its functions as customization points
to be overridden as needed. Likewise, externally defined algorithms can be used to operate on other data structures that meet the GCI requirements.
This achieves the same goals as the STL, where algorithms can be used on any container that meets the requirements of the algorithm.

The GCI is designed to support a wider scope of graph containers than required by the views and algorithms in this 
proposal. This enables for future growth of the graph data model (e.g. incoming edges on a vertex), or as a framework for graph implementations 
outside of the standard. For instance, existing implementations may have requirements that cause them to define features with looser constraints, 
such as sparse vertex\_ids, non-integral vertex\_ids, or storing vertices in associative bi-directional containers (e.g. std::map or std::unordered\_map). 

Such features require specialized implementations for views and algorithms. The performance for such algorithms will be sub-optimal, but may be
preferable to run them on the existing container rather than loading the graph into a high-performance graph container and then running the 
algorithm on it, where the loading time can far outweigh the time to run the sub-optimal algorithm. To achieve this, care has been taken to 
make sure that the use of concepts chosen is appropriate for algorithm, view and container.

All algorithms in this and related proposals require that adjacency list vertices are stored in random access containers and that \tcode{vertex_id_t<G>} 
is integral. Future designs may relax these requirements, but for now they are required.

\section{Adjacency List Interface}

\subsection{Namespace}
All adjacency list concepts, type aliases, CPOs and descriptors are defined in the \tcode{std::graph::adj\_list}
namespace. Every symbol is also re-exported into \tcode{std::graph} via \tcode{using} declarations so that
callers who include \tcode{<graph>} can write \tcode{graph::vertices(g)} rather than
\tcode{graph::adj\_list::vertices(g)}. Code that wishes to be explicit about the choice of abstract data
type, or that uses both an adjacency list and an edge list simultaneously, may qualify names with the
\tcode{graph::adj\_list::} prefix.

\subsection{Concepts}

This section describes the concepts used for adjacency lists in the Graph Library. Three qualifiers are used in concept names.
\begin{itemize}
    \item \textbf{index} — the vertex range is random-access and the vertex id is an integral type.
    \item \textbf{mapped} — vertices are stored in an associative container (e.g.\tcode{map} or \tcode{unordered_map}); vertex lookup uses \tcode{find_vertex(g, uid)}.
    \item \textbf{bidirectional} — the graph exposes in-edges via \tcode{in_edges(g, u)} in addition to out-edges.
\end{itemize}

\emph{While we believe the use of concepts is appropriate for graphs as a range-of-ranges, we are marking them as "For exposition only" 
until we have consensus of whether they belong in the standard or not.}

\subsubsection{Edge Concepts}
The \tcode{edge<G,E>} concept unifies all edge types in a single concept: an edge must provide \tcode{source_id},
\tcode{source}, \tcode{target_id}, and \tcode{target} access given the graph.
{\small
     \lstinputlisting{D3130_Container_Interface/src/concepts_edges.hpp}
}

The previous version of this interface defined separate concepts for targeted edges (with only \tcode{target_id}) and sourced edges
(with both \tcode{source_id} and \tcode{target_id}). The descriptor design makes this distinction
unnecessary: every edge descriptor carries enough information to provide both ids, so the single \tcode{edge<G,E>} concept is sufficient.

Return types are not validated in order to provide flexibility. Constraining the return type of \tcode{target_id(g,uv)} to \tcode{integral}
would exclude graphs whose vertex id is a non-integral key (e.g.\ a \tcode{string} for a \tcode{map}-based graph). Constraining to
\tcode{vertex_id_t<G>} gives obscure error messages on failure. Leaving the return type unconstrained allows the compiler to
report the actual type mismatch at the point of use, which is easier to diagnose. There is precedent for this choice in the
\href{https://en.cppreference.com/w/cpp/ranges/sized_range}{\tcode{sized_range}} concept.

\paragraph{Edge Range Concepts}

\tcode{out_edge_range<R,G>} constrains the range returned by \tcode{out_edges(g,u)}, and \tcode{in_edge_range<R,G>} constrains
the range returned by \tcode{in_edges(g,u)} for bidirectional graphs. Both require a \tcode{forward_range} whose value type
satisfies \tcode{edge<G, range_value_t<R>>}.
{\small
     \lstinputlisting{D3130_Container_Interface/src/concepts_target_edge_range.hpp}
}

\subsubsection{Vertex Concepts}
The \tcode{vertex<G,V>} concept requires that a vertex supports \tcode{vertex_id(g,u)} and \tcode{find_vertex(g,uid)}.
Building on it, \tcode{vertex_range<R,G>} requires a \tcode{forward_range} and \tcode{sized_range} whose
value type satisfies \tcode{vertex<G,...>}.

Two specializations of vertex range are provided:
\begin{itemize}
    \item \tcode{index_vertex_range<G>} — vertex id is integral and the vertex range is random-access with an integral \tcode{storage_type}. Used for high-performance graphs such as \tcode{index_adjacency_list}.
    \item \tcode{mapped_vertex_range<G>} — vertex id is non-integral and vertices are found via \tcode{find_vertex(g, uid)}. Used for graphs whose vertices are stored in associative containers.
\end{itemize}
{\small
     \lstinputlisting{D3130_Container_Interface/src/concepts_vertex_range.hpp}
}

\subsubsection{Adjacency List Concepts}
The adjacency list concepts combine the vertex and edge concepts. Six concepts are defined, covering three axes:
\begin{itemize}
    \item \tcode{adjacency_list<G>} — base concept: \tcode{vertices(g)} returns a \tcode{vertex_range<G>} and \tcode{out_edges(g,u)} returns an \tcode{out_edge_range<G>}.
    \item \tcode{index_adjacency_list<G>} — \tcode{adjacency_list} with \tcode{index_vertex_range}. All algorithms currently proposed for the Graph Library use this concept.
    \item \tcode{bidirectional_adjacency_list<G>} — \tcode{adjacency_list} that also exposes \tcode{in_edges(g,u)} as an \tcode{in_edge_range<G>} with accessible \tcode{source_id}.
    \item \tcode{index_bidirectional_adjacency_list<G>} — \tcode{bidirectional_adjacency_list} with \tcode{index_vertex_range}.
    \item \tcode{mapped_adjacency_list<G>} — \tcode{adjacency_list} with \tcode{mapped_vertex_range}.
    \item \tcode{mapped_bidirectional_adjacency_list<G>} — \tcode{bidirectional_adjacency_list} with \tcode{mapped_vertex_range}.
\end{itemize}

The previous interface defined \tcode{basic_*} and \tcode{sourced_*} variants to distinguish edges with or without a source id.
The descriptor design makes this distinction unnecessary: all edge descriptors carry source information, so the \tcode{basic_*}
and \tcode{sourced_*} concept families have been removed.
{\small
     \lstinputlisting{D3130_Container_Interface/src/concepts_adj_list.hpp}
}

\subsection{Traits}
Table \ref{tab:graph_traits} summarizes the type traits in the Graph Container Interface, allowing views and algorithms to query the graph's characteristics.

\begin{table}[h!]
\begin{center}
%\resizebox{\textwidth}{!}
{\begin{tabular}{l l L{7.0cm}}
\hline
    \textbf{Trait} & \textbf{Type} & \textbf{Comment} \\
\hline
    \tcode{has_degree<G>} & concept & Is the \tcode{out_degree(g,u)} function available? \\
    \tcode{has_find_vertex<G>} & concept & Are the \tcode{find_vertex(g,\_)} functions available? \\
    \tcode{has_find_vertex_edge<G>} & concept & Are the \tcode{find_out_edge(g,\_)} functions available?\\
    \tcode{has_contains_edge<G,V>} & concept & Is the \tcode{contains_out_edge(g,uid,vid)} function available?\\
\hdashline
    \tcode{has_in_degree<G>} & concept & Is the \tcode{in_degree(g,u)} function available? \\
    \tcode{has_find_in_edge<G>} & concept & Are the \tcode{find_in_edge(g,\_)} functions available? \\
    \tcode{has_contains_in_edge<G,V>} & concept & Is the \tcode{contains_in_edge(g,uid,vid)} function available?\\
\hdashline
    \tcode{has_basic_queries<G>} & concept & \tcode{has_degree<G> \&\& has_find_vertex<G> \&\& has_find_vertex_edge<G>} \\
    \tcode{has_full_queries<G>} & concept & \tcode{has_basic_queries<G> \&\& has_contains_edge<G, vertex_t<G>>} \\
\hline
    \tcode{define_adjacency_matrix<G> : false_type} & struct & Specialize for graph implementation to derive from \tcode{true_type} for edges stored as a square 2-dimensional array. Not yet in reference implementation. \\
    \tcode{is_adjacency_matrix<G>} & struct & \\
    \tcode{is_adjacency_matrix_v<G>} & type alias & \\
    \tcode{adjacency_matrix<G>} & concept & \\
\hline
\end{tabular}}
\caption{Graph Container Interface Type Traits}
\label{tab:graph_traits}
\end{center}
\end{table}


\subsection{Types}
Table \ref{tab:graph_type} summarizes the type aliases in the Graph Container Interface. These are the types used to define the objects 
in a graph container, no matter its internal design and organization. Thus, it is designed to be able to reflect all forms of adjacency 
graphs including a vector of lists, compressed\_graph and adjacency matrix.

The type aliases are defined by either a function specialization for the underlying graph container, or a refinement of one of those types (e.g. an iterator of a range). Table \ref{tab:graph_func} describes the functions in more detail.

\tcode{graph_value(g)}, \tcode{vertex_value(g,u)} and \tcode{edge_value(g,uv)} can be optionally implemented, depending on whether the graph container supports values on the graph, vertex and edge types.

\begin{table}[h!]
\begin{center}
%\resizebox{\textwidth}{!}
{\begin{tabular}{l l L{2.5cm}}
\hline
    \textbf{Type Alias} & \textbf{Definition} & \textbf{Comment} \\
\hline
    \tcode{graph_value_t<G>} & \tcode{decltype(graph_value(g))} & optional \\
\hline
    \tcode{vertex_range_t<G>} & \tcode{decltype(vertices(g))} & \\    
    \tcode{vertex_iterator_t<G>} & \tcode{iterator_t<vertex_range_t<G>>} & \\    
    \tcode{vertex_t<G>} & \tcode{range_value_t<vertex_range_t<G>>} & descriptor \\    
    \tcode{vertex_id_t<G>} & \tcode{remove_cvref_t<decltype(vertex_id(g,vertex_t<G>))>} & \\
    \tcode{vertex_id_store_t<G>} & (see below) & normative \\
    \textit{\tcode{raw-vertex-id-type}} & \tcode{decltype(vertex_id(g,vertex_t<G>))} & exposition only \\
    \tcode{vertex_value_t<G>} & \tcode{decltype(vertex_value(g,u))} & optional \\
\hline
    \tcode{out_edge_range_t<G>} & \tcode{decltype(out_edges(g,vertex_t<G>))} & \\
    \tcode{out_edge_iterator_t<G>} & \tcode{iterator_t<out_edge_range_t<G>>} & \\
    \tcode{out_edge_t<G>} & \tcode{range_value_t<out_edge_range_t<G>>} & descriptor \\
    \tcode{vertex_edge_range_t<G>} & \tcode{out_edge_range_t<G>} & alias \\
    \tcode{vertex_edge_iterator_t<G>} & \tcode{out_edge_iterator_t<G>} & alias \\
    \tcode{edge_t<G>} & \tcode{out_edge_t<G>} & alias \\
    \tcode{edge_value_t<G>} & \tcode{decltype(edge_value(g,uv))} & optional \\
\hdashline
    \tcode{in_edge_range_t<G>} & \tcode{decltype(in_edges(g,vertex_t<G>))} & \\
    \tcode{in_edge_iterator_t<G>} & \tcode{iterator_t<in_edge_range_t<G>>} & \\
    \tcode{in_edge_t<G>} & \tcode{range_value_t<in_edge_range_t<G>>} & descriptor \\
\hline
    \tcode{partition_id_t<G>} & \tcode{decltype(partition_id(g,u))} & optional \\
    \tcode{partition_vertex_range_t<G>} & \tcode{vertices(g,pid)} & optional \\    
\hline
\end{tabular}}
\caption{Graph Container Interface Type Aliases}
\label{tab:graph_type}
\end{center}
\end{table}

There is no contiguous requirement for \tcode{vertex_id} from one partition to the next, though
in practice they will often be assigned contiguously. Gaps in \tcode{vertex_id}s between partitions should be allowed.

\subsection{Classes and Structs}

The \tcode{graph_error} exception class is available, inherited from \tcode{runtime_error}. While any function may use it, it 
is only anticipated to be used by the \tcode{load} functions at this time.
No additional functionality is added beyond that provided by \tcode{runtime_error}.

\emph{While we believe the use of concepts is appropriate for graphs as a range-of-ranges, we are marking them as "For exposition only" 
until we have consensus of whether they belong in the standard or not.}

\subsubsection{Data Structs}

\tcode{vertex_data}, \tcode{edge_data} and \tcode{neighbor_data} are aggregates used by
views for structured-binding iteration. They are defined with multiple specializations
that void-out unused members so that the structured binding is compact.

\tcode{vertex_data<VId, V, VV>} has the following members depending on the template
parameters (a \tcode{void} parameter omits the corresponding member):

\begin{table}[h!]
\begin{center}
{\begin{tabular}{l l L{7.5cm}}
\hline
    \textbf{Members present} & \textbf{Specialization} & \textbf{Typical use} \\
\hline
    \tcode{id, vertex, value} & \tcode{<VId, V, VV>} & Full vertexlist with descriptor and value \\
    \tcode{id, vertex}        & \tcode{<VId, V, void>} & Vertexlist with descriptor, no value \\
    \tcode{id, value}         & \tcode{<VId, void, VV>} & Vertexlist with value, no descriptor \\
    \tcode{id}                & \tcode{<VId, void, void>} & ID-only vertexlist \\
    \tcode{vertex, value}     & \tcode{<void, V, VV>} & Descriptor-based, no id \\
    \tcode{vertex}            & \tcode{<void, V, void>} & Descriptor only \\
    \tcode{value}             & \tcode{<void, void, VV>} & Value only \\
    (empty)                   & \tcode{<void, void, void>} & No members \\
\hline
\hline
\end{tabular}}
\caption{\tcode{vertex\_data} Specializations}
\label{tab:vertex_data}
\end{center}
\end{table}

\tcode{edge_data<VId, Sourced, E, EV>} has the following members. The \tcode{Sourced}
boolean controls whether \tcode{source_id} is included. A \tcode{void} VId or EV omits
the corresponding member; a \tcode{void} E omits the edge descriptor member.

\begin{table}[h!]
\begin{center}
{\begin{tabular}{l l L{6.5cm}}
\hline
    \textbf{Members present} & \textbf{Specialization} & \textbf{Typical use} \\
\hline
    \tcode{source\_id, target\_id, edge, value} & \tcode{<VId,true,E,EV>} & Sourced incidence with descriptor and value \\
    \tcode{source\_id, target\_id, edge}        & \tcode{<VId,true,E,void>} & Sourced incidence with descriptor \\
    \tcode{source\_id, target\_id, value}       & \tcode{<VId,true,void,EV>} & Sourced incidence with value \\
    \tcode{source\_id, target\_id}              & \tcode{<VId,true,void,void>} & Sourced incidence, IDs only \\
    \tcode{target\_id, edge, value}             & \tcode{<VId,false,E,EV>} & Unsourced incidence with descriptor and value \\
    \tcode{target\_id, edge}                    & \tcode{<VId,false,E,void>} & Unsourced incidence with descriptor \\
    \tcode{target\_id, value}                   & \tcode{<VId,false,void,EV>} & Unsourced incidence with value \\
    \tcode{target\_id}                          & \tcode{<VId,false,void,void>} & Target ID only \\
\hline
\hline
\end{tabular}}
\caption{\tcode{edge\_data} Specializations (VId present)}
\label{tab:edge_data}
\end{center}
\end{table}

\tcode{neighbor_data<VId, Sourced, V, VV>} is analogous to \tcode{edge_data} but replaces
the edge descriptor with a target vertex descriptor. It is used by adjacency (neighbor)
views.

The following helper type aliases simplify the most common structured-binding patterns:

\begin{table}[h!]
\begin{center}
{\begin{tabular}{l L{8cm}}
\hline
    \textbf{Alias} & \textbf{Definition} \\
\hline
    \tcode{copyable_vertex_t<VId, VV>}   & \tcode{vertex_data<VId, void, VV>} --- \texttt{\{id, value\}} \\
    \tcode{copyable_edge_t<VId, EV>}     & \tcode{edge_data<VId, true, void, EV>} --- \texttt{\{source\_id, target\_id, value\}} \\
    \tcode{copyable_neighbor_t<VId, VV>} & \tcode{neighbor_data<VId, true, void, VV>} --- \texttt{\{source\_id, target\_id, value\}} \\
\hline
\hline
\end{tabular}}
\caption{Helper type aliases for common structured-binding patterns}
\label{tab:copyable_types}
\end{center}
\end{table}

\subsection{Value Function Concepts}

\tcode{vertex_value_function} and \tcode{edge_value_function} define the required
interface for callables passed to views and algorithms to extract per-vertex or per-edge
scalar values (e.g.\ a weight or a label). Both concepts use the two-argument
\tcode{f(const G\&, descriptor)} convention so that stateless lambdas may be captured
into \tcode{std::views} pipelines without holding a reference to the graph.

\begin{lstlisting}
template <class VVF, class Graph, class VertexDescriptor>
concept vertex_value_function =
    invocable<VVF, const Graph&, VertexDescriptor> &&
    (!is_void_v<invoke_result_t<VVF, const Graph&, VertexDescriptor>>);

template <class EVF, class Graph, class EdgeDescriptor>
concept edge_value_function =
    invocable<EVF, const Graph&, EdgeDescriptor> &&
    (!is_void_v<invoke_result_t<EVF, const Graph&, EdgeDescriptor>>);
\end{lstlisting}

\tcode{basic_edge_weight_function} (defined in the algorithms proposal) refines
\tcode{edge_value_function} with the additional requirement that the return type is
arithmetic.

\subsection{Vertex Property Map}

Algorithms that maintain per-vertex state (visited flags, distances, component labels,
etc.) need a container indexed by vertex ID. For index-based graphs a \tcode{vector} is
natural; for mapped (key-based) graphs an \tcode{unordered_map} is required.
\tcode{vertex_property_map<G, T>} abstracts over this difference.

\begin{lstlisting}
template <class G, class T>
using vertex_property_map =
    conditional_t<index_vertex_range<G>,
                  vector<T>,
                  unordered_map<vertex_id_t<G>, T>>;
\end{lstlisting}

Four helper functions complete the interface:

\begin{table}[h!]
\begin{center}
\resizebox{\textwidth}{!}
{\begin{tabular}{l L{9cm}}
\hline
    \textbf{Function} & \textbf{Description} \\
\hline
    \tcode{make_vertex_property_map(g, init)} & Eager factory: creates a map with every vertex pre-populated to \tcode{init}. O(V). \\
    \tcode{make_vertex_property_map<G,T>(g)}  & Lazy factory: creates an empty map with capacity reserved. O(V) for index, O(1) for mapped. \\
    \tcode{vertex_property_map_contains(m, uid)} & Returns \tcode{true} if \tcode{uid} has an entry (\tcode{uid < size(m)} for \tcode{vector}; \tcode{m.contains(uid)} for \tcode{unordered\_map}). \\
    \tcode{vertex_property_map_get(m, uid, dflt)} & Returns the stored value or \tcode{dflt} if absent. Does not insert. \\
\hline
\hline
\end{tabular}}
\caption{Vertex Property Map Functions}
\label{tab:vpm_func}
\end{center}
\end{table}

\subsection{Functions}

Tables \ref{tab:graph_func}, \ref{tab:vertex_func} and \ref{tab:edge_func} summarize the primitive functions in the Graph Container Interface.
used to access an adjacency graph, no matter its internal design and organization. Thus, it is designed 
to be able to reflect all forms of adjacency graphs including a vector of lists, CSR-based graph and adjacency matrix.


\begin{table}[h!]
\begin{center}
\resizebox{\textwidth}{!}
{\begin{tabular}{l l p{2.3cm} L{6.2cm}}
\hline
    \textbf{Function} & \textbf{Return Type} & \textbf{Complexity} & \textbf{Default Implementation} \\
\hline
    \tcode{graph_value(g)} & \tcode{graph_value_t<G>} & constant & n/a, optional \\
    \tcode{vertices(g)} & \tcode{vertex_range_t<G>} & constant & \tcode{g} if \tcode{random_access_range<G>}, n/a otherwise \\
    \tcode{num_vertices(g)} & \tcode{integral} & constant & \tcode{size(vertices(g))} \\
    \tcode{num_edges(g)} & \tcode{integral} & |E| & \tcode{n=0; for(u: vertices(g)) n+=distance(out\_edges(g,u)); return n;} \\
    \tcode{has_edges(g)} & \tcode{bool} & |V| & \tcode{for(u: vertices(g)) if !empty(out\_edges(g,u)) return true; return false;} \\
\hdashline
    \tcode{num_partitions(g)} & \tcode{integral} & constant & 1 \\
    \tcode{vertices(g,pid)} & \tcode{partition_vertex_range_t<G>} & constant & \tcode{vertices(g)} \\
    \tcode{num_vertices(g,pid)} & \tcode{integral} & constant & \tcode{size(vertices(g))}  \\
\hline
\end{tabular}}
\caption{Graph Functions}
\label{tab:graph_func}
\end{center}
\end{table}
The complexity shown above for \tcode{num_edges(g)} and \tcode{has_edges(g)} is for the default implementation.
Specific graph implementations may have better characteristics.

The only vertex function that requires a vertex id \tcode{(uid)} is \tcode{find_vertex(g,uid)}. The other 
functions that use it are convenience functions that imply a call to \tcode{find_vertex(g,uid)} to get the 
vertex descriptor before a call to the overloaded function that takes a descriptor is made.

The complexity shown above for \tcode{vertices(g,pid)} and \tcode{num_vertices(g,pid)} is for the default
implementation. Specific graph implementations may have different characteristics.

\begin{table}[h!]
\begin{center}
\resizebox{\textwidth}{!}
{\begin{tabular}{l l p{1.5cm} L{7.0cm}}
\hline
    \textbf{Function} & \textbf{Return Type} & \textbf{Cmplx} & \textbf{Default Implementation} \\
\hline
    \tcode{find_vertex(g,uid)} & \tcode{vertex_iterator_t<G>} & constant & \tcode{begin(vertices(g)) + uid} \\
                               &                              &          & \hspace{3mm}if \tcode{random_access_range<vertex_range_t<G>>} \\
    \tcode{vertex_id(g,u)} & \tcode{vertex_id_t<G>} & constant & (see Determining the vertex\_id type below) \\
                           &                       &          & Override to define a different \\
                           &                       &          & \tcode{vertex_id_t<G>} type (e.g. int32\_t). \\
    \tcode{vertex_value(g,u)} & \tcode{vertex_value_t<G>} & constant & n/a, optional \\
    \tcode{vertex_value(g,uid)} & \tcode{vertex_value_t<G>} & constant & \tcode{vertex_value(g,*find_vertex(g,uid))}, \\
                                &                           &          & optional \\
    \tcode{out_edges(g,u)} & \tcode{out_edge_range_t<G>} & constant & \tcode{u} if \tcode{forward_range<vertex_t<G>>}, n/a otherwise \\
    \tcode{out_edges(g,uid)} & \tcode{out_edge_range_t<G>} & constant & \tcode{out\_edges(g,*find\_vertex(g,uid))} \\
    \tcode{out_degree(g,u)} & \tcode{integral} & constant & \tcode{size(out\_edges(g,u))} if \tcode{sized_range<out_edge_range_t<G>>} \\
    \tcode{out_degree(g,uid)} & \tcode{integral} & constant & \tcode{size(out\_edges(g,uid))} if \tcode{sized_range<out_edge_range_t<G>>} \\
    (\tcode{edges(g,u), edges(g,uid)} & & & aliases for \tcode{out\_edges}) \\
    (\tcode{degree(g,u), degree(g,uid)} & & & aliases for \tcode{out\_degree}) \\
\hdashline
    \tcode{in_edges(g,u)} & \tcode{in_edge_range_t<G>} & constant & n/a (requires bidirectional graph) \\
    \tcode{in_edges(g,uid)} & \tcode{in_edge_range_t<G>} & constant & \tcode{in\_edges(g,*find\_vertex(g,uid))} \\
    \tcode{in_degree(g,u)} & \tcode{integral} & constant & \tcode{size(in\_edges(g,u))} if \tcode{sized_range<in_edge_range_t<G>>} \\
    \tcode{in_degree(g,uid)} & \tcode{integral} & constant & \tcode{size(in\_edges(g,uid))} if \tcode{sized_range<in_edge_range_t<G>>} \\
\hdashline
    \tcode{partition_id(g,u)} & \tcode{partition_id_t<G>} & constant &  \\
    \tcode{partition_id(g,uid)} & \tcode{partition_id_t<G>} & constant &  \\
\hline
\hline
\end{tabular}}
\caption{Vertex Functions}
\label{tab:vertex_func}
\end{center}
\end{table}

The default implementation for the \tcode{out_degree} and \tcode{in_degree} functions assumes that \tcode{out_edge_range_t<G>} or \tcode{in_edge_range_t<G>} is a sized range
to have constant complexity. If the underlying container has a non-linear \tcode{size(R)} function, the degree functions will
also be non-linear. This is expected to be an uncommon case.

\begin{table}[h!]
\begin{center}
\resizebox{\textwidth}{!}
{\begin{tabular}{l l p{1.3cm} L{6.9cm}}
\hline
    \textbf{Function} & \textbf{Return Type} & \textbf{Cmplx} & \textbf{Default Implementation} \\
\hline
    \tcode{target_id(g,uv)} & \tcode{vertex_id_t<G>} & constant & (see below) \\
    \tcode{target(g,uv)} & \tcode{vertex_t<G>} & constant & \tcode{*(begin(vertices(g)) + target_id(g, uv))} \\
                         &                     &          & if \tcode{random_access_range<vertex_range_t<G>>} \\
                         &                     &          & \hspace{3mm}\tcode{\&\& integral<decltype(target_id(g,uv))>} \\
    \tcode{edge_value(g,uv)} & \tcode{edge_value_t<G>} & constant & \tcode{uv} if \tcode{forward_range<vertex_t<G>>}, \\
                             &                         &          & n/a otherwise, optional \\
    \tcode{find_out_edge(g,u,vid)} & \tcode{out_edge_t<G>} & linear & \tcode{find(out\_edges(g,u),} \\
                                   &                      &        & \tcode{[](uv) \{ return target\_id(g,uv)==vid; \})} \\
    \tcode{find_out_edge(g,uid,vid)} & \tcode{out_edge_t<G>} & linear & \tcode{find\_out\_edge(} \\
                                     &                      &        & \hspace{8mm}\tcode{g, *find\_vertex(g,uid), vid)} \\
    \tcode{contains_out_edge(g,uid,vid)} & \tcode{bool} & constant & \tcode{uid < size(vertices(g))} \\
                                         &              &          & \tcode{\&\& vid < size(vertices(g))} \\
                                         &              &          & \hspace{3mm}if \tcode{is_adjacency_matrix_v<G>}.\\
                                         &              & linear   & \tcode{find\_out\_edge(g,uid,vid)} \\
                                         &              &          & \tcode{!= end(out\_edges(g,uid))} otherwise. \\
    (\tcode{find\_vertex\_edge(g,u,vid)} & & & alias for \tcode{find\_out\_edge}) \\
    (\tcode{find\_vertex\_edge(g,uid,vid)} & & & alias for \tcode{find\_out\_edge}) \\
    (\tcode{contains\_edge(g,uid,vid)} & & & alias for \tcode{contains\_out\_edge}) \\
\hdashline
    \tcode{source_id(g,uv)} & \tcode{vertex_id_t<G>} & constant & mandatory for descriptor-based edges \\
    \tcode{source(g,uv)} & \tcode{vertex_t<G>} & constant & \tcode{*(begin(vertices(g)) + source\_id(g,uv))} \\
                         &                     &          & if \tcode{random_access_range<vertex_range_t<G>>} \\
                         &                     &          & \hspace{3mm}\tcode{\&\& integral<decltype(source\_id(g,uv))>} \\
\hdashline
    \tcode{find_in_edge(g,u,vid)} & \tcode{in_edge_t<G>} & linear & n/a (requires bidirectional graph) \\
    \tcode{find_in_edge(g,uid,vid)} & \tcode{in_edge_t<G>} & linear & \tcode{find\_in\_edge(} \\
                                    &                     &        & \hspace{8mm}\tcode{g, *find\_vertex(g,uid), vid)} \\
    \tcode{contains_in_edge(g,uid,vid)} & \tcode{bool} & linear & \tcode{find\_in\_edge(g,uid,vid)} \\
                                        &              &        & \tcode{!= end(in\_edges(g,uid))} \\
\hline
\end{tabular}}
\caption{Edge Functions}
\label{tab:edge_func}
\end{center}
\end{table}

When the graph matches the pattern \tcode{random_access_range<forward_range<integral>>} or \tcode{random_access_range<forward_range<tuple<integral,...>>>},
the default implementation for \tcode{target_id(g,uv)} will return the \tcode{integral}. Additionally, if the caller does not override 
\tcode{vertex_id(g,u)}, the \tcode{integral} value will define the \tcode{vertex_id_t<G>} type.

Functions that have n/a for their Default Implementation must be defined by the author of a Graph Container implementation. 

Value functions (\tcode{graph_value(g)}, \tcode{vertex_value(g,u)} and \tcode{edge_value(g,uv)}) can be optionally implemented, 
depending on whether the graph container supports values on the graph, vertex and edge types. They return a single value and can 
be scalar, struct, class, union, or tuple. These are abstract types used by the GVF, VVF and EVF function objects to retrieve
values used by algorithms. As such it's valid to return the "enclosing" owning class (graph, vertex or edge), or some other
embedded value in those objects.

\tcode{find_vertex(g,uid)} is constant complexity because all algorithms in this proposal require that \tcode{vertex_range_t<G>} is a random access range. 

If the concept requirements for the default implementation aren't met by the graph container the function will need to be overridden.

\subsection{Determining the vertex\_id and its type}
To determine the type for \tcode{vertex_id_t<G>} the following steps are taken, in order, to determine its type.
\begin{enumerate}
    \item Use the type returned by \tcode{vertex_id(g,u)} when overridden for a graph.
    \item When the graph matches the pattern \tcode{random_access_range<forward_range<integral>>} or \\
          \tcode{random_access_range<forward_range<tuple<integral,...>>>},
          use the \tcode{integral} type specified, which is assumed to be the target\_id on an edge.
    \item Use \tcode{size_t} in all other cases.
\end{enumerate}

\tcode{vertex_id_t<G>} is defined by the type returned by \tcode{vertex_id(g)} and it defaults to the difference\_type of the underlying container used for vertices (e.g. int64\_t for 64-bit systems). 
This is sufficient for all situations. However, there are often space and performance advantages if a smaller type is used, such as int32\_t or even int16\_t. It is recommended to consider overriding 
this function for optimal results, assuring that it is also large enough for the number of possible vertices and edges in the application. It will also need to be overridden if the implementation doesn't 
expose the vertices as a range.

\tcode{vertex_id(g,u)} is evaluated in the context of a descriptor using the following rules:
\begin{enumerate}
    \item Use the value returned by \tcode{vertex_id(g,u)} when overridden for a graph.
    \item Use the index value on the descriptor.
\end{enumerate}

\subsection{Vertex and Edge Descriptor Views}
Graph containers may store vertices and edges in very different data structures: a
\tcode{vector}-based adjacency list uses integer indices as vertex identifiers while a
\tcode{map}-based graph uses the map key. Exposing raw iterators directly would require
concepts, algorithms, and user code to handle each storage strategy separately.
Descriptors solve this by providing a lightweight, opaque handle that works uniformly
regardless of the underlying container type. Table~\ref{tab:desc_vs_iter} compares raw
iterators with descriptors.

\begin{table}[h!]
\begin{center}
\resizebox{\textwidth}{!}
{\begin{tabular}{l L{5.5cm} L{7.0cm}}
\hline
    \textbf{Concern} & \textbf{Raw iterators} & \textbf{Descriptors} \\
\hline
    Abstraction & Expose container layout & Hide storage strategy --- one concept covers indexed and keyed graphs \\
    Overload reduction & Separate overloads for index vs.\ iterator & A single overload accepts the descriptor \\
    Invalidation & Invalidated by container mutation & Can be re-resolved via \tcode{find_vertex} \\
    Copyability & Varies (e.g.\ \tcode{forward_list::iterator}) & Always trivially copyable or cheap \\
    Size & May be pointer-sized or larger & Index-based: one integer \\
\hline
\hline
\end{tabular}}
\caption{Raw iterators vs.\ descriptors}
\label{tab:desc_vs_iter}
\end{center}
\end{table}

Index-based vertex descriptors provide an additional level of abstraction: they can
refer to a vertex even when no physical vertex object exists in the graph.
A \tcode{compressed_graph}, for example, stores only edges; there is no C++ object
corresponding to each vertex. Because the vertex is identified purely by its index, the
GCI's algorithms and views work identically whether or not the container materialises
explicit vertex storage. An iterator-based approach cannot offer this property --- an
iterator must point to an actual object.

CPOs automatically detect the storage pattern of an adjacency list and select the
appropriate descriptor strategy: \textbf{random-access} containers (\tcode{vector},
\tcode{deque}) produce index-based vertex descriptors; \textbf{associative} containers
(\tcode{map}, \tcode{unordered_map}) produce key-based (iterator) descriptors. Edge
patterns --- \tcode{int}, \tcode{pair<int,W>}, \tcode{tuple<int,...>}, or custom structs
--- are recognised automatically so that \tcode{target_id}, \tcode{source_id}, and
\tcode{edge_value} CPOs are provided without user specialisation.

The vertex and edge descriptors are defined as the \tcode{vertex_t<G>} and
\tcode{edge_t<G>} types, respectively. \tcode{vertex_descriptor<VertexIter>} stores
either a \tcode{size_t} index (for random-access containers) or an iterator (for
associative containers) as an exposition-only private member.
\tcode{edge_descriptor<EdgeIter,VertexIter,EdgeDirection>} additionally stores the
source \tcode{vertex_descriptor} and an \tcode{EdgeDirection} tag (\tcode{out_edge_tag}
or \tcode{in_edge_tag}) that controls source/target semantics, allowing bidirectional
graphs to use the same edge type for both outgoing and incoming edges.

Both descriptor types support equality comparison (\tcode{==}, \tcode{!=}), ordering
(\tcode{<}, \tcode{<=}, \tcode{>}, \tcode{>=} where supported), copy, assignment, and
default construction. They also provide \tcode{underlying_value(c)} and
\tcode{inner_value(c)} member function templates, where \tcode{c} is the owning
container, to retrieve a reference to the underlying element. These are needed only
when overriding CPOs to adapt an external graph container to the GCI.

Because descriptors are small and trivially copyable, CPO signatures accept them by
value. A \tcode{vertex_descriptor} is at most pointer-sized (one index or iterator).
An \tcode{edge_descriptor} is larger because it also carries the source
\tcode{vertex_descriptor}, but is still cheap to copy. The \tcode{const\&} form appears
only in \tcode{requires} clauses of concepts where value copies are not permitted.
{\small
      \lstinputlisting{D3130_Container_Interface/src/descriptor.hpp}
}

The following type traits identify whether a type is a vertex or edge descriptor.
{\small
      \lstinputlisting{D3130_Container_Interface/src/descriptor_traits.hpp}
}

The only vertex function that requires a vertex id \tcode{(uid)} is \tcode{find_vertex(g,uid).} All other functions
that accept vertex id are convenience functions that imply a call to \tcode{find_vertex(g,uid)} to get the vertex 
descriptor before a call.

The following are the descriptor views used by \tcode{vertices(g)} and \tcode{out_edges(g,u)}.
{\small
      \lstinputlisting{D3130_Container_Interface/src/descriptor_view.hpp}
}

\subsection{Unipartite, Bipartite and Multipartite Graph Representation}

\tcode{num_partitions(g)} returns the number of partitions, or partiteness, of the graph. It has a range of 1 to n, where 1 identifies 
a unipartite graph, 2 is a bipartite graph, and a value of 2 or more can be considered a multipartite graph. 

If a graph data structure doesn't support partitions then it is unipartite with one partition and partite functions will reflect that. 
For instance, \tcode{num_partitions(g)} returns a value of 1, and \tcode{vertices(g,0)} (vertices in the first partition) will return a range that includes all vertices in the graph.

A partition identifies a type of a vertex, where the vertex value types are assumed to be uniform in each partition. This creates a dilemma because 
the existing \tcode{vertex_value(g,u)} returns a single type based template parameter for the vertex value type. Supporting 
multiple types can be addressed in different ways using C++ features. The key to remember is that the actual value used by algorithms is done
by calling a function object that retrieves the value to be used. That function is specific to the graph data structure, using the partition to 
determine how to get the appropriate value.
\begin{itemize}
    \item
          \tcode{std::variant}: The lambda returns the appropriate variant value based on the partition.
    \item
          Base class pointer: The lambda can call a member function to return the value based on the partition.
    \item
          \tcode{void*}: The lambda can cast the pointer to a concrete type based on the partition, and
          then return the appropriate value.
\end{itemize}

\tcode{out_edges(g,uid,pid)} and \tcode{out_edges(g,u,pid)} filter the edges where the target is in the partition \tcode{pid} passed.
This isn't needed for bipartite graphs.

\section{Edgelist Interface}
An edgelist is a range of values where we can get the source\_id and target\_id, and an optional edge\_value. It is similar to edges in an
adjacency list or edges in the incidence view, but is a distinct range of values that are separate from the others.

Like the adjacency list, the edgelist has default implementations that use the standard library for simple implementations out of the box.
It's also able to easily adapt to externally defined edge types by overriding the \tcode{source_id(el,uv)}, \tcode{target_id(el,uv)} and 
\tcode{edge_value(el,uv)} functions.

\subsection{Namespace}
Edge list concepts, type aliases and CPOs are defined in the \tcode{std::graph::edge\_list} namespace.
Unlike the adjacency list, these symbols are \emph{not} re-exported into \tcode{std::graph} because
certain names have distinct meanings in each context --- most notably \tcode{vertex\_id\_t<G>} is derived
from \tcode{vertex\_id(g,u)} for an adjacency list but from \tcode{target\_id(el,uv)} for an edge list.
Bringing both into the root namespace would create an irresolvable ambiguity.

\tcode{graph::adj\_list} and \tcode{graph::edge\_list} are \emph{peer} namespaces: neither is a
subset or refinement of the other. An adjacency list provides per-vertex edge ranges and supports fast
neighbour traversal; an edge list is a flat range of \tcode{(source, target [, value])} tuples suited to
edge-centric algorithms such as Kruskal's MST or bulk construction. CPOs that operate directly on
edge data --- \tcode{source\_id}, \tcode{target\_id} and \tcode{edge\_value} --- are shared and
live in \tcode{std::graph}.


\subsection{Concepts}
The edgelist concepts use a 2-argument form \tcode{f(el, uv)} where \tcode{el} is the edgelist and \tcode{uv} is the edge value,
matching the adjacency list CPO convention. All edgelist CPOs require both the container and the element to allow for
container-specific dispatch.

\tcode{basic_sourced_edgelist<EL>} is the base concept: an input range whose element type is not itself a range, supporting
\tcode{source_id(el,uv)} and \tcode{target_id(el,uv)} with compatible types. \tcode{basic_sourced_index_edgelist<EL>} additionally
requires both ids to be integral. \tcode{has_edge_value<EL>} refines \tcode{basic_sourced_edgelist} to also provide \tcode{edge_value(el,uv)}.
{\small
     \lstinputlisting{D3130_Container_Interface/src/edgelist_concepts.hpp}
}
\subsection{Traits}
Table \ref{tab:edgelist_traits} summarizes the type traits in the Edgelist Interface, allowing views and algorithms to query the graph's characteristics.

\begin{table}[h!]
\begin{center}
%\resizebox{\textwidth}{!}
{\begin{tabular}{l l L{7.0cm}}
\hline
    \textbf{Trait} & \textbf{Type} & \textbf{Comment} \\
\hline
    \tcode{is_directed<EL> : false_type} & struct & When specialized for an edgelist to derive \\
    \tcode{is_directed_v<EL>}          & & from \tcode{true_type}, it may be used during graph 
                                                     construction to add a second edge with source\_id and target\_id reversed. \\
\hline
\end{tabular}}
\caption{Graph Container Interface Type Traits}
\label{tab:edgelist_traits}
\end{center}
\end{table}

%{\small
%     \lstinputlisting{D3130_Container_Interface/src/edgelist_typetraits.hpp}
%}
\subsection{Types}
Table \ref{tab:edgelist_type} summarizes the type aliases in the Edgelist Interface.

The type aliases are defined by either a function specialization for the edgelist implementation, or a refinement of one of 
those types (e.g. an iterator of a range). Table \ref{tab:edgelist_func} describes the functions in more detail.

\tcode{edge_value(el,uv)} can be optionally implemented, depending on whether or not the edgelist has values on the edge types.

\begin{table}[h!]
\begin{center}
%\resizebox{\textwidth}{!}
{\begin{tabular}{l l L{1.7cm}}
\hline
    \textbf{Type Alias} & \textbf{Definition} & \textbf{Comment} \\
\hline
    \tcode{edge_range_t<EL>} & \tcode{EL} & \\    
    \tcode{edge_iterator_t<EL>} & \tcode{iterator_t<edge_range_t<EL>>} & \\    
    \tcode{edge_t<EL>} & \tcode{range_value_t<edge_range_t<EL>>} & \\       
    \tcode{edge_value_t<EL>} & \tcode{decltype(edge_value(el,uv))} & optional \\
\hline
    \tcode{vertex_id_t<EL>} & \tcode{remove_cvref_t<decltype(target_id(el,uv))>} & \\
    \textit{\tcode{raw-vertex-id-type}} & \tcode{decltype(target_id(el,uv))} & exposition only \\
\hline
\end{tabular}}
\caption{Edgelist Interface Type Aliases}
\label{tab:edgelist_type}
\end{center}
\end{table}

%{\small
%     \lstinputlisting{D3130_Container_Interface/src/edgelist_types.hpp}
%}
\subsection{Functions}
Table \ref{tab:edgelist_func} shows the functions available in the Edgelist Interface. Unlike the adjacency list, \tcode{source_id(el,uv)}
is always available.


\begin{table}[h!]
    \begin{center}
    \resizebox{\textwidth}{!}
    {\begin{tabular}{l l p{1.5cm} L{6.7cm}}
    \hline
        \textbf{Function} & \textbf{Return Type} & \textbf{Cmplx} & \textbf{Default Implementation} \\
    \hline
        \tcode{target_id(el,uv)} & \tcode{vertex_id_t<EL>} & constant & (see below) \\
        \tcode{source_id(el,uv)} & \tcode{vertex_id_t<EL>} & constant & (see below) \\
        \tcode{edge_value(el,uv)} & \tcode{edge_value_t<EL>} & constant & optional, see below \\
    \hdashline
        \tcode{contains_edge(el,uid,vid)} & \tcode{bool} & linear & \tcode{find_if(el, [](auto\&\& uv) \{} \\
                                          &              &        & \hspace{7mm}\tcode{return source\_id(el,uv)==uid} \\
                                          &              &        & \hspace{13.5mm}\tcode{\&\& target\_id(el,uv)==vid; \} )} \\
        \tcode{num_edges(el)} & \tcode{integral} & constant & \tcode{size(el)} \\
        \tcode{has_edges(el)} & \tcode{bool} & constant & \tcode{num_edges(el)>0} \\
    \hline
    \end{tabular}}
    \caption{Edgelist Interface Functions}
    \label{tab:edgelist_func}
    \end{center}
\end{table}

% \tcode{find_if(el, )}
% [](edge_reference_t<EL>) \{return source_id(e)==uid && target_id(e)==vid\}

\subsection{Determining the source\_id, target\_id and edge\_value types}
Special patterns are recognized for edges based on the \tcode{pair}, \tcode{tuple} and \tcode{edge_data} types. When they are used the \tcode{source_id(el,uv)},
\tcode{target_id(el,uv)} and \tcode{edge_value(el,uv)} functions will be defined automatically.

The \tcode{pair} pattern is
\begin{itemize}
    \item \tcode{pair<integral,integral>} for \tcode{source_id(el,uv)} and \tcode{target_id(el,uv)} respectively.
\end{itemize}

The \tcode{tuple} patterns are
\begin{itemize}
    \item \tcode{tuple<integral,integral>} for \tcode{source_id(el,uv)} and \tcode{target_id(el,uv)} respectively.
    \item \tcode{tuple<integral,integral,scalar>} for \tcode{source_id(el,uv)}, \tcode{target_id(el,uv)} and \tcode{edge_value(el,uv)} respectively.
\end{itemize}

The \tcode{edge_data} patterns are
\begin{itemize}
    \item \tcode{edge_data<VId,true,void,void>} with \tcode{source_id(el,uv)} and \tcode{target_id(el,uv)}.
    \item \tcode{edge_data<VId,true,void,EV>} with \tcode{source_id(el,uv)}, \tcode{target_id(el,uv)} and \tcode{edge_value(el,uv)}.
    \item \tcode{edge_data<VId,true,E\&,void>} with \tcode{source_id(el,uv)}, \tcode{target_id(el,uv)} and an edge descriptor reference.
    \item \tcode{edge_data<VId,true,E\&,EV>} with \tcode{source_id(el,uv)}, \tcode{target_id(el,uv)}, an edge descriptor reference and \tcode{edge_value(el,uv)}.
\end{itemize}

In all other cases the functions will need to be overridden for the edge type.

\section{Using Existing Data Structures}
Reasonable defaults have been defined for the adjacency list and edgelist functions to minimize the amount of work
needed to adapt existing data structures to be used by the views and algorithms.

Useful defaults have been created using types and containers in the standard library, with the ability
to override them for external data structures. This is described in more detail in the paper for Graph Library 
Containers.

\subsection{Recognized Vertex Patterns}

When a graph type \tcode{G} is itself a \tcode{forward_range} whose elements are also ranges, the
GCI CPOs can detect and adapt to the storage strategy automatically. Two broad categories of vertex
container are recognized, each producing a different descriptor and vertex-id strategy.

\subsubsection{Random-Access Vertex Pattern}
When \tcode{G} satisfies \tcode{random_access_range} and its elements satisfy \tcode{forward_range},
the vertices are assumed to reside in a contiguous or random-access container such as
\tcode{std::vector} or \tcode{std::deque}.

\begin{itemize}
    \item \tcode{vertex_id_t<G>} defaults to an integral type (index into the range).
    \item \tcode{find_vertex(g, uid)} uses indexed access: \tcode{begin(vertices(g)) + uid},
          giving $\mathcal{O}(1)$ lookup.
    \item \tcode{out_edges(g, u)} returns the inner range at that index.
\end{itemize}

Examples of standard containers matching this pattern:
\begin{itemize}
    \item \tcode{vector<vector<int>>} --- vertices in a \tcode{vector}, edges as \tcode{int} target ids.
    \item \tcode{vector<list<pair<int,double>>>} --- vertices in a \tcode{vector}, edges as weighted pairs in a \tcode{list}.
    \item \tcode{deque<forward_list<int>>} --- vertices in a \tcode{deque}, edges in a \tcode{forward_list}.
\end{itemize}

\subsubsection{Associative Vertex Pattern}
When \tcode{G} is an associative container (\tcode{std::map} or \tcode{std::unordered_map})
whose mapped values are forward ranges, vertices are keyed by an arbitrary id type rather than
a dense index.

\begin{itemize}
    \item \tcode{vertex_id_t<G>} defaults to the \tcode{key_type} of the container.
    \item \tcode{find_vertex(g, uid)} uses \tcode{g.find(uid)}, giving
          $\mathcal{O}(\log |V|)$ for \tcode{map} or amortized $\mathcal{O}(1)$ for
          \tcode{unordered_map}.
    \item \tcode{out_edges(g, u)} extracts the \tcode{.second} member (the mapped edge
          range) from the key-value pair.
\end{itemize}

Examples of standard containers matching this pattern:
\begin{itemize}
    \item \tcode{map<int, vector<int>>} --- sparse vertex ids in a sorted \tcode{map}, edges as target ids.
    \item \tcode{unordered_map<int, vector<pair<int,double>>>} --- hash-based vertex lookup, weighted edges.
\end{itemize}

\subsubsection{Edge Container Patterns}
The edge range returned by \tcode{out_edges(g, u)} is the inner range stored at (or
referenced by) each vertex. Any standard container satisfying \tcode{forward_range}
whose element type matches a recognized edge element pattern (described below) is
accepted automatically. The library does \emph{not} maintain a fixed list of containers;
it relies on concept-based detection.

Table~\ref{tab:edge_container_patterns} groups the standard containers by category and
summarises their trade-offs when used as edge containers.

\begin{table}[h!]
\begin{center}
\resizebox{\textwidth}{!}
{\begin{tabular}{l l L{7.5cm}}
\hline
    \textbf{Category} & \textbf{Container} & \textbf{Characteristics} \\
\hline
    Sequence &
    \tcode{vector<E>} &
    Random-access, cache-friendly, amortized $\mathcal{O}(1)$ push-back. \\
    &
    \tcode{deque<E>} &
    Random-access, stable references on push-back. \\
    &
    \tcode{list<E>} &
    Bidirectional, $\mathcal{O}(1)$ splice, stable iterators. \\
    &
    \tcode{forward_list<E>} &
    Forward-only, minimal overhead, $\mathcal{O}(1)$ push-front. \\
\hdashline
    Ordered associative &
    \tcode{set<E>} &
    Sorted, automatic deduplication, $\mathcal{O}(\log n)$ lookup; element is the edge itself. \\
    &
    \tcode{map<VId, EV>} &
    Sorted by target id, $\mathcal{O}(\log n)$ lookup, no parallel edges; \tcode{.first} is
    target id, \tcode{.second} is edge value. \\
\hdashline
    Unordered associative &
    \tcode{unordered_set<E>} &
    Hash-based deduplication, amortized $\mathcal{O}(1)$ insert/lookup. \\
    &
    \tcode{unordered_map<VId, EV>} &
    Hash-based by target id, amortized $\mathcal{O}(1)$ lookup, no parallel edges;
    \tcode{.first} is target id, \tcode{.second} is edge value. \\
\hline
\end{tabular}}
\caption{Recognized Edge Container Patterns}
\label{tab:edge_container_patterns}
\end{center}
\end{table}

For associative containers whose \tcode{value_type} is a \tcode{pair<const Key, Mapped>}
(i.e.\ \tcode{map} and \tcode{unordered_map}), the key is
treated as the target vertex id and the mapped type as the edge value. This means that
\tcode{target_id(g,uv)} returns \tcode{.first} and \tcode{edge_value(g,uv)} returns
\tcode{.second}, following the same pair convention described below.

For set-like containers (\tcode{set}, \tcode{unordered_set}), each element is the edge
itself. If the element is integral it serves as the target id directly; if it is a
\tcode{pair} or \tcode{tuple} the element-level rules below apply.

\subsubsection{Edge Element Patterns}
Within each vertex's edge range the GCI recognizes the following element forms and
automatically provides \tcode{target_id(g,uv)} and, where applicable,
\tcode{edge_value(g,uv)}:

\begin{table}[h!]
\begin{center}
{\begin{tabular}{l L{8.5cm}}
\hline
    \textbf{Edge Element Type} & \textbf{Automatic CPOs} \\
\hline
    \tcode{integral} & \tcode{target_id(g,uv)} returns the value directly.
                        No \tcode{edge_value}. \\
    \tcode{pair<integral, EV>} & \tcode{target_id(g,uv)} returns \tcode{.first};
                                  \tcode{edge_value(g,uv)} returns \tcode{.second}. \\
    \tcode{tuple<integral, ...>} & \tcode{target_id(g,uv)} returns \tcode{get<0>(uv)};
                                    remaining elements accessible via \tcode{edge_value}. \\
    custom struct &
    The user overrides \tcode{target_id(g,uv)} and optionally
    \tcode{edge_value(g,uv)} for the type. \\
\hline
\end{tabular}}
\caption{Recognized Edge Element Patterns}
\label{tab:edge_element_patterns}
\end{center}
\end{table}

\noindent
These vertex and edge patterns combine freely: any recognized vertex container can hold
any recognized edge container, and any recognized edge container can hold any recognized
edge element type. For example,
\tcode{vector<vector<pair<int,double>>>} is a random-access graph with weighted edges,
\tcode{vector<set<int>>} is a random-access graph with sorted deduplicated unweighted edges,
\tcode{vector<map<int,double>>} is a random-access graph with $\mathcal{O}(\log n)$ edge lookup by target,
and \tcode{map<string, list<int>>} is an associative graph with unweighted edges.
When the element type does not match a recognized pattern, the user must override the
appropriate CPOs for the graph type.
