\section{Revision History}

\subsection*{\paperno r0}

\begin{itemize}
      \item Split from P1709r5. Added \textit{Getting Started} section.
      \item Add default implementation for \tcode{target_id(g,uv)} when the graph type matches the 
            pattern \tcode{random_access_range<forward_range<integral>>} or \tcode{random_access_range<forward_range<tuple<integral,...>>>};
            \tcode{vertex_id_t<G>} also defaults to the \tcode{integral} type given.
      \item Revised concept definitions, adding \tcode{sourced_targeted_edge} and \tcode{target_edge_range}, and replaced
            summary table with code for clarity. Also assured that all combinations of adjacency list concepts for
            \textit{basic}, \textit{sourced} and \textit{index} exist.
      \item Move text for graph data structures created from std containers from Graph Container Interface to 
            Container Implementation paper.
      \item Identify all \tcode{concept} definitions as "For exposition only" until we have consensus of whether they 
            belong in the standard or not.
\end{itemize}

\subsection*{\paperno r1}

\begin{itemize}
      \item Add \tcode{num_edges(g)} and \tcode{has_edge(g)} functions. Split function table into 3 tables for graph,
            vertex and edge functions because it was getting too big.
      \item Removed the Load Graph Data section with it's load functions from \href{https://www.wg21.link/P3130}{P3130 Graph Container Interface}
            because it unnecessarily complicates the interface with constructors for graph data structures. To complement this, constructors have 
            been added for \tcode{compressed_graph} in \href{https://www.wg21.link/P3131}{P3131 Graph Containers}.
      \item Revised partition functions after implementation in \tcode{compressed_graph} to reflect usage, including: 
            renaming \tcode{partition_count(g)} 
            to \tcode{num_partitions(g)} to match other names used, changed \tcode{partition_id(g,u)} to \tcode{partition_id(g,uid)}
            because vertices may not exist when the function is called, and removing \tcode{edges(g,u,pid)} because it can easily be 
            implemented as a filter using ranges functionality when target vertices can be in different partitions.
\end{itemize}

\subsection*{\paperno r2}

\begin{itemize}
      \item Add the edgelist as an abstract data structure as a peer to the adjacency list. 
            This causes a reorganization of this paper and the addition of a new section for the edgelist.
      \item Remove unnecessary \tcode{E} edge template parameter in concepts. 
      \item Remove type traits \tcode{is_unordered_edge} and \tcode{is_ordered_edge} because 
            their matching concepts, \tcode{unordered_edge} and \tcode{ordered_edge}, don't need them.
      \item Remove \tcode{edge_id(g,uv)} and \tcode{edge_id_t<G>} because they don't add value to the interface
            and can easily be implemented if needed.
      \item Added description of why the return type isn't validated for \tcode{target_id(g,uv)} in the 
            \tcode{basic_targeted_edge} concept.
\end{itemize}

\subsection*{\paperno r3}

\begin{itemize}
      \item Introduction of new boost::graph-like descriptors with the following changes:
            \begin{itemize}
                  \item Concrete vertex and edge types defined by a graph container are replaced with absract 
                        vertex and edge descriptors. This allows the graph container to be decoupled from the 
                        underlying container type. This enables future expansion by allowing a graph container 
                        that uses associative containers with minimal impact to the interface.

                        While this is a major revision to the implementation, it isn't a big conceptual change to the 
                        user. The visible changes include replacing \tcode{vertex_id(g,ui)} with \tcode{vertex_id(g,u),}
                        the addition of \tcode{partition_id(g,ui),}
                        and \tcode{vertices(g)} and \tcode{edges(g,u)} now return a \tcode{descriptor_view} instead 
                        of a range of the underlying container.
                  \item A descriptor replaces the combination of id and reference parameters from the previous version 
                        of the interface. This reduces the number of concepts by the removal of the "basic" name qualifiers 
                        for vertex\_id-only concepts. It also enables consolidation of the views in \href{https://www.wg21.link/P3129}{P3129},
                        where the "basic" views are no longer needed, reducing the number of view functions in half.
                  \item The previous \tcode{descriptor} structs in \href{https://www.wg21.link/P3129}{P3129 Views} 
                        have been renamed to \tcode{info} structs to avoid name clashes. Additionally, the "copyable" 
                        type aliases for the \tcode{info} structs are no longer needed because reference was replaced
                        with \tcode{descriptor}, which is now copyable.
            \end{itemize}
\end{itemize}

\subsection*{\paperno r4}

\begin{itemize}
      \item Adopt the new split-descriptor architecture: separate \tcode{vertex_descriptor<VertexIter>} and
            \tcode{edge_descriptor<EdgeIter,VertexIter,EdgeDirection>} types, each with exposition-only internals.
            Descriptor traits (\tcode{is_vertex_descriptor_v}, \tcode{is_edge_descriptor_v}, etc.) are defined in a
            dedicated header. \tcode{vertex_descriptor_view} and \tcode{edge_descriptor_view} replace the
            former \tcode{descriptor_view}.
      \item Simplify the concept hierarchy: the single \tcode{edge<G,E>} concept replaces the previous
            \tcode{targeted_edge}/\tcode{sourced_edge} split; \tcode{out_edge_range<R,G>} and
            \tcode{in_edge_range<R,G>} replace \tcode{targeted_edge_range}. Six adjacency list concepts
            cover the index/mapped/bidirectional axes; \tcode{basic_*} and \tcode{sourced_*} variants are removed.
            \tcode{mapped_vertex_range} is added alongside \tcode{index_vertex_range}.
      \item CPO renames: \tcode{edges(g,u)} $\rightarrow$ \tcode{out_edges(g,u)} (\tcode{edges} kept as
            backward-compatible alias); \tcode{degree(g,u)} $\rightarrow$ \tcode{out_degree(g,u)};
            \tcode{find_vertex_edge} $\rightarrow$ \tcode{find_out_edge};
            \tcode{contains_edge} (adj list) $\rightarrow$ \tcode{contains_out_edge}. In-edge CPOs
            (\tcode{in_edges}, \tcode{in_degree}, \tcode{find_in_edge}, \tcode{contains_in_edge}) are added.
      \item Type alias updates: \tcode{vertex_edge_range_t<G>} / \tcode{edge_t<G>} noted as aliases for
            \tcode{out_edge_range_t<G>} / \tcode{out_edge_t<G>}. New: \tcode{out_edge_range_t},
            \tcode{out_edge_t}, \tcode{in_edge_range_t}, \tcode{in_edge_t}, \tcode{vertex_id_store_t<G>},
            \textit{raw-vertex-id-type}. Removed: \tcode{graph_reference_t}, \tcode{vertex_reference_t},
            \tcode{edge_reference_t}.
      \item New data structures: \tcode{vertex_data}, \tcode{edge_data} and \tcode{neighbor_data} aggregates
            with partial-specialisation families; \tcode{copyable_vertex_t}, \tcode{copyable_edge_t},
            \tcode{copyable_neighbor_t} helper aliases.
      \item New concepts: \tcode{vertex_value_function} and \tcode{edge_value_function} (value-function
            concepts for views and algorithms).
      \item New utility: \tcode{vertex_property_map<G,T>} type alias; \tcode{make_vertex_property_map}
            (eager and lazy), \tcode{vertex_property_map_contains}, \tcode{vertex_property_map_get}.
      \item Edgelist namespace renamed from \tcode{std::graph::edgelist} to \tcode{std::graph::edge_list}.
            Adjacency list symbols now formally reside in \tcode{std::graph::adj_list} and are re-exported
            to \tcode{std::graph}; the two sub-namespaces are peers.
      \item Edgelist CPOs moved to 2-argument form \tcode{f(el,uv)}; \tcode{edge_info} renamed to
            \tcode{edge_data}; \tcode{edge_reference_t<EL>} removed; \tcode{raw-vertex-id-type} added.
\end{itemize}
