\section{Revision History}

\subsection*{\paperno r0}

\begin{itemize}
      \item Split from P1709r5. Added \textit{Getting Started} section.
      \item Removed allocator parameters on views, for consistency with existing views in the standard. 
\end{itemize}

\subsection*{\paperno r1}
\begin{itemize}
      \item Add the edgelist as an abstract data structure as a peer to the adjacency list. 
            The range returned by \tcode{edgelist_view} adheres to the \tcode{basic_sourced_index_edgelist} concept,
            and to the \tcode{has_edge_value} concept if a \tcode{evf(uv)} function is passed. The same applies to
            all \textit{sourced} versions of the BFS, DFS and Topological Sort views.
      \item Restore the allocator parameters on the DFS, BFS and Topological Sort views, based on feedback and 
            by SG14/SG19 joint meeting.
      \item Add a note that we will be unable to support a freestanding graph library in this proposal because
            of the need for \tcode{stack}, \tcode{queue} and potential \tcode{bad_alloc} exception in many of 
            the views.
      \item Rename \tcode{descriptor} structs to \tcode{info} structs in preparation for new BGL-like descriptors.
\end{itemize}

\subsection*{\paperno r3}
\begin{itemize}
    \item Renamed \tcode{vertex_info}, \tcode{edge_info}, and \tcode{neighbor_info} to
          \tcode{vertex_data}, \tcode{edge_data}, and \tcode{neighbor_data} respectively,
          with the addition of separate vertex ID and vertex descriptor template parameters.
    \item Updated all value functions from 1-arg \tcode{vvf(u)}/\tcode{evf(uv)} form to
          2-arg \tcode{vvf(g,u)}/\tcode{evf(g,uv)} form, taking the graph as the first
          parameter.
    \item Added \tcode{basic\_} view variants (\tcode{basic\_vertexlist},
          \tcode{basic\_incidence}, \tcode{basic\_neighbors}, \tcode{basic\_edgelist}) that
          yield only vertex/edge IDs without descriptors.
    \item Added incoming-edge view variants (\tcode{in\_incidence},
          \tcode{in\_neighbors}) for bidirectional graphs.
    \item Added \tcode{transpose} view for direction-swapping graph adaption.
    \item Added pipe syntax support via range adaptors for all view functions.
    \item Added \tcode{\_safe} topological sort factories with cycle detection
          via \tcode{std::expected} (C++23; backward compatible to C++20 via external
          \tcode{expected} library).
    \item Changed topological sort from seeded (single source) to all-vertex traversal.
    \item Removed \tcode{sourced\_edges\_dfs}, \tcode{sourced\_edges\_bfs}, and
          \tcode{sourced\_edges\_topological\_sort} --- sourced search variants are no longer
          in the implementation.
    \item Changed search view control from free functions (\tcode{cancel()}, \tcode{depth()},
          \tcode{size()}) to member functions (\tcode{.cancel()}, \tcode{.depth()},
          \tcode{.num\_visited()}).
    \item Added \tcode{search\_view} concept.
    \item Replaced all uses of \tcode{vertex\_reference\_t<G>} with \tcode{vertex\_t<G>}
          and \tcode{edge\_reference\_t<G>} with \tcode{edge\_t<G>}, consistent with the
          descriptor-based architecture defined in D3130.
    \item Documented allocator parameters for search views (DFS, BFS, topological sort).
\end{itemize}

\subsection*{\paperno r2}
\begin{itemize}
      \item Replace the use of \textit{id} and \textit{reference} with \textit{descriptor}, leading to a simpler
            interace. It also creates a more flexible interface that can support associative containers in the future. 
            The following changes were made:
            \begin{itemize}
                  \item The number of View functions has been halved because we no longer need separate functions that 
                        only have \tcode{id}, and another set that has has both \tcode{id} and \tcode{reference}. Only functions
                        with \tcode{descriptor} are needed.
                  \item The \tcode{vertex_id} member has been removed from the \tcode{vertex_info} struct, and the \tcode{vertex}
                        member can hold either an id or a descriptor, depending on the context it's used. The same changes have 
                        also been applied to the \tcode{edge_info} and \tcode{neighbor_info} structs.
                  \item The copyable info type aliases and concepts have been removed. \tcode{vertex_info}, \tcode{edge_info} 
                        and \tcode{neighbor_info} are always copyable because they no longer contain references.
                  \item See \href{https://www.wg21.link/P3130}{P3130 Graph Container Interface} for more details about
                        descriptors.
            \end{itemize}
\end{itemize}
