\section{Revision History}

\subsection*{\paperno r0}

\begin{itemize}
      \item Split from P1709r5. Added \textit{Getting Started} section.
      \item Rewrite \textit{Goals and Priorities} section to reflect the structure of the papers and to include a section on our \textit{Future Roadmap}.
      \item Added \textit{Notes and Considerations} section.
      \item Concepts will be identified as "For exposition only" until we have consensus of whether they belong in the standard or not.
\end{itemize}

\subsection*{\paperno r1}
\begin{itemize}
      \item {Added Issues Status section to be open with the issues that have been reported and that we are working on.}
\end{itemize}

\subsection*{\paperno r2}
\begin{itemize}
      \item Add the edgelist as an abstract data structure as a peer to the adjacency list. 
            This completes an open issue for completing the definition of the edgelist.
      \item Added the \tcode{std::graph::edgelist} namespace for edgelist concepts, traits and types
            to keep identically named types separate from those for adjacency lists.
      \item Added a reference to the new \href{https://www.wg21.link/P3337}{P3337} \textit{Graph Comparisons} paper
            to the Getting Started section.
      \item Update text to make it clear parallel algorithms will \textit{not} be included in the proposal.
\end{itemize}

\subsection*{\paperno r3}
\begin{itemize}
      \item Change the reference implementation from the \tcode{std::graph} to the \tcode{graph} namespace. This will
            make it more accessible to the community and allow for easier experimentation outside of this proposal.
      \item Update the status on supporting more versatile BFS and DFS algorithms.
      \item Add additional motivation for a graph library in the Overview section.
      \item Extend the Six Degrees of Kevin Bacon example to include the output and additional description.
\end{itemize}
